### 2016-06-06 doc/om-paper/paper.tex
model-checking paper: first commit.

\documentclass{IEEEtran/IEEEtran}

\usepackage{ifthen}
\usepackage{xcolor}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}

\newtheorem{definition}{Definition}

\newboolean{submission}  %set to true for the submission version
\setboolean{submission}{false}
%\setboolean{submission}{true}
\ifthenelse
{\boolean{submission}}
{ %
  \newcommand{\lee}[1]{ } %
  \newcommand{\ben}[1]{ } %
  % Put your own TODO macros here and below
} %hide todo
{ %
  \newcommand{\lee}[1]{ {\color{blue}$<$lee: #1$>$} } %
  \newcommand{\ben}[1]{ {\color{purple}$<$ben: #1$>$} } %
  \usepackage{hyperref} %
  \usepackage[inline]{showlabels} %
}

\begin{document}

\title{Towards Systematic Specification and Verification of Fault-Tolerant Protocols}
\author{Ben and Lee}

\maketitle

\begin{abstract}
  Some abstract.
\end{abstract}

\section{Introduction}

Fault-tolerant distributed systems are essential to life-critical systems, like avionics. Consequently, this class of systems demand high-assurance of correct design and implementation. Formal verification can be used to prove correctness.

There is substantial literature on verifying fault-tolerant distributed systems~\cite{}. Some of this work uses mechanical theorem-proving~\cite{}. Two problems with mechanical verification are that it requires specialized expertise and it is labor intensive. In contrast, model-checking is automated but has limited scalability. To address scalability, abstractions and optimizations are employed, both of which lead to problems. Verification abstractions are often ad-hoc and system-specific. For example, to model a node that can broadcast the values $\{0, 1\}$, or no value if when faulty, one might define an enumerated type of broadcast values sent such as

$$\{0, 1, NONE\}$$

However, the enumeration conflates environment artifacts (i.e., faults) and system artifacts. Conflating them makes formal model-to-code correspondence more difficult to achieve. An example of a verification optimization is to model message passing as shared state between nodes. While these optimizations might be safe under particular assumptions, they can lead to unsound models when a system or its environment is modified.

\paragraph{Contributions}

Our work combines a number of ideas in the formal verification literature to build scalable, generic formal models of fault-tolerant distributed systems.

We describe a generic formal model for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}. The model combines several formalisms from the literature to efficiently but systematically model and verify a broad class of systems. The model extends \emph{calendar automata}, originally developed by Dutetre and Sorea~\cite{}. A calendar automata is a model inspired by discrete-event simulation, used in testing real-time systems~\cite{}. Calendar automata have previously been used to verify bounded asynchronous systems~\cite{}; here we verify a synchronous system.

The model also employs \emph{synchronous observers}~\cite{}. Synchronous observers are a modeling abstraction in which an ``observer'' actor is synchronously composed with the system under verification. We use synchronous observers to impose fault model constraints on the system. Doing so makes the fault-model compositional with the system and easy to modify. We also use synchronous observers to observe safety properties, eliding the need for complex linear temporal logic specifications. Finally, we use an \emph{abstract state machine}~\cite{}, also composed with the system under verification using synchronous observation. The abstract state machine simplifies discovering invariants and debugging counterexamples.

We demonstrate the model with an extended case-study, verifying the Oral Messages (1) protocol (OM(1))~\cite{}. While our model is more detailed---it explicitly models channels---and also more general---it allows bounded asynchrony between nodes---it is the most scalable model-checking verification of the protocol.

\lee{finish intro...}

 %% Timed automata are an alternative to the \emph{timed automata} formalism~\cite{}, as implemented in model-checkers like UPPAAL~\cite{} and Kronos~\cite{}.

% ------------------------------------------------------------
\section{Formal Model}\label{sec:model}
Here we describe our formal model specialized for fault-tolerant distributed systems. There are three principal abstractions we describe: calendar automata, synchronous observers, and abstract state machines; we describe each below.

\subsection{Calendar Automata}\label{sec:calendar}
\emph{Timed automata} are specialized formalisms for verifying real-time systems~\cite{}. Time automata assume the existence of continuously-varying real-valued variables, known as \emph{real-time clocks} in a solver. While successful timed automata model-checkers exist, they are specialized for verifying real-time systems in which the only infinite-valued variables are the real-time clocks. Modern general-purpose model-checkers like \lee{name some} use SMT solvers for general infinite-state system verification that is not limited to real-time clocks.

However, real-time system verification in general-purpose model-checkers requires an explicit formalism of real-time progression. Trying to encode real-time clocks directly is difficult; in particular, one must avoid Zeno's paradox in which no progress is made because state transitions simply update real-valued variables by an infinitesimally small amount~\cite{bruno,lamport}. To avoid this problem, Dutetre and Sorea developed \emph{calendar automata}~\cite{bruno}, which is itself inspired by event calendars used in discrete-event simulation.

Calendar automata avoid the problems of encoding timed automata. Rather than encoding ``how much time has passed since the last event'', it encodes ``how far into the future is the next scheduled event'', and a real-valued variable representing the current time is updated to the next event time. Doing so avoids Zeno updates.

Define a set of \emph{events} $e_0, e_1, \ldots, e_n \in E$. We do not define events; intuitively, an event is a set of state variables. When an event is \emph{enabled}, the transitions over events are enabled; otherwise, the variables stutter and maintain the same value.

\begin{definition}[Event Calendar and Calender Events]
An \emph{event calendar} $\{ <e_0, t_0>, <e_1, t_1>, \ldots, <e_n, t_n> \}$ is a finite set of ordered pairs $<e_i, t_i>$ called \emph{calendar events} where $e_i \in E$ is an event and $t_i \in \mathcal{R}$ is the time at which the event is scheduled.
\end{definition}

For an event calendar $c$, $c_i \in c = <e_i, t_i>$. Define an ordering on event calendars such that $<e_i, t_i> \leq <e_j, t_j>$ iff $t_i \leq t_j>$, and $min(c) = \{ c_i \| \forall c_j \in c, \, c_i \leq c_j  \}$ are the minimum elements of $c$.

Let a transition system $\mathcal{M} = <S, I, \rightarrow>$, be a set of states $S$, a set of initial states $I \subseteq S$, and a transition relation $\rightarrow \subseteq S \times S$. We implicitly assume a set of state variables such that a state $\sigma \in S$ is a total function that maps state variables to values.

We distinguish two special state variables: (1) $now \in \mathcal{R}$ denotes the current time in the state, and (2) $c$ is an event calendar.

The following laws must hold of a calendar automaton:

\begin{itemize}
\item Time is initialized to 0 in the initial state: $\forall \sigma \in I$, $\sigma(now) = 0$.
\item In all states, every calendar event occurs no sooner than the current time: $\forall \sigma \in S, \, <e_i, t_i> \in C$, $\sigma(t_i) /geq \sigma(now)$.
\item In all states, if the current time is strictly less than every calendar event, then the only enabled transition is a \emph{time progress} update: $\forall \sigma \in S \, <e_i, t_i> \in C$, if $\sigma(now) < t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma' = \sigma \backslash now = min$
\end{itemize}

\emph{give theorems for calendar, describe calendar for messages---all you need}

\subsection{Synchronous Observers}\label{sec:sync}

\paragraph{Fault model}

\paragraph{Property Observer}

\subsection{Abstract State Machines}\label{sec:abstract}

% ------------------------------------------------------------
\section{Experimental Results}\label{sec:experimental}

\subsection{Byzantine Agreement}\label{sec:byz}

% ------------------------------------------------------------
\section{Conclusions}\label{sec:conclusions}

\end{document}

### 2016-06-07 doc/om-paper/paper.tex
more wibbles on calendar automata formalism.

@@
 
-\begin{definition}[Event Calendar and Calender Events]
-An \emph{event calendar} $\{ <e_0, t_0>, <e_1, t_1>, \ldots, <e_n, t_n> \}$ is a finite set of ordered pairs $<e_i, t_i>$ called \emph{calendar events} where $e_i \in E$ is an event and $t_i \in \mathcal{R}$ is the time at which the event is scheduled.
-\end{definition}
+An \emph{event calendar} $\{ (e_0, t_0), (e_1, t_1), \ldots, (e_n, t_n) \}$ is a finite set of ordered pairs $(e_i, t_i)$ called \emph{calendar events} where $e_i \in E$ is an event and $t_i \in \mathcal{R}$ is a \emph{timeout}, the time at which the event is scheduled.
+We denote element $(e_1, t_1)$ of an event calendar by $c_i$.
 
-For an event calendar $c$, $c_i \in c = <e_i, t_i>$.
-Define an ordering on event calendars such that $<e_i, t_i> \leq <e_j, t_j>$ iff $t_i \leq t_j>$, and $min(c) = \{ c_i \| \forall c_j \in c, \, c_i \leq c_j  \}$ are the minimum elements of $c$.
+Let $c$ be an event calendar and $c_i, c_j \in c$ be calendar events.
+Define an ordering on event calendars such that $c_i \leq c_j$ iff $t_i \leq t_j$, and $min(c) = \{ c_i | \forall c_j \in c, \, c_i \leq c_j  \}$ are the minimum elements of $c$.
 
-Let a transition system $\mathcal{M} = <S, I, \rightarrow>$, be a set of states $S$, a set of initial states $I \subseteq S$, and a transition relation $\rightarrow \subseteq S \times S$.
+Let a transition system $\mathcal{M} = (S, I, \rightarrow)$, be a set of states $S$, a set of initial states $I \subseteq S$, and a transition relation $\rightarrow \subseteq S \times S$.
 We implicitly assume a set of state variables such that a state $\sigma \in S$ is a total function that maps state variables to values.
+We distinguish two special state variables in a transition system: (1) $now \in \mathcal{R}$ denotes the current time in the state, and (2) $c$ is an event calendar.
 
-We distinguish two special state variables: (1) $now \in \mathcal{R}$ denotes the current time in the state, and (2) $c$ is an event calendar.
-
-The following laws must hold of a calendar automaton:
+The following laws must hold of a transition system $\mathcal{M}$ implementing a calendar automaton:
 
 \begin{itemize}
-\item Time is initialized to 0 in the initial state: $\forall \sigma \in I$, $\sigma(now) = 0$.
-\item In all states, every calendar event occurs no sooner than the current time: $\forall \sigma \in S, \, <e_i, t_i> \in C$, $\sigma(t_i) /geq \sigma(now)$.
-\item In all states, if the current time is strictly less than every calendar event, then the only enabled transition is a \emph{time progress} update: $\forall \sigma \in S \, <e_i, t_i> \in C$, if $\sigma(now) < t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma' = \sigma \backslash now = min$
+\item Time is initialized to be less than or equal to every calendar timeout: $\forall \sigma \in I, \, \forall (e_i, t_i) \in \sigma(c)$, $\sigma(now) \leq t_i$.
+\item In all states, every calendar event occurs no sooner than the current time: $\forall \sigma \in S \, \forall (e_i, t_i) \in \sigma(c)$, $t_i \geq \sigma(now)$.
+\item In all states, if the current time is strictly less than every calendar event, then the only enabled transition is a \emph{time progress} update: $\forall \sigma \in S \, \forall (e_i, t_i) \in \sigma(c)$, if $\sigma(now) < t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma' = \sigma[now := min(c)]$.
 \end{itemize}

### 2016-06-09 doc/om-paper/paper.tex
Add paper.bib and one citation

@@
 \newtheorem{definition}{Definition}
+\newtheorem{theorem}{Theorem}
+\newtheorem{lemma}{Lemma}
 
@@
 
-\paragraph{Contributions}
-
-Our work combines a number of ideas in the formal verification literature to build scalable, generic formal models of fault-tolerant distributed systems.
+Our contributions combine a number of ideas in the formal verification literature to build scalable, generic formal models of fault-tolerant distributed systems.
 
@@
 The model combines several formalisms from the literature to efficiently but systematically model and verify a broad class of systems.
-The model extends \emph{calendar automata}, originally developed by Dutetre and Sorea~\cite{}.
+The model extends \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{Dutertre-Sorea}.
 A calendar automata is a model inspired by discrete-event simulation, used in testing real-time systems~\cite{}.
@@
 
-An \emph{event calendar} $\{ (e_0, t_0), (e_1, t_1), \ldots, (e_n, t_n) \}$ is a finite set of ordered pairs $(e_i, t_i)$ called \emph{calendar events} where $e_i \in E$ is an event and $t_i \in \mathcal{R}$ is a \emph{timeout}, the time at which the event is scheduled.
+An \emph{event calendar} $\{ (e_0, t_0), (e_1, t_1), \ldots, (e_n, t_n) \}$ is a set of ordered pairs $(e_i, t_i)$ called \emph{calendar events} where $e_i \in E$ is an event and $t_i \in \mathbb{R}$ is a \emph{timeout}, the time at which the event is scheduled.
 We denote element $(e_1, t_1)$ of an event calendar by $c_i$.
@@
 We implicitly assume a set of state variables such that a state $\sigma \in S$ is a total function that maps state variables to values.
-We distinguish two special state variables in a transition system: (1) $now \in \mathcal{R}$ denotes the current time in the state, and (2) $c$ is an event calendar.
+We distinguish two special state variables in a transition system: (1) $now \in \mathbb{R}$ denotes the current time in the state, and (2) $c$ is an event calendar.
 
@@
 
-\begin{itemize}
+\begin{enumerate}
 \item Time is initialized to be less than or equal to every calendar timeout: $\forall \sigma \in I, \, \forall (e_i, t_i) \in \sigma(c)$, $\sigma(now) \leq t_i$.
 \item In all states, every calendar event occurs no sooner than the current time: $\forall \sigma \in S \, \forall (e_i, t_i) \in \sigma(c)$, $t_i \geq \sigma(now)$.
-\item In all states, if the current time is strictly less than every calendar event, then the only enabled transition is a \emph{time progress} update: $\forall \sigma \in S \, \forall (e_i, t_i) \in \sigma(c)$, if $\sigma(now) < t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma' = \sigma[now := min(c)]$.
-\end{itemize}
+\item In all states, if the current time is strictly less than every calendar event, then the only enabled transition is a \emph{time progress} update: $\forall \sigma \in S, \, \forall (e_i, t_i) \in \sigma(c)$, if $\sigma(now) < t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma' = \sigma[now := min(c)]$.
+\item In all states, if the current time equals a timeout, then the only transition enabled is a calendar event update associated with the timeout: $\forall \sigma \in S, \, \exists (e_i, t_i) \in \sigma(c)$ such that $\sigma(now) = t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma'(now) = \sigma(now)$, $\sigma'(c_j) = \sigma(c_j)$ for all $c_j \in \sigma(c)$ such that $c_j \neq c_i$, and $\sigma'(t_i) > \sigma(t_i)$.
+\end{enumerate}
 
-\emph{give theorems for calendar, describe calendar for messages---all you need}
+\begin{lemma}[Monotonic time]
+$\forall \sigma, \sigma' \in S$, if $\sigma \rightarrow \sigma'$, then $\sigma'(now) \geq \sigma(now)$.
+\end{lemma}
+
+In a distributed system, it is convenient to distinguish global actions and local actions.
+Global actions are principally interprocess communication, while local actions are those carried out by each process to update its local state and produce new messages to broadcast.
+While both global and local actions can both be modeled as events in a calendar automata, doing so is generally overkill and complicates the model.
+Real-time constraints matter at the system-level but rarely within an individual process.
+From the global perspective, individual processes can update their local state atomically.
+
+Following Dutetre and Sorea, we associate calendar events with channels in a distributed system~\cite{dutetre}.
+Specializing calendars to message passing does not lose generality since all external communication from an individual process can be abstracted as message passing.
+Furthermore, typical fault models can be abstracted to act over channels rather than processes~\cite{abstractions}.
+The calendar introduces real-time constraints on when processes send and receive messages.
+
+Assume processes are indexed with the natural numbers $\mathbb{N}$.
+A \emph{channel} from process $i$ to $j$ is an ordered pair $(i,j)$.
+Define a set of messages $Msg$, with a designated element $null \in Msg$ denoting the absence of a message.
+Given a channel and a timeout, let $send$ return the message sent on the channel at that time:
+$$send : \mathbb{N} \times \mathbb{N} -> \mathbb{R} -> Msg$$
+So $send(i, j, t)$ returns the message that $i$ sends to $j$ at time $t$. Likewise, let
+$$recv : \mathbb{N} \times \mathbb{N} -> \mathbb{R} -> Msg$$
+compute the message received on that channel at that time, so that $recv(i, j, t)$ returns the message received by $j$ from $i$ at time $t$.
+
+We require that messages received were previously sent and not previously received: if $recv(i, j, t) \neq null$, then $\exists t'$ such that $send(i, j, t')$ where $t' < t$, $recv(i, j, t) = send(i, j, t')$, and $\neg\exists t''$ such that $t' < t'' < t$ and $recv(i, j, t'') = recv(i, j, t)$
+
+Then an event calendar for sending and receiving messages on channels is a set of send or receive events together with their timeouts that yield non-null messages; that is,
+$$
+\{ (e, t) | \forall i,j: e \neq null \textnormal{ and } (e = send(i, j, t) \textnormal{ or } e = recv(i, j, t))
+\}
+$$
+
+\lee{technically, atomic time is just a convenience.}
+The event of receiving a message initiates a process to update its local state machine and generate additional messages to send.
+When the process is updating its local state machine, the event calendar is paused.
+That is, updating an event $recv(i, j, t)$ implicitly includes updating $j$'s state machine.
+
+\lee{note that we don't need full generality of send and receive events in OM(1)}
 
 \subsection{Synchronous Observers}\label{sec:sync}
+Synchronous observers are state machines that are synchronously composed with a system specification to check properties over the system's state variables.
+Synchronous observers are a method for composing specifications with a program.
+Synchronous observers were first introduced in the context of the synchronous programming languages, such as Lustre~\cite{}.
+While synchronous observers have been part of model-checking folklore for some time, Rushby systematically describes their application to the domain~\cite{}.
+The general form for specifying synchronous observers is the following:
+$$system | assumptions | requirements$$
+\noindent
+where `$|$' denotes synchronous composition of state machines.
 
@@
 
-\subsection{Abstract State Machines}\label{sec:abstract}
+% ------------------------------------------------------------
+\section{Abstract State Machines}\label{sec:abstract}
+
+% ------------------------------------------------------------
+\section{An Extended Example: Byzantine Agreement}\label{sec:byz}
 
@@
 
-\subsection{Byzantine Agreement}\label{sec:byz}
-
 % ------------------------------------------------------------
@@
 
+\bibliographystyle{IEEEtran}
+\bibliography{paper}
+
 \end{document}

### 2016-06-14 doc/om-paper/paper.tex
more paper text

@@
 
-\lee{finish intro...}
+\lee{finish intro...  One more contribution is that we abstract faults using partially interpreted functions in model-checking, which is new? Note focus is on infinite-state model-checking}
 
@@
 \noindent
-where `$|$' denotes synchronous composition of state machines.
+where `$|$' denotes synchronous composition of state machines. The assumptions constrain the system's state variables, while the requirements monitor the system's state variables.
 
-\paragraph{Fault model}
+The benefit of the synchronous observer approach is that we can develop a system specification that contains no extraneous modeling context from which to synthesize an implementation.
+For example, we do not want synthesize an implementation that injects faults!
+Furthermore, modifying assumptions and requirements is simpler when they are not distributed throughout the model.
 
-\paragraph{Property Observer}
+%% As a simple example, consider the following transition system.
+In the following, \texttt{'} denotes the next-state value of a variable, and texttt{-->} is a two-place operator, preceded by a guard, such that if the guard is true, the state update following the arrow is applied.
+One transition with a true guard is nondeterministically applied.
+
+%% \begin{verbatim}
+%% system:
+%%   initialization: i = 0, b = True
+%%   transitions:
+%%     b == True  --> i' = i+2, b' = False
+%%     b == False --> i' = i-1, b' = True
+%% \end{verbatim}
+
+%% The linear temporal logic (LTL) property \texttt{G(i >= 0)} holds for this system. Let us define a requirements transition system to check the same property:
+
+%% \begin{verbatim}
+%% requirements:
+%%   initialization: nonneg = True
+%%   transitions:
+%%     i >= 0 -->
+%% \end{verbatim}
+
+\subsection{Property Observer}
+
+% ------------------------------------------------------------
+\section{The Synchronous Kibitzer}
+The typical approach to injecting faults into a formal model of a distributed system is to add a new state variables to each process representing its fault state.
+Then additional state variables are added to represent the nondeterminism introduced by faults into the state or messages by a process.
+Finally, additional transitions are added to each process to modify the process's state or messages based on it's current fault state.
+
+This straightforward approach introduces substantial additional state and nondeterminism, reducing the scalability of model-checking.
+Modeling faults is particularly expensive given that additional state and nondeterminism is added to each process.
+These scalability issues are noted throughout the literature.
+For example, \lee{TTA startup paper, helmut's paper, etc.}
+
+Not only is modeling faults expensive, but it is generally non-compositional.
+For example, to model-check the Oral Messages algorithm~\cite{} that provides interactive consistency in the presence of Byzantine faults, Rushby every transition of the General and Lieutenant processes acts on their fault state~\cite{}.
+
+Another difficulty with adding faults to a model is that the number of states that must be introduced may be dependent non-obviously on other aspects of the fault model, specific protocol, and system size.
+Furthermore, since SAT-based model-checking is NP-complete, it is important not to introduce any state beyond the minimal required.
+Such constraints lead to ``meta-model'' reasoning, such as the following, as described by Rushby:
+
+\begin{quote}
+To achieve the full range of faulty behaviors, it seems that a faulty source should be able to send a different incorrect value to each relay, and this requires n different values.
+It might seem that we need some additional incorrect values 3 so that faulty relays can exhibit their full range of behaviors.
+It would certainly be safe to introduce additional values for this purpose, but the performance of model checking is very sensitive to the size of the state space, so there is a countervailing argument against introducing additional values.
+A little thought will show that the way faulty relays have their most significant impact is by tipping the majority vote in a receiver one way or the other, and this can only be achieved if they use the same values as nonfaulty relays.
+Hence, we decide against further extension to the range of values~\cite{}.
+\end{quote}
+
+To address these problems, we introduce a synchronous observer that injects faults that we call a \emph{synchronous kibitzer}.
+The kibitzer decomposes the state and transitions associated with the fault model from the system.
+
+For the sake of concreteness, we introduce
+ under the hybrid fault model of Thambidurai and Park~\cite{}.
+The fault model distinguishes Byzantine (or arbitrary), symmetric, and manifest faults.
+This fault model applies to broadcast systems, in which a process is expected to broad the same value to multiple receivers.
+A Byzantine faulty
 
@@
 
+\lee{let's also check out scalability when we ``turn off'' faults.
+Also, how hard is it to turn them off or change the fault model?
+E.g., how many lines of spec need to be changed?
+Compare to rushby's?}
+
 % ------------------------------------------------------------

### 2016-07-05 doc/om-paper/paper.tex

@@
 Verification abstractions are often ad-hoc and system-specific.
-For example, to model a node that can broadcast the values $\{0, 1\}$, or no value if when faulty, one might define an enumerated type of broadcast values sent such as
+For example, to model a process that can broadcast the values $\{0, 1\}$, or no value if when faulty, one might define an enumerated type of broadcast values sent such as
 
@@
 Conflating them makes formal model-to-code correspondence more difficult to achieve.
-An example of a verification optimization is to model message passing as shared state between nodes.
+An example of a verification optimization is to model message passing as shared state between processes.
 While these optimizations might be safe under particular assumptions, they can lead to unsound models when a system or its environment is modified.
@@
 We demonstrate the model with an extended case-study, verifying the Oral Messages (1) protocol (OM(1))~\cite{}.
-While our model is more detailed---it explicitly models channels---and also more general---it allows bounded asynchrony between nodes---it is the most scalable model-checking verification of the protocol.
+While our model is more detailed---it explicitly models channels---and also more general---it allows bounded asynchrony between processes---it is the most scalable model-checking verification of the protocol.
 
@@
 
-\subsection{Calendar Automata}\label{sec:calendar}
+\subsection{Calendar Automata}\label{sec:calendar}\label{sec:calendar}
 \emph{Timed automata} are specialized formalisms for verifying real-time systems~\cite{}.
@@
 
-Let $c$ be an event calendar and $c_i, c_j \in c$ be calendar events.
-Define an ordering on event calendars such that $c_i \leq c_j$ iff $t_i \leq t_j$, and $min(c) = \{ c_i | \forall c_j \in c, \, c_i \leq c_j  \}$ are the minimum elements of $c$.
+Let $cal$ be an event calendar and $c_i, c_j \in c$ be calendar events.
+Define an ordering on event calendars such that $c_i \leq c_j$ iff $t_i \leq t_j$, and $min(c) = \{ c_i | \forall c_j \in cal, \, c_i \leq c_j  \}$ are the minimum elements of $cal$.
 
@@
 We implicitly assume a set of state variables such that a state $\sigma \in S$ is a total function that maps state variables to values.
-We distinguish two special state variables in a transition system: (1) $now \in \mathbb{R}$ denotes the current time in the state, and (2) $c$ is an event calendar.
+We distinguish two special state variables in a transition system: (1) $now \in \mathbb{R}$ denotes the current time in the state, and (2) $cal$ is an event calendar.
 
@@
 \begin{enumerate}
-\item Time is initialized to be less than or equal to every calendar timeout: $\forall \sigma \in I, \, \forall (e_i, t_i) \in \sigma(c)$, $\sigma(now) \leq t_i$.
-\item In all states, every calendar event occurs no sooner than the current time: $\forall \sigma \in S \, \forall (e_i, t_i) \in \sigma(c)$, $t_i \geq \sigma(now)$.
-\item In all states, if the current time is strictly less than every calendar event, then the only enabled transition is a \emph{time progress} update: $\forall \sigma \in S, \, \forall (e_i, t_i) \in \sigma(c)$, if $\sigma(now) < t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma' = \sigma[now := min(c)]$.
-\item In all states, if the current time equals a timeout, then the only transition enabled is a calendar event update associated with the timeout: $\forall \sigma \in S, \, \exists (e_i, t_i) \in \sigma(c)$ such that $\sigma(now) = t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma'(now) = \sigma(now)$, $\sigma'(c_j) = \sigma(c_j)$ for all $c_j \in \sigma(c)$ such that $c_j \neq c_i$, and $\sigma'(t_i) > \sigma(t_i)$.
+\item Time is initialized to be less than or equal to every calendar timeout: $\forall \sigma \in I, \, \forall (e_i, t_i) \in \sigma(cal)$, $\sigma(now) \leq t_i$.
+
+\item In all states, every calendar event occurs no sooner than the current time: $\forall \sigma \in S \, \forall (e_i, t_i) \in \sigma(cal)$, $t_i \geq \sigma(now)$.
+
+\item In all states, if the current time is strictly less than every calendar event, then the only enabled transition is a \emph{time progress} update: $\forall \sigma \in S, \, \forall (e_i, t_i) \in \sigma(cal)$, if $\sigma(now) < t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma' = \sigma[now := min(c)]$.
+
+\item In all states, if the current time equals a timeout, then the only transition enabled is a calendar event update associated with the timeout: $\forall \sigma \in S, \, \exists (e_i, t_i) \in \sigma(cal)$ such that $\sigma(now) = t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma'(now) = \sigma(now)$, $\sigma'(c_j) = \sigma(c_j)$ for all $c_j \in \sigma(c)$ such that $c_j \neq c_i$, and $\sigma'(t_i) > \sigma(t_i)$.
 \end{enumerate}
@@
 
-Assume processes are indexed with the natural numbers $\mathbb{N}$.
+Assume processes are indexed from a finite set $Id \subset \mathbb{N}$.
 A \emph{channel} from process $i$ to $j$ is an ordered pair $(i,j)$.
@@
 Given a channel and a timeout, let $send$ return the message sent on the channel at that time:
-$$send : \mathbb{N} \times \mathbb{N} -> \mathbb{R} -> Msg$$
+$$send : Id \times Id \rightarrow \mathbb{R} \rightarrow Msg$$
 So $send(i, j, t)$ returns the message that $i$ sends to $j$ at time $t$. Likewise, let
-$$recv : \mathbb{N} \times \mathbb{N} -> \mathbb{R} -> Msg$$
+$$recv : Id \times Id \rightarrow \mathbb{R} \rightarrow Msg$$
 compute the message received on that channel at that time, so that $recv(i, j, t)$ returns the message received by $j$ from $i$ at time $t$.
 
-We require that messages received were previously sent and not previously received: if $recv(i, j, t) \neq null$, then $\exists t'$ such that $send(i, j, t')$ where $t' < t$, $recv(i, j, t) = send(i, j, t')$, and $\neg\exists t''$ such that $t' < t'' < t$ and $recv(i, j, t'') = recv(i, j, t)$
+In the non-faulty case, we require that messages received were previously sent and not previously received: if $recv(i, j, t) \neq null$, then $\exists t'$ such that $send(i, j, t')$ where $t' < t$, $recv(i, j, t) = send(i, j, t')$, and $\neg\exists t''$ such that $t' < t'' < t$ and $recv(i, j, t'') = recv(i, j, t)$.
+(We address faults in Section~\ref{sec:kibitzer}.)
 
@@
 % ------------------------------------------------------------
-\section{The Synchronous Kibitzer}
+\section{The Synchronous Kibitzer}\label{sec:kibitzer}
 The typical approach to injecting faults into a formal model of a distributed system is to add a new state variables to each process representing its fault state.
@@
 The kibitzer decomposes the state and transitions associated with the fault model from the system.
-
-For the sake of concreteness, we introduce
- under the hybrid fault model of Thambidurai and Park~\cite{}.
+For the sake of concreteness, we introduce a particular fault model, the hybrid fault model of Thambidurai and Park~\cite{}.
 The fault model distinguishes Byzantine (or arbitrary), symmetric, and manifest faults.
 This fault model applies to broadcast systems, in which a process is expected to broad the same value to multiple receivers.
-A Byzantine faulty
+A \emph{Byzantine fault} is one in which a process that is intended to broadcast the same value to other processes may instead broadcast arbitrary values to different receivers (including no value or the correct value).
+A \emph{symmetric fault} is one in which a process may broadcast the same, but incorrect, value to other processes, and a \emph{manifest fault} is one in which a process's broadcast fault is detectable by the receivers; e.g., by performing a cyclic redundancy check (CRC) or because the value arrives outside of a predetermined window.
+
+In specifying a fault-tolerant system, we must also specify a \emph{maximum fault assumption} (MFA), which specifies the maximum number of of each kind of fault that may be present in the system for some system property to hold.
+Define a enumeration of fault types $$\textnormal{FAULT\_TYPE} = \{none, byz, sym, man\}$$.
+As in the previous section, let $Id \subset \mathbb{N}$ be a finite set of indices, and let the variable $$faults: Id \rightarrow \textnormal{FAULT\_TYPE}$$ range over possible mappings from process indices to faults.
+
+The hybrid fault model assumes a broadcast model of communication.
+% Define $rnd: \mathbb{R} \rightarrow \mathbb{N}$ such that if $rnd(t_0) < rnd(t_1)$, then $t_0 < t_1$.
+A $broadcast: Id \rightarrow 2^{Id} \rightarrow \mathbb{R} \rightarrow Msg \rightarrow 2^{E}$ takes a sender, a set of receivers, a real-time, and a message to send each receiver, and returns a set of calendar events:
+$$broadcast(i, R, t, m) = \{m | j \in R \textnormal{ and } send(i, j, t) = m\}$$
+
+\lee{msg in SAL is recv here}
+
+With this machinery, we can define the semantics of faults by constraining the relationship between a message broadcast and the values received by the recipients.
+For a nonfaulty process that broadcasts, every recipient receives the sent message.
+\begin{align*}
+&nonfaulty\_constraint =\\
+  &\quad \forall i, j \in Id, t \in \mathbb{R}.\\
+  &\quad\quad (faults(i) = none\\
+  &\quad\quad\quad \textnormal{ implies } recv(i, j, t) = send(i, j, t)
+\end{align*}
+\noindent
+For symmetric faults, we there is no requirement that the messages sent are the ones received, only that every recipient receives the same value.
+\begin{align*}
+&sym\_constraint =\\
+  &\quad \forall i, j, k \in Id, t \in \mathbb{R}.\\
+  &\quad\quad (faults(i) = sym \textnormal{ and } broadcast(i, \{j, k\}, t, m))\\
+  &\quad\quad\quad \textnormal{ implies } recv(i, j, t) = recv(i, k, t)
+\end{align*}
+\noindent
+Byzantine faults are left completely unconstrained.
 
 % ------------------------------------------------------------
-\section{Abstract State Machines}\label{sec:abstract}
+\section{Abstract State Machines}\label{sec:abstract}\label{sec:asms}
+When model-checking a system, counterexamples to properties are traces of states.
+A state is a collection of variable values.
+For industrial-scale models, there are many 10s or even 100s of state variables.
+So a single state includes assignments to each of the variables, and counterexample trace may be many 10s of states in length.
+In practice, counterexamples depend on a small number of state variables.
+The user has to manually separate the state variables that matter from those that do not for a specific property.
+It is not enough prevent variables that stutter from displaying since some state variables might be updated without affecting the property.
+The problem is particularly pertinent when model-checking models that have sufficient detail for implementation synthesis, as is the goal in our work.
+
+Systems are usually designed in modes such predicates on state variables are parameterized on which mode the system is in.
+The transition between modes implements a finite state-machine.
+The modes and the state-machine are usually implicit, but a property violation usually requires an unexpected mode transition.
+Abstracting counterexample traces over all of the state variables by traces over the modes helps to identify why a property fails quickly.
+
+Our solution to the problem of constructing a mode-based state machine uses abstract state machines (ASMs).
+ASMs are a generalization of finite state machines originally developed by Gurevich to generalize the Church-Turing thesis for abstract data structures~\cite{}.
+Since then, ASMs have become an engineering technique for simulating a system at different levels of abstraction.
+
+\lee{http://www2.informatik.hu-berlin.de/sam/preprint/reisig203.pdf}
+More precisely, we implement a sequential small-step ASM~\cite{}.
+\begin{definition}[Sequential small-step ASM]
+Let $V$ be a finite set of state variables. Let $s$ be an assignment of variables to values, and let $b$ be a relation Å
+\end{definition}
 

### 2016-07-25 doc/om-paper/paper.tex
om1 paper updates

@@
 
-Fault-tolerant distributed systems are essential to life-critical systems, like avionics.
+Fault-tolerant distributed systems are essential to reliable life-critical systems, like avionics.
 Consequently, this class of systems demand high-assurance of correct design and implementation.
@@
 There is substantial literature on verifying fault-tolerant distributed systems~\cite{}.
-Some of this work uses mechanical theorem-proving~\cite{}.
-Two problems with mechanical verification are that it requires specialized expertise and it is labor intensive.
-In contrast, model-checking is automated but has limited scalability.
-To address scalability, abstractions and optimizations are employed, both of which lead to problems.
+Most of this work uses interactive theorem-proving~\cite{}.
+Interactive theorem-proving is useful in the domain because of the complexity of the models; automated methods are difficult to scale.
+However, two problems with interactive verification are that it requires specialized expertise and it is labor intensive.
+
+In contrast, model-checking is automated but has limited scalability. To address scalability, abstractions and optimizations are employed, both of which lead to problems.
+
 Verification abstractions are often ad-hoc and system-specific.
-For example, to model a process that can broadcast the values $\{0, 1\}$, or no value if when faulty, one might define an enumerated type of broadcast values sent such as
+For example, to model a process that can broadcast the values $\{0, 1\}$, or no value when faulty, one might define an enumerated type of broadcast values sent such as
 
@@
 
-Our contributions combine a number of ideas in the formal verification literature to build scalable, generic formal models of fault-tolerant distributed systems.
+Our contributions combine a number of ideas in the formal verification literature to build scalable, generic formal models of fault-tolerant distributed systems in which verification is more automated than in traditional interactive theorem-proving.
 
@@
 
-We demonstrate the model with an extended case-study, verifying the Oral Messages (1) protocol (OM(1))~\cite{}.
+We validate the model with a number of small examples \lee{todo}, as well as an extended case-study, verifying the Oral Messages (1) protocol (OM(1))~\cite{}.
 While our model is more detailed---it explicitly models channels---and also more general---it allows bounded asynchrony between processes---it is the most scalable model-checking verification of the protocol.
@@
 \emph{Timed automata} are specialized formalisms for verifying real-time systems~\cite{}.
-Time automata assume the existence of continuously-varying real-valued variables, known as \emph{real-time clocks} in a solver.
-While successful timed automata model-checkers exist, they are specialized for verifying real-time systems in which the only infinite-valued variables are the real-time clocks.
+Timed automata assume the existence of continuously-varying real-valued variables, known as \emph{real-time clocks} in a solver.
+Timed automata model-checkers are specialized for verifying real-time systems in which the only infinite-valued variables are the real-time clocks.
 Modern general-purpose model-checkers like \lee{name some} use SMT solvers for general infinite-state system verification that is not limited to real-time clocks.
@@
 Define a set of \emph{events} $e_0, e_1, \ldots, e_n \in E$.
-We do not define events; intuitively, an event is a set of state variables.
+For now, we do not define events; intuitively, an event is a set of state variables (shortly, we will associate events with messages sent in a distributed system).
 When an event is \emph{enabled}, the transitions over events are enabled; otherwise, the variables stutter and maintain the same value.
@@
 An \emph{event calendar} $\{ (e_0, t_0), (e_1, t_1), \ldots, (e_n, t_n) \}$ is a set of ordered pairs $(e_i, t_i)$ called \emph{calendar events} where $e_i \in E$ is an event and $t_i \in \mathbb{R}$ is a \emph{timeout}, the time at which the event is scheduled.
-We denote element $(e_1, t_1)$ of an event calendar by $c_i$.
+We denote element $(e_i, t_i)$ of an event calendar by $c_i$.
 
@@
 While both global and local actions can both be modeled as events in a calendar automata, doing so is generally overkill and complicates the model.
-Real-time constraints matter at the system-level but rarely within an individual process.
 From the global perspective, individual processes can update their local state atomically.
@@
 A \emph{channel} from process $i$ to $j$ is an ordered pair $(i,j)$.
-Define a set of messages $Msg$, with a designated element $null \in Msg$ denoting the absence of a message.
-Given a channel and a timeout, let $send$ return the message sent on the channel at that time:
-$$send : Id \times Id \rightarrow \mathbb{R} \rightarrow Msg$$
-So $send(i, j, t)$ returns the message that $i$ sends to $j$ at time $t$. Likewise, let
-$$recv : Id \times Id \rightarrow \mathbb{R} \rightarrow Msg$$
-compute the message received on that channel at that time, so that $recv(i, j, t)$ returns the message received by $j$ from $i$ at time $t$.
+Define a set of messages $Msg$.
+Given a channel and a timeout, let $send$ be a relation on messages sent on a channel at a given time:
+$$send \subseteq Id \times Id \times \mathbb{R} \times Msg$$
+So $send(i, j, t, m)$ holds iff $i$ sends to $j$ message $m$ at time $t$. Likewise, let
+$$recv \subseteq Id \times Id \times \mathbb{R} \times Msg$$
+be a relation on messages received on a channel at a time, so that $recv(i, j, t, m)$ holds iff the message $m$ received by $j$ from $i$ at time $t$.
 
-In the non-faulty case, we require that messages received were previously sent and not previously received: if $recv(i, j, t) \neq null$, then $\exists t'$ such that $send(i, j, t')$ where $t' < t$, $recv(i, j, t) = send(i, j, t')$, and $\neg\exists t''$ such that $t' < t'' < t$ and $recv(i, j, t'') = recv(i, j, t)$.
+In the non-faulty case, we require that messages received were previously sent and not previously received: if $recv(i, j, t, m)$, then $\exists t'$ such that $send(i, j, t', m)$ where $t' < t$, and $\neg\exists t''$ such that $t' < t'' < t$ and $recv(i, j, t'', m) = recv(i, j, t, m)$.
 (We address faults in Section~\ref{sec:kibitzer}.)
 
-Then an event calendar for sending and receiving messages on channels is a set of send or receive events together with their timeouts that yield non-null messages; that is,
-$$
-\{ (e, t) | \forall i,j: e \neq null \textnormal{ and } (e = send(i, j, t) \textnormal{ or } e = recv(i, j, t))
-\}
-$$
+Then an event calendar for sending and receiving messages on channels is the union of the $send$ and $recv$ relations.
 
@@
 When the process is updating its local state machine, the event calendar is paused.
-That is, updating an event $recv(i, j, t)$ implicitly includes updating $j$'s state machine.
+That is, updating an event $recv(i, j, t, m)$ also includes updating $j$'s state machine.
 
@@
 
+\lee{todo}
+
 % ------------------------------------------------------------
@@
 Not only is modeling faults expensive, but it is generally non-compositional.
-For example, to model-check the Oral Messages algorithm~\cite{} that provides interactive consistency in the presence of Byzantine faults, Rushby every transition of the General and Lieutenant processes acts on their fault state~\cite{}.
+For example, in a classic model-checking model of the fault-tolerant Oral Messages algorithm~\cite{}, every transition of the sending processes updates environmental state variables associated with the fault state.
 

### 2016-10-12 doc/om-paper/paper.tex
edits to the paper.

@@
-\documentclass{IEEEtran/IEEEtran}
+%\documentclass{IEEEtran/IEEEtran}
+\documentclass{llncs/llncs}
 
@@
 
-\newtheorem{definition}{Definition}
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
+%% IEEE
+%% \newtheorem{definition}{Definition}
+%% \newtheorem{theorem}{Theorem}
+%% \newtheorem{lemma}{Lemma}
 
@@
 
-Fault-tolerant distributed systems are essential to reliable life-critical systems, like avionics.
+Fault-tolerant distributed systems are the backbone of reliable life-critical systems, like avionics.
 Consequently, this class of systems demand high-assurance of correct design and implementation.
@@
 Most of this work uses interactive theorem-proving~\cite{}.
-Interactive theorem-proving is useful in the domain because of the complexity of the models; automated methods are difficult to scale.
+Interactive theorem-proving is useful in the domain because of the complexity of the models and because distributed system verification is typically parameterized by the number of nodes; automated methods perform poorly along these two metrics.
 However, two problems with interactive verification are that it requires specialized expertise and it is labor intensive.
@@
 
-Verification abstractions are often ad-hoc and system-specific.
+For example, verification abstractions are often ad-hoc and system-specific.
 For example, to model a process that can broadcast the values $\{0, 1\}$, or no value when faulty, one might define an enumerated type of broadcast values sent such as
@@
 Our contributions combine a number of ideas in the formal verification literature to build scalable, generic formal models of fault-tolerant distributed systems in which verification is more automated than in traditional interactive theorem-proving.
-
 We describe a generic formal model for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}.
@@
 \begin{enumerate}
-\item Time is initialized to be less than or equal to every calendar timeout: $\forall \sigma \in I, \, \forall (e_i, t_i) \in \sigma(cal)$, $\sigma(now) \leq t_i$.
+\item \label{cal:a} Time is initialized to be less than or equal to every calendar timeout: $\forall \sigma \in I$, $\forall (e_i, t_i) \in \sigma(cal)$, $\sigma(now) \leq t_i$.
 
-\item In all states, every calendar event occurs no sooner than the current time: $\forall \sigma \in S \, \forall (e_i, t_i) \in \sigma(cal)$, $t_i \geq \sigma(now)$.
+%% \item \label{cal:b} In all states, every calendar event occurs no sooner than the current time: $\forall \sigma \in S$, $(e_i, t_i) \in \sigma(cal)$, $t_i \geq \sigma(now)$.
 
-\item In all states, if the current time is strictly less than every calendar event, then the only enabled transition is a \emph{time progress} update: $\forall \sigma \in S, \, \forall (e_i, t_i) \in \sigma(cal)$, if $\sigma(now) < t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma' = \sigma[now := min(c)]$.
+\item \label{cal:c} In all states, if the current time is strictly less than every calendar event, then the only enabled transition is a \emph{time progress} update: $\forall \sigma \in S$, $\forall (e_i, t_i) \in \sigma(cal)$, if $\sigma(now) < t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma' = \sigma[now := min(c)]$.
 
-\item In all states, if the current time equals a timeout, then the only transition enabled is a calendar event update associated with the timeout: $\forall \sigma \in S, \, \exists (e_i, t_i) \in \sigma(cal)$ such that $\sigma(now) = t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma'(now) = \sigma(now)$, $\sigma'(c_j) = \sigma(c_j)$ for all $c_j \in \sigma(c)$ such that $c_j \neq c_i$, and $\sigma'(t_i) > \sigma(t_i)$.
+\item \label{cal:d} In all states, if the current time equals a timeout, then the only transition enabled is a calendar event update associated with the timeout: $\forall \sigma \in S$, $\exists (e_i, t_i) \in \sigma(cal)$ such that $\sigma(now) = t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma'(now) = \sigma(now)$, $\sigma'(c_j) = \sigma(c_j)$ for all $c_j \in \sigma(c)$ such that $c_j \neq c_i$, and $\sigma'(t_i) > \sigma(t_i)$.
 \end{enumerate}
 
+We prove some intermediate results from the laws. First, in every state, the timeouts are never in the past:
+\begin{lemma}[Future timeouts]\label{lem:ft}
+$\forall \sigma \in S$, $(e_i, t_i) \in \sigma(cal)$, $\sigma(now) \leq t_i$.
+\end{lemma}
+\begin{proof}
+By induction on states.
+In an initial state, the property holds immediately (Law~\ref{cal:a}).
+In the induction step, suppose $\sigma(now) \leq t_i$, for all $t_i$.
+If $\sigma(now) < t_i$, for all $t_i$, then the only possible transition is a time progress update to the minimum timeout (Law~\ref{cal:c}).
+Otherwise, if $\sigma(now) = t_i$, for some $t_i$, then the only possible transition is a calendar event update, incrementing $t_i$ but not $\sigma(now)$ (Law~\ref{cal:d}).
+\end{proof}
+
+Second, we prove that time is monotonic:
 \begin{lemma}[Monotonic time]
@@
 \end{lemma}
+\begin{proof}
+From Lemma~\ref{lem:ft}, in $\sigma(now) \leq t_i$, for all $(e_i, t_i) \in \sigma(cal)$.
+For any $\sigma'$ such that $\sigma \rightarrow \sigma'$, either there is a time progress update (Law~\ref{cal:c}) or a calendar event update (Law~\ref{cal:d}).
+In the first case, $\sigma(now) < \sigma'(now)$, and in the second case, $\sigma(now) = \sigma'(now)$.
+\end{proof}
 
@@
 \noindent
-where `$|$' denotes synchronous composition of state machines. The assumptions constrain the system's state variables, while the requirements monitor the system's state variables.
+where `$|$' denotes synchronous composition of transition systems.
+The assumptions constrain the system's state variables, while the requirements monitor the system's state variables.
+More formally, define a transition subset relation:
+
+\begin{definition}[Transition system subset]
+For transition systems $M = (S, I, \rightarrow)$ and $M' = (S', I', \rightarrow')$, $M \subseteq M'$ if and only if $S \subseteq S'$, $I \subseteq I'$, and $\rightarrow \subseteq \rightarrow'$.
+\end{definition}
+
+Let the system, assumptions, and requirements each be the transition systems $\mathcal{S}$, $\mathcal{A}$, and $\mathcal{R}$, respectively.
+Then we require $A \subseteq S$.
+If the system is \emph{realizable}~\cite{} under its assumptions, then the set of states, initial states, and transition relation of $A$ are all nonempty.
+A system is \emph{correct} if $\mathcal{S} \subseteq \mathcal{R}$.
 
 The benefit of the synchronous observer approach is that we can develop a system specification that contains no extraneous modeling context from which to synthesize an implementation.
-For example, we do not want synthesize an implementation that injects faults!
+For example, we do not want to synthesize an implementation that injects faults!
 Furthermore, modifying assumptions and requirements is simpler when they are not distributed throughout the model.
 
-%% As a simple example, consider the following transition system.
-In the following, \texttt{'} denotes the next-state value of a variable, and texttt{-->} is a two-place operator, preceded by a guard, such that if the guard is true, the state update following the arrow is applied.
-One transition with a true guard is nondeterministically applied.
+% ------------------------------------------------------------
+\subsection{Abstract State Machines}\label{sec:abstract}\label{sec:asms}
+When model-checking a system, counterexamples to properties are traces of states.
+A state is a collection of variable values, and in large models, there are a lot.
+A single state includes assignments to each of the variables, and counterexample trace may be many 10s of states in length.
+In practice, counterexamples depend on a small number of state variables.
+The user has to manually separate the state variables that matter from those that do not for a specific property.
+It is not enough prevent variables that stutter from displaying since some state variables might be updated without affecting the property.
+The problem is particularly pertinent when model-checking models that have sufficient detail for implementation synthesis, as is the goal in our work.
 
-%% \begin{verbatim}
-%% system:
-%%   initialization: i = 0, b = True
-%%   transitions:
-%%     b == True  --> i' = i+2, b' = False
-%%     b == False --> i' = i-1, b' = True
-%% \end{verbatim}
+Systems are usually designed in modes such that predicates on state variables are parameterized on which mode the system is in.
+The transition between modes implements a finite state-machine.
+The modes and the state-machine are usually implicit, but a property violation usually requires an unexpected mode transition.
+Abstracting counterexample traces over all of the state variables by traces over the modes helps to identify why a property fails quickly.
 
-%% The linear temporal logic (LTL) property \texttt{G(i >= 0)} holds for this system. Let us define a requirements transition system to check the same property:
+Our solution to the problem of constructing a mode-based state machine uses abstract state machines (ASMs).
+ASMs are a generalization of finite state machines originally developed by Gurevich to generalize the Church-Turing thesis for abstract data structures~\cite{}.
+Since then, ASMs have become an engineering technique for simulating a system at different levels of abstraction.
 
-%% \begin{verbatim}
-%% requirements:
-%%   initialization: nonneg = True
-%%   transitions:
-%%     i >= 0 -->
-%% \end{verbatim}
-
-\subsection{Property Observer}
-
-\lee{todo}
+\lee{http://www2.informatik.hu-berlin.de/sam/preprint/reisig203.pdf}
+More precisely, we implement a sequential small-step ASM~\cite{}.
+\begin{definition}[Sequential small-step ASM]
+Let $V$ be a finite set of state variables. Let $s$ be an assignment of variables to values, and let $b$ be a relation Å
+\end{definition}
 
@@
   &\quad \forall i, j \in Id, t \in \mathbb{R}.\\
-  &\quad\quad (faults(i) = none\\
+  &\quad\quad faults(i) = none\\
   &\quad\quad\quad \textnormal{ implies } recv(i, j, t) = send(i, j, t)
@@
 Byzantine faults are left completely unconstrained.
-
-% ------------------------------------------------------------
-\section{Abstract State Machines}\label{sec:abstract}\label{sec:asms}
-When model-checking a system, counterexamples to properties are traces of states.
-A state is a collection of variable values.
-For industrial-scale models, there are many 10s or even 100s of state variables.
-So a single state includes assignments to each of the variables, and counterexample trace may be many 10s of states in length.
-In practice, counterexamples depend on a small number of state variables.
-The user has to manually separate the state variables that matter from those that do not for a specific property.
-It is not enough prevent variables that stutter from displaying since some state variables might be updated without affecting the property.
-The problem is particularly pertinent when model-checking models that have sufficient detail for implementation synthesis, as is the goal in our work.
-
-Systems are usually designed in modes such predicates on state variables are parameterized on which mode the system is in.
-The transition between modes implements a finite state-machine.
-The modes and the state-machine are usually implicit, but a property violation usually requires an unexpected mode transition.
-Abstracting counterexample traces over all of the state variables by traces over the modes helps to identify why a property fails quickly.
-
-Our solution to the problem of constructing a mode-based state machine uses abstract state machines (ASMs).
-ASMs are a generalization of finite state machines originally developed by Gurevich to generalize the Church-Turing thesis for abstract data structures~\cite{}.
-Since then, ASMs have become an engineering technique for simulating a system at different levels of abstraction.
-
-\lee{http://www2.informatik.hu-berlin.de/sam/preprint/reisig203.pdf}
-More precisely, we implement a sequential small-step ASM~\cite{}.
-\begin{definition}[Sequential small-step ASM]
-Let $V$ be a finite set of state variables. Let $s$ be an assignment of variables to values, and let $b$ be a relation Å
-\end{definition}
 

### 2016-10-19 doc/om-paper/paper.tex
More paper text

@@
 
-\title{Towards Systematic Specification and Verification of Fault-Tolerant Protocols}
+\title{Systematic Specification for Scalable Verification of Fault-Tolerant Algorithms}
 \author{Ben and Lee}
@@
 Consequently, this class of systems demand high-assurance of correct design and implementation.
-Formal verification can be used to prove correctness.
+Formal verification can be used to prove the correctness of their designs.
 
@@
 
-However, the enumeration conflates environment artifacts (i.e., faults) and system artifacts.
+However, the enumeration conflates environmental artifacts (i.e., faults) and system artifacts (i.e., the non-faulty broadcast values).
 Conflating them makes formal model-to-code correspondence more difficult to achieve.
-An example of a verification optimization is to model message passing as shared state between processes.
+
+An example of a verification optimization is to model message passing as shared state between processes to reduce the number of state variables in the model, since the message passing mechanism is elided.
 While these optimizations might be safe under particular assumptions, they can lead to unsound models when a system or its environment is modified.
@@
 Synchronous observers are a modeling abstraction in which an ``observer'' actor is synchronously composed with the system under verification.
-We use synchronous observers to impose fault model constraints on the system.
-Doing so makes the fault-model compositional with the system and easy to modify.
-We also use synchronous observers to observe safety properties, eliding the need for complex linear temporal logic specifications.
+We use synchronous observers to compose distinct transition systems modeling the system, environment, assumptions, and requirements.
 Finally, we use an \emph{abstract state machine}~\cite{}, also composed with the system under verification using synchronous observation.
@@
 While synchronous observers have been part of model-checking folklore for some time, Rushby systematically describes their application to the domain~\cite{}.
-The general form for specifying synchronous observers is the following:
-$$system | assumptions | requirements$$
+The general form for specifying a system using synchronous observers is the following:
+$$\mathcal{S} | \mathcal{E} | \mathcal{A} | \mathcal{R}$$
 \noindent
 where `$|$' denotes synchronous composition of transition systems.
-The assumptions constrain the system's state variables, while the requirements monitor the system's state variables.
-More formally, define a transition subset relation:
+$\mathcal{S}$ is the \emph{system}, or machine to be realized in an implementation.
+$\mathcal{E}$ is the \emph{environment} in which the system operates.
+For example, the environment can be used to specify the behavior of faults in the system.
+$\mathcal{A}$ are the \emph{assumptions} on the system, and $\mathcal{R}$ are the \emph{requirements}.
+Synchronous observers can be thought of as a generalization of state predicates,like Hoare triples~\cite{}, to transitions systems.
 
-\begin{definition}[Transition system subset]
-For transition systems $M = (S, I, \rightarrow)$ and $M' = (S', I', \rightarrow')$, $M \subseteq M'$ if and only if $S \subseteq S'$, $I \subseteq I'$, and $\rightarrow \subseteq \rightarrow'$.
+To give a more precise treatment, we introduce some definitions.
+
+\begin{definition}[Consistent states]
+  States $s, s'$ are consistent iff for every assignment $(x = v) \in s$ and $(x' =
+  v') \in s'$, $x \neq x'$ or $v = v'$.
 \end{definition}
 
-Let the system, assumptions, and requirements each be the transition systems $\mathcal{S}$, $\mathcal{A}$, and $\mathcal{R}$, respectively.
-Then we require $A \subseteq S$.
-If the system is \emph{realizable}~\cite{} under its assumptions, then the set of states, initial states, and transition relation of $A$ are all nonempty.
-A system is \emph{correct} if $\mathcal{S} \subseteq \mathcal{R}$.
+\begin{definition}[Reachable]
+For a transition relation $\rightarrow$ over the states of $S$, for states $s_0, s_1, \ldots, s_k \in S$, $s_k$ is reachable from $s_0$ iff
+$$s_0 \rightarrow s_1 \rightarrow \ldots \rightarrow s_k$$
+\end{definition}
+\noindent
+In particular, a state $s$ is reachable from $s$.
 
-The benefit of the synchronous observer approach is that we can develop a system specification that contains no extraneous modeling context from which to synthesize an implementation.
-For example, we do not want to synthesize an implementation that injects faults!
-Furthermore, modifying assumptions and requirements is simpler when they are not distributed throughout the model.
+In the following definitions, let
+$M = (S, I, \rightarrow)$
+and
+$M' = (S', I', \rightarrow')$
+be transitions systems.
+
+\begin{definition}[Deadlock]
+$M$ is deadlocked iff either the set of initial states $I$ is empty, or there exists a state $s_0 \in S$ and state $s_1 \in I$ such that $s_0$ is reachable from $s_1$ and there exists no $s_2 \in S$ such that $s_0 \rightarrow s_2$.
+\end{definition}
+
+%% \begin{definition}[Transition system union]
+%% %% $M \cup M' = (S \cup S', I \cup I', \rightarrow_{M|M'})$ where $\rightarrow_{M|M'}$ is the smallest relation such that $(s,s') \in \rightarrow_{M|M'}$ iff 
+%% \end{definition}
+
+In defining the synchronous composition of two transition systems, we allow the possibility that composing two states can lead to inconsistent states (as opposed to the semantics in which composition of states $s_0$ and $s_1$ is the ordered pair $(s_0, s_1)$).
+
+\begin{definition}[Synchronous composition]
+The synchronous composition of $M$ and $M'$ is a transition system
+$$M | M' = (S_{M | M'}, I_{M | M'}, \rightarrow_{M | M'})$$
+\noindent
+such that $S_{M | M'} = S \cup S'$,
+$$I_{M | M'} = \{ s_I \cup s_{I'} | s_I \in I\textnormal{, }s_{I'} \in I'\textnormal{, and } s_I \textnormal{ and } s_{I'}\textnormal{ are consistent}\}$$
+\noindent
+and for $s, s' \in S_{M | M'}$, $s \rightarrow_{M | M'} s'$ iff
+\begin{itemize}
+\item there exist states $s_M, s'_M \in S$ such that $s_M \rightarrow s'_M$, $s_M \subseteq s$, and $s'_M \subseteq s'$,
+\item there exist states $s_{M'}, s'_{M'} \in S'$ such that $s_{M'} \rightarrow s'_{M'}$, $s_{M'} \subseteq s$, and $s'_{M'} \subseteq s'$,
+\item $s_M$ and $s_{M'}$ are consistent,
+\item and $s'_M$ and $s'_{M'}$ are consistent.
+\end{itemize}
+\end{definition}
+
+\begin{definition}[Transition system refinement]
+$M \subseteq M'$ if and only if $S \subseteq S'$, $I \subseteq I'$, and $\rightarrow \subseteq \rightarrow'$.
+\end{definition}
+
+By convention, the assumptions $\mathcal{A}$ should be a refinement of the system $\mathcal{S}$ and environment $\mathcal{E}$.
+The system is \emph{realizable}~\cite{} under its assumptions if $\mathcal{S} | \mathcal{E} | \mathcal{A}$ is not deadlocked.
+A system is \emph{correct} if $\mathcal{S} | \mathcal{E} | \mathcal{A}$ refines the requirements $\mathcal{R}$.
+
+There are two practical benefits of the synchronous observer approach.
+First, the system $\mathcal{S}$ contains no extraneous modeling context about the environment, assumptions, or requirements, making it easier to decompose $\mathcal{S}$ and to synthesize an implementation directly from it.
+Second, as described by Rushby~\cite{}, specifying requirements is simpler.
 
@@
 % ------------------------------------------------------------
-\section{An Extended Example: Byzantine Agreement}\label{sec:byz}
+\section{Extended Example: OM(1)}\label{sec:byz}
 

### 2016-10-19 doc/om-paper/paper.tex
add figs.

@@
 \usepackage{xcolor}
+
+\usepackage{graphicx}
 
@@
 
+We have instantiated the formal model described above to model and verify the classic OM(1) Byzantine fault-tolerant algorithm~\cite{}.
+OM(1) and its variants have been verified in other settings, allowing us to benchmark our results against others (see Section~\ref{sec:experimental} for experimental results).
+
+The verification is carried out in the Symbolic Analysis Laboratory~\cite{}, which contains a suite of model-checkers and a high-level specification language.
+In our work, we use the infinite-state model-checker using $k$-induction~\cite{}.
+
+\subsection{OM(1) Overview}
+
+\subsection{Model Sketch}
+
+\subsection{Invariants}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=8cm]{figs/proof-structure-validity}
+  \caption{Proof structure for the validity property.}
+  \label{fig:compare}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=8cm]{figs/proof-structure-agreement}
+  \caption{Proof structure for the agreement property.}
+  \label{fig:compare}
+\end{figure}
+
 % ------------------------------------------------------------
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=8cm]{figs/benchmark}
+  \caption{Comparison of number of nodes vs. verification time.}
+  \label{fig:compare}
+\end{figure}
+
+% ------------------------------------------------------------
+\section{Toward an Architectural Domain-Specific Language}\label{sec:adsl}
+
 % ------------------------------------------------------------

### 2016-10-20 doc/om-paper/paper.tex
paper edits, tweak SAL file for paper

@@
 \usepackage{amsfonts}
+
+\usepackage{tikz}
+\newcommand{\OM}[1]{\ensuremath{\mathrm{OM}(#1)}}
 
@@
 
-\lee{finish intro...  One more contribution is that we abstract faults using partially interpreted functions in model-checking, which is new? Note focus is on infinite-state model-checking}
+\lee{finish intro...
+One more contribution is that we abstract faults using partially interpreted functions in model-checking, which is new?
+Note focus is on infinite-state model-checking: will need user-provided lemmas.
+compositional approach to providing lemmas }
 
@@
 
-\subsection{Calendar Automata}\label{sec:calendar}\label{sec:calendar}
+\subsection{Calendar Automata}\label{sec:calendar}
 \emph{Timed automata} are specialized formalisms for verifying real-time systems~\cite{}.
@@
 \subsection{OM(1) Overview}
+Our first case is the ``Oral Messages'' algorithm to solve the Byzantine
+Generals Problem~\cite{Lamport-OM}. The algorithm, $\OM{m}$, is parameterized by
+the number of rounds of message relay. Here, we focus on Oral Messages with one
+round of communication, $\OM{1}$, that tolerates a single Byzantine fault, so
+long as there are a sufficient number of nodes nodes.
+
+Figure~\ref{fig:om1} depicts the communication pattern for \OM{1}. Each of the
+$n$ nodes in the system represents a (Byzantine) General. In this formulation, the
+center node is the commanding general, while the $n-1$ outer generals are the
+lieutenants. The algorithm starts with the commanding general sending each
+lieutenant the same message $v$. Upon receipt of the commander's message, each
+lieutenant in turn sends the message received to each of the other lieutenants.
+Once a lieutenant has received all $n-1$ expected messages, a majority vote is
+taken among the $n$ messages and the lieutenant declares its output (course of
+action) to be whichever message is in the majority.
+
+\begin{figure}[ht]
+\centering
+\begin{tikzpicture}[
+ell/.style={x radius=1, y radius=0.5, draw, shape=circle},
+arr/.style={semithick}]
+\node[ell] (c)  at (0,0)   {Commander};
+\node[ell] (l1) at (0:4)   {Lieutenant 1};
+\node[ell] (l2) at (120:4) {Lieutenant 2};
+\node[ell] (l3) [fill=gray!30] at (240:4) {Lieutenant 3};
+%
+\draw[->, arr] (c) -- node[auto] {$v$} (l1);
+\draw[->, arr] (c) -- node[auto] {$v$} (l2);
+\draw[->, arr] (c) -- node[auto] {$v$} (l3);
+\draw[<->, arr] (l1) -- node[auto,swap,near start] {$v$}
+                        node[auto,swap,near end] {$v$} (l2);
+\draw[<->, arr] (l1) -- node[auto,near start,color=red] {$x$}
+                        node[auto,near end] {$v$} (l3);
+\draw[<->, arr] (l2) -- node[auto,swap,near start,color=red] {$x$}
+                        node[auto,swap,near end] {$v$} (l3);
+\end{tikzpicture}
+\caption{\OM{1} Four generals, one traitor}
+\label{fig:om1}
+\end{figure}
+
+For a fault model we follow~\cite{Rushby-OM1} and assume that nodes are the
+only source of faults. Moreover, node faults are assumed to be either
+\emph{manifest}, \emph{symmetric}, or \emph{byzantine}. A fault is called manifest if
+it is detectable by the non-faulty nodes; for example a node that sends to
+another node a message that explicitly indicates there is a fault. On the
+other hand, a fault is called symmetric if its presence is not necessarily
+detectable as a fault to the other components; for example a node which sends
+wrong messages, as opposed to invalid or missing messages.
+
+Given a formal model of \OM{1}, the standard target for verification is
+\emph{validity} and \emph{agreement}. If we let $l_i$ denote the output for
+lieutenant $i$ (where $1 \le i \le n-1$), then validity states:
+%
+\begin{equation}
+    \forall \,i. \quad l_i = v
+\end{equation}
+%
+whereas agreement states:
+%
+\begin{equation}
+    \forall \,i, j. \quad l_i = l_j
+\end{equation}
+%
+where $i,j$ range over the non-faulty nodes.
+
+It is a classical result that \OM{1} can tolerate at most 1 (byzantine) faulty
+node as long as there are at least 4 nodes in total. By adding additional
+communication rounds, \OM{1} can be extended to an algorithm \OM{m} which
+tolerates at most $m$ faults. On the other hand it is known that any system
+which tolerates $m$ byzantine faults must involve at least $3m+1$ nodes.
+
+In the results above there are several assumptions made regarding the
+underlying computational platform which aren't obvious from the informal
+description, but become so when one considers formally modelling it. To make
+as many assumptions as possible explicit, we require:
+\begin{enumerate}
+    \item every message sent by a node is received successfully by the
+        addressed node,
+    \item when a node receives a message it may determine who sent it,
+    \item each node can detect the absence of a message.
+\end{enumerate}
+These are the assumptions made by Lamport in his constructed solution for
+$\OM{m}$ and the corresponding impossibility result (see
+\cite{Lamport-OM}~\S2). It follows that a system implementing $\OM{1}$ must
+transition synchronously. In particular, in an asynchronous system there is no
+general method for detecting when a message is absent.
 
@@
   \centering
-  \includegraphics[width=8cm]{figs/proof-structure-validity}
-  \caption{Proof structure for the validity property.}
-  \label{fig:compare}
+  \includegraphics[width=\textwidth]{figs/proof-structure}
+  \caption{Proof structure.}
+  \label{fig:proof}
 \end{figure}
 
-\begin{figure}
-  \centering
-  \includegraphics[width=8cm]{figs/proof-structure-agreement}
-  \caption{Proof structure for the agreement property.}
-  \label{fig:compare}
-\end{figure}
+The proof structure is shown in Figure~\ref{fig:proof}.
+The two graphs show the structure to prove validity and agreement, respectively.
+The nodes group related lemmas, and a directed edge from node $a$ to $b$ means lemmas of $b$ depend on $a$.
+For readability, we have replicated lemmas shared in the proofs of validity and agreement.
+Additionally, we have elided the dependencies between the lemmas within each node.
+The proofs are all $k=1$ inductive.
+
+The lemmas are classified into five groups: calendar automata lemmas, abstract state machine lemmas, protocol behavior lemmas, majority vote lemmas, and fault model lemmas.
+
+The calendar automata lemmas do not concern the system.
+The \emph{cal\_invt} lemma is a conjunction of basic properties about the calendar automata, essentially characterizing the axiomatization described in Section~\ref{sec:calendar}.
+Additionally, the \emph{missing\_buffer\_size} lemma relates the calendar to a node's message buffer: the messages received added to the empty messages equals the total number of possible messages to receive.
+
+The \emph{abstract\_all} lemmas characterizes the transition between states of the abstract state machine.
+Technically, \emph{abstract\_all} also depends on ``boilerplate'' lemmas relating each state of the abstract state machine to a state variable representing it, which are elided for space.
+
+The system itself includes both the lemmas characterizing the global protocol behavior as well as majority vote\lee{have we talked about the voting algorithm earlier?}.
+The global behavior concerns the entire distributed system, whereas the lemmas regarding the majority vote algorithm only concern the behavior of a single node.
+The proofs mechanize closely the proofs found in Boyer and Moore's original treatment~\cite{}.
+
+Finally, the fault model lemmas relate the fault model to the system.
+The \emph{fault\_propagation} lemma connects the presence of faulty messages in the calendar with the fault status of nodes that receive those messages in subsequent rounds.
+The \emph{symmetric\_relay} and \emph{manifest\_relay} lemmas characterize the effect of symmetric and manifest faults, respectively, on the kinds of messages a receiver obtains from a faulty node.
+All the fault model lemmas mentioned so far are necessary for the agreement proof, but for validity, we need only \emph{candidate\_good}, proving that if the maximum fault assumption holds and the sender is nonfaulty, then no modified messages are received.
 
@@
 % ------------------------------------------------------------
+\section{Related Work}\label{sec:related}
+
+% ------------------------------------------------------------
 \section{Conclusions}\label{sec:conclusions}

### 2016-10-26 doc/om-paper/paper.tex
more paper edits

@@
 \noindent
-such that $S_{M | M'} = S \cup S'$,
-$$I_{M | M'} = \{ s_I \cup s_{I'} | s_I \in I\textnormal{, }s_{I'} \in I'\textnormal{, and } s_I \textnormal{ and } s_{I'}\textnormal{ are consistent}\}$$
-\noindent
-and for $s, s' \in S_{M | M'}$, $s \rightarrow_{M | M'} s'$ iff
+such that
+\begin{itemize}
+\item $S_{M | M'} = S \cup S'$,
+\item $I_{M | M'} = \{ s_I \cup s_{I'} | s_I \in I\textnormal{, }s_{I'} \in I'\textnormal{, and } s_I \textnormal{ and } s_{I'}\textnormal{ are consistent}\}$
+\item and for $s, s' \in S_{M | M'}$, $s \rightarrow_{M | M'} s'$ iff
   \begin{itemize}
@@
   \item and $s'_M$ and $s'_{M'}$ are consistent.
+  \end{itemize}
 \end{itemize}
@@
 
-% ------------------------------------------------------------
-\subsection{Abstract State Machines}\label{sec:abstract}\label{sec:asms}
-When model-checking a system, counterexamples to properties are traces of states.
-A state is a collection of variable values, and in large models, there are a lot.
-A single state includes assignments to each of the variables, and counterexample trace may be many 10s of states in length.
-In practice, counterexamples depend on a small number of state variables.
-The user has to manually separate the state variables that matter from those that do not for a specific property.
-It is not enough prevent variables that stutter from displaying since some state variables might be updated without affecting the property.
-The problem is particularly pertinent when model-checking models that have sufficient detail for implementation synthesis, as is the goal in our work.
-
-Systems are usually designed in modes such that predicates on state variables are parameterized on which mode the system is in.
-The transition between modes implements a finite state-machine.
-The modes and the state-machine are usually implicit, but a property violation usually requires an unexpected mode transition.
-Abstracting counterexample traces over all of the state variables by traces over the modes helps to identify why a property fails quickly.
-
-Our solution to the problem of constructing a mode-based state machine uses abstract state machines (ASMs).
-ASMs are a generalization of finite state machines originally developed by Gurevich to generalize the Church-Turing thesis for abstract data structures~\cite{}.
-Since then, ASMs have become an engineering technique for simulating a system at different levels of abstraction.
-
-\lee{http://www2.informatik.hu-berlin.de/sam/preprint/reisig203.pdf}
-More precisely, we implement a sequential small-step ASM~\cite{}.
-\begin{definition}[Sequential small-step ASM]
-Let $V$ be a finite set of state variables. Let $s$ be an assignment of variables to values, and let $b$ be a relation Å
-\end{definition}
-
-% ------------------------------------------------------------
-\section{The Synchronous Kibitzer}\label{sec:kibitzer}
+\subsection{The Synchronous Kibitzer}\label{sec:kibitzer}
+\lee{We use synchronous composition to implement fault injection. Mention in the intro. This is fundamental to the model, so we have section devoted to it.}
 The typical approach to injecting faults into a formal model of a distributed system is to add a new state variables to each process representing its fault state.
@@
 In specifying a fault-tolerant system, we must also specify a \emph{maximum fault assumption} (MFA), which specifies the maximum number of of each kind of fault that may be present in the system for some system property to hold.
-Define a enumeration of fault types $$\textnormal{FAULT\_TYPE} = \{none, byz, sym, man\}$$.
-As in the previous section, let $Id \subset \mathbb{N}$ be a finite set of indices, and let the variable $$faults: Id \rightarrow \textnormal{FAULT\_TYPE}$$ range over possible mappings from process indices to faults.
+Define a enumeration of fault types $$\textnormal{FAULT\_TYPE} = \{none, byz, sym, man\}$$ As in the previous section, let $Id \subset \mathbb{N}$ be a finite set of indices, and let the variable $$faults: Id \rightarrow \textnormal{FAULT\_TYPE}$$ range over possible mappings from process indices to faults.
 
@@
 Byzantine faults are left completely unconstrained.
+
+% ------------------------------------------------------------
+\subsection{Abstract State Machines}\label{sec:abstract}\label{sec:asms}
+When model-checking a system, counterexamples to properties are traces of states.
+A state is a collection of variable values, and in large models, there are a lot.
+A single state includes assignments to each of the variables, and counterexample trace may be many 10s of states in length.
+In practice, counterexamples depend on a small number of state variables.
+The user has to manually separate the state variables that matter from those that do not for a specific property.
+It is not enough prevent variables that stutter from displaying since some state variables might be updated without affecting the property.
+The problem is particularly pertinent when model-checking models that have sufficient detail for implementation synthesis, as is the goal in our work.
+
+Systems are usually designed in modes such that predicates on state variables are parameterized on which mode the system is in.
+The transition between modes implements a finite state-machine.
+The modes and the state-machine are usually implicit, but a property violation usually requires an unexpected mode transition.
+Abstracting counterexample traces over all of the state variables by traces over the modes helps to identify why a property fails quickly.
+
+Our solution to the problem of constructing a mode-based state machine uses abstract state machines (ASMs).
+ASMs are a generalization of finite state machines originally developed by Gurevich to generalize the Church-Turing thesis for abstract data structures~\cite{}.
+Since then, ASMs have become an engineering technique for simulating a system at different levels of abstraction.
+
+\lee{http://www2.informatik.hu-berlin.de/sam/preprint/reisig203.pdf}
+More precisely, we implement a sequential small-step ASM~\cite{}.
+\begin{definition}[Sequential small-step ASM]
+Let $V$ be a finite set of state variables. Let $s$ be an assignment of variables to values, and let $b$ be a relation Å
+\end{definition}
+
+% ------------------------------------------------------------
 

### 2016-11-08 source/experiments/atom/OM1/AtomOM1.hs
Update AtomOM1 example to the latest atom-sally API and update the generated model

@@ -84,9 +84,10 @@ relay ident inC outCs = period relayPeriod
   done <- bool "done" False
   msg  <- msgVar (tg "relay_msg" ident)
 
-  -- activation condition
-  cond $ isMissing msg  -- we haven't stored a value yet
-  condChannel inC       -- there is a value available
+  -- activation condition:
+  --   we haven't stored a value yet and there is a message waiting
+  --   on the channel 'inC'
+  cond $ (isMissing msg) &&. (fullChannel inC)
 
   -- behavior
   msg  <== readChannel inC
@@ -104,15 +105,14 @@ recv :: Int           -- ^ receiver id
      -> V MsgType
      -> Atom ()
 recv ident inCs vote = period recvPeriod
-                . atom (tg "recv" ident) $ do
+                     . atom (tg "recv" ident) $ do
   done <- bool "done" False
   buffer <- mapM msgVar [ tg (tg "buffer" ident) i | i <- relaySet ]
 
   -- declare multiple "pollers", one for each buffer location
   forM_ relaySet $ \i -> do
     atom (tg2 "recv_poll" ident i) $ do
-      cond $ isMissing (buffer !! i)
-      condChannel (inCs !! i)
+      cond $ (isMissing (buffer !! i)) &&. (fullChannel (inCs !! i))
       (buffer !! i) <== readChannel (inCs !! i)
 
   -- declare a voter
@@ -153,6 +153,7 @@ observer = period observerPeriod
 -- Messages ------------------------------------------------------------
 
 type MsgType = Int64
+msgType = Int64
 
 -- | Specially designated intended message to be send in the absense of faults
 goodMsg :: E MsgType
@@ -170,7 +171,7 @@ isMissing = (==. missingMsgValueE) . value
 
 -- | Declare a new channel with 'missingMsgValue' as its initial value
 newChannel :: String -> Atom (ChanInput, ChanOutput)
-newChannel = flip channel missingMsgValue
+newChannel = flip channel msgType
 
 -- | Declare a variable of message type and add a probe for it to the
 -- environment

### 2016-11-08 source/experiments/atom/OM1/Main.hs

@@ -11,7 +11,7 @@ main = do
   compileOM1ToC
 
   putStrLn "Compiling OM1 to Sally... (om1.mcmt)"
-  compileToSally "om1" TrConfig "om1.mcmt" om1 Nothing
+  compileToSally "om1" defaultCfg "om1.mcmt" om1 Nothing
 
   putStrLn "Done."
 
@@ -22,8 +22,8 @@ main = do
 -- Also print out info on the generated schedule.
 compileOM1ToC :: IO ()
 compileOM1ToC = do
-  (sched, _, _, _, _) <- compile "om1" cfg om1
-  putStrLn $ reportSchedule sched
+  res <- compile "om1" cfg om1
+  putStrLn $ reportSchedule (compSchedule res)
   where
     cfg = defaults { cCode = prePostCode }
 

### 2016-11-08 source/experiments/atom/OM1/AtomOM1.hs
Change whitespace and comments

@@ -1,3 +1,15 @@
+-- |
+-- Module      :  AtomOM1
+-- Copyright   :  Benjamin Jones 2016
+-- License     :  BSD3
+--
+-- Maintainer  :  bjones@galois.com
+-- Stability   :  experimental
+-- Portability :  unknown
+--
+-- A specification for the distributed, fault tolerant system OM(1) written
+-- using Atom (https://github.com/GaloisInc/atom).
+--
 module AtomOM1
   ( om1
   )
@@ -9,8 +21,8 @@ import Data.Int
 import Language.Atom
 import Language.Sally
 
--- Parameters ----------------------------------------------------------
 
+-- Parameters ----------------------------------------------------------
 
 -- Node clock periods (in ticks)
 initPeriod     = 100
@@ -149,8 +161,8 @@ observer = period observerPeriod
   ps <- probes
   mapM_ printProbe ps
 
--- Miscellaneous helper functions and definittions ---------------------
--- Messages ------------------------------------------------------------
+
+-- Helper functions and definitions for Channels and Messages ----------
 
 type MsgType = Int64
 msgType = Int64

### 2016-11-08 source/experiments/atom/smp/AtomSMP.hs
Update AtomSMP to new atom-sally/atom APIs

@@ -23,6 +23,7 @@ recvPeriod     = 5
 -- Messages ------------------------------------------------------------
 
 type MsgType = Int64
+msgType = Int64
 
 -- | Special message value indicating "no message present"
 missingMsgValue :: MsgType
@@ -42,8 +43,8 @@ goodMsg' = Const 2
 smp :: Atom ()
 smp = do
   -- setup channel for communication between source and receiver
-  (s2rIn, s2rOut)   <- channel "s2r_1" missingMsgValue
-  (s2rIn', s2rOut') <- channel "s2r_2" missingMsgValue
+  (s2rIn, s2rOut)   <- channel "s2r_1" msgType
+  (s2rIn', s2rOut') <- channel "s2r_2" msgType
 
   -- declare system nodes:
 
@@ -94,7 +95,7 @@ recv nm c = period recvPeriod . atom nm $ do
   vote <- int64 (nm ++ "_vote") missingMsgValue
   probe (nm ++ "_vote") (value vote)
 
-  condChannel c           -- execute only if channel 'c' has new message
+  cond $ fullChannel c    -- execute only if channel 'c' has new message
   vote <== readChannel c  -- read message and assign to 'vote'
   done <== Const True
 

### 2016-11-08 source/experiments/atom/smp/Main.hs

@@ -10,7 +10,7 @@ main = do
   putStrLn "Compiling SMP to C... (smp.{c,h})"
   compileSMPToC
   putStrLn "Compiling SMP to Sally... (smp.mcmt)"
-  compileToSally "smp" TrConfig "smp.mcmt" smp Nothing
+  compileToSally "smp" defaultCfg "smp.mcmt" smp Nothing
   putStrLn "Done."
 
 
@@ -20,8 +20,8 @@ main = do
 -- Also print out info on the generated schedule.
 compileSMPToC :: IO ()
 compileSMPToC = do
-  (sched, _, _, _, _) <- compile "smp" cfg smp
-  putStrLn $ reportSchedule sched
+  res <- compile "smp" cfg smp
+  putStrLn $ reportSchedule (compSchedule res)
   where
     cfg = defaults { cCode = prePostCode }
 

### 2016-11-09 doc/om-paper/paper.tex
paper checkpoint

@@
 
-\title{Systematic Specification for Scalable Verification of Fault-Tolerant Algorithms}
+\title{Systematic Models for $k$-induction and Synthesis of Fault-Tolerant Distributed Systems}
 \author{Ben and Lee}
@@
 
+\lee{needs to be redone.}
 Fault-tolerant distributed systems are the backbone of reliable life-critical systems, like avionics.
@@
 
-In contrast, model-checking is automated but has limited scalability. To address scalability, abstractions and optimizations are employed, both of which lead to problems.
+In contrast, model-checking is automated but has limited scalability. To address scalability, model-specific abstractions are employed, both of can lead to problems.
 
@@
 
-An example of a verification optimization is to model message passing as shared state between processes to reduce the number of state variables in the model, since the message passing mechanism is elided.
+Another example is to model message passing as shared state between processes to reduce the number of state variables in the model, since the message passing mechanism is elided.
 While these optimizations might be safe under particular assumptions, they can lead to unsound models when a system or its environment is modified.
@@
 Here we describe our formal model specialized for fault-tolerant distributed systems.
-There are three principal abstractions we describe: calendar automata, synchronous observers, and abstract state machines; we describe each below.
+The model draws on three principal abstractions: calendar automata, synchronous observers, and abstract state machines; we describe each below.
 
@@
 
-We prove some intermediate results from the laws. First, in every state, the timeouts are never in the past:
+From the definitions, it follows that in every state, the timeouts are never in the past:
 \begin{lemma}[Future timeouts]\label{lem:ft}
@@
 
-Second, we prove that time is monotonic:
+\noindent
+Furthermore, time is monotonic:
 \begin{lemma}[Monotonic time]
@@
 
-Following Dutetre and Sorea, we associate calendar events with channels in a distributed system~\cite{dutetre}.
+Again following Dutetre and Sorea, we associate calendar events with channels in a distributed system~\cite{dutetre}.
 Specializing calendars to message passing does not lose generality since all external communication from an individual process can be abstracted as message passing.
-Furthermore, typical fault models can be abstracted to act over channels rather than processes~\cite{abstractions}.
+Furthermore, fault models can be abstracted to act over channels rather than processes~\cite{abstractions}.
 The calendar introduces real-time constraints on when processes send and receive messages.
@@
 \noindent
-where `$|$' denotes synchronous composition of transition systems.
+where `$|$' denotes synchronous composition of transition systems (defined formally below).
 $\mathcal{S}$ is the \emph{system}, or machine to be realized in an implementation.
@@
 
-To give a more precise treatment, we introduce some definitions.
+We define synchronous composition more precisely below, beginning with some preliminary definitions:
 
@@
 
-\begin{definition}[Reachable]
+\begin{definition}[Reachable states]
 For a transition relation $\rightarrow$ over the states of $S$, for states $s_0, s_1, \ldots, s_k \in S$, $s_k$ is reachable from $s_0$ iff
@@
 First, the system $\mathcal{S}$ contains no extraneous modeling context about the environment, assumptions, or requirements, making it easier to decompose $\mathcal{S}$ and to synthesize an implementation directly from it.
-Second, as described by Rushby~\cite{}, specifying requirements is simpler.
+Second, as described by Rushby~\cite{}, requirements can be specified using transition systems rather than temporal logic formula, which is often more natural.
 
@@
 \lee{We use synchronous composition to implement fault injection. Mention in the intro. This is fundamental to the model, so we have section devoted to it.}
-The typical approach to injecting faults into a formal model of a distributed system is to add a new state variables to each process representing its fault state.
+In Section~\ref{sec:sync}, we mentioned an environment transition system synchronously composed with the requirements, system, and assumptions.
+A particularly important part of the environment is the model of faults, which we address here.
+The typical approach to injecting faults into a formal model of a distributed system is to add new state variables (or increase the domain of existing state variables) to each process representing its fault state.
 Then additional state variables are added to represent the nondeterminism introduced by faults into the state or messages by a process.
-Finally, additional transitions are added to each process to modify the process's state or messages based on it's current fault state.
 
@@
 
-Not only is modeling faults expensive, but it is generally non-compositional.
-For example, in a classic model-checking model of the fault-tolerant Oral Messages algorithm~\cite{}, every transition of the sending processes updates environmental state variables associated with the fault state.
+Not only is modeling faults expensive, but it is generally non-compositional insofar as the fault model pervades the entire model; for example, consider the model of Oral Messages presented by Rushby~\cite{}.
 
@@
 \subsection{Abstract State Machines}\label{sec:abstract}\label{sec:asms}
-When model-checking a system, counterexamples to properties are traces of states.
+In model-checking, counterexamples to properties are traces of states.
 A state is a collection of variable values, and in large models, there are a lot.
@@
 ASMs are a generalization of finite state machines originally developed by Gurevich to generalize the Church-Turing thesis for abstract data structures~\cite{}.
-Since then, ASMs have become an engineering technique for simulating a system at different levels of abstraction.
+Since then, ASMs have become an engineering technique for simulating systems at different levels of abstraction.
 
@@
 More precisely, we implement a sequential small-step ASM~\cite{}.
-\begin{definition}[Sequential small-step ASM]
-Let $V$ be a finite set of state variables. Let $s$ be an assignment of variables to values, and let $b$ be a relation Å
+\begin{definition}[Abstract state machine]
+Let $\mathcal{M} = (S, I, \rightarrow), \mathcal{M}' = (S', I', \rightarrow')$ be transition systems.
+Let $B \subseteq 2^S$ be a set of relations on $S$.
+We say that $\mathcal{M'}$ is an abstract state machine for $\mathcal{M} | \mathcal{M'}$ iff $S \cap S' = \emptyset$, and each transition $t \in \rightarrow'$ is of the form
+$$\textnormal{if } b_0(s_0) \textnormal{ then } s'_0 \textnormal{ else if } b_1(s_1) \textnormal{ then } s'_1 \ldots \textnormal{ else } b_n(s_n) \textnormal{ then } s'_n$$
 \end{definition}

### 2016-11-10 doc/om-paper/paper.tex
paper wibbles

@@
 
-\title{Systematic Models for $k$-induction and Synthesis of Fault-Tolerant Distributed Systems}
+\title{Systematic Models for Inductive Verification of Fault-Tolerant Distributed Systems}
 \author{Ben and Lee}
@@
 \lee{needs to be redone.}
-Fault-tolerant distributed systems are the backbone of reliable life-critical systems, like avionics.
+Fault-tolerant distributed systems are famously complex to design and build, yet are the  backbone of life-critical systems.
 Consequently, this class of systems demand high-assurance of correct design and implementation.
@@
 
-There is substantial literature on verifying fault-tolerant distributed systems~\cite{}.
+There is substantial literature on verifying fault-tolerant distributed systems.
 Most of this work uses interactive theorem-proving~\cite{}.
@@
 
-In contrast, model-checking is automated but has limited scalability. To address scalability, model-specific abstractions are employed, both of can lead to problems.
+In contrast, model-checking for this class of systems is automated~\cite{}, but has limited scalability, particularly in the context of modeling the nondetermism of faults and distributed computation.
+The way that scalability is usually addressed is by specialization and ad-hoc model-specific abstractions.
+Specialization includes fixing a (small number) of nodes and topology for distributed protocols that are parameterized by the number of nodes and network.
 
-For example, verification abstractions are often ad-hoc and system-specific.
-For example, to model a process that can broadcast the values $\{0, 1\}$, or no value when faulty, one might define an enumerated type of broadcast values sent such as
+Two examples of model-specific abstractions include the following: (1) Suppose we model a system in which a process that broadcasts the values $\{0, 1\}$ in the nominal case, but when the process suffers a fault, it broadcasts no value.
+To model the system, it is typical to define an enumerated type of broadcast values as $$\{0, 1, NONE\}$.
+However, the enumeration conflates the environment (i.e., faults) and system (i.e., the non-faulty broadcast values).
+(2) Another example is to model message passing as shared state between processes to reduce the number of state variables in the model, since the message passing mechanism is elided.
+While these optimizations might be safe under particular assumptions, they require careful meta-level reasoning and justification and divorce the model from any realization of it.
 
-$$\{0, 1, NONE\}$$
+However, recent advances in model-checking allow for better scalability and model fidelity, and they begin to blur the lines between model-checking and interactive theorem-proving.
+In particular, $k$-induction for infinite-state systems allows the specifier to assert $k$-inductive lemmas
 
-However, the enumeration conflates environmental artifacts (i.e., faults) and system artifacts (i.e., the non-faulty broadcast values).
-Conflating them makes formal model-to-code correspondence more difficult to achieve.
-
-Another example is to model message passing as shared state between processes to reduce the number of state variables in the model, since the message passing mechanism is elided.
-While these optimizations might be safe under particular assumptions, they can lead to unsound models when a system or its environment is modified.
+  (We discuss the applicability of property-directed reachability in Section~\ref{sec:related}.)
 
@@
 
+\lee{Talk about how to define an MFA in the environment}
+
 % ------------------------------------------------------------
@@
 Our solution to the problem of constructing a mode-based state machine uses abstract state machines (ASMs).
-ASMs are a generalization of finite state machines originally developed by Gurevich to generalize the Church-Turing thesis for abstract data structures~\cite{}.
-Since then, ASMs have become an engineering technique for simulating systems at different levels of abstraction.
+ASMs are a generalization of finite state machines originally developed by Gurevich to generalize state machines to operate over arbitrary data structures~\cite{asm}.
 
-\lee{http://www2.informatik.hu-berlin.de/sam/preprint/reisig203.pdf}
-More precisely, we implement a sequential small-step ASM~\cite{}.
-\begin{definition}[Abstract state machine]
-Let $\mathcal{M} = (S, I, \rightarrow), \mathcal{M}' = (S', I', \rightarrow')$ be transition systems.
-Let $B \subseteq 2^S$ be a set of relations on $S$.
-We say that $\mathcal{M'}$ is an abstract state machine for $\mathcal{M} | \mathcal{M'}$ iff $S \cap S' = \emptyset$, and each transition $t \in \rightarrow'$ is of the form
+In our context, we require only a restricted form of ASMs known as a \emph{sequential small-step ASM}:
+
+\begin{definition}[Abstract state machine (ASM)]
+Let $\mathcal{M} = (S, I, \rightarrow)$ be a transition system and let $B \subseteq 2^S$ be a set of predicates over $S$.
+Then $\mathcal{M}$ is an ASM iff the transition relation $\rightarrow$ is of the form
 $$\textnormal{if } b_0(s_0) \textnormal{ then } s'_0 \textnormal{ else if } b_1(s_1) \textnormal{ then } s'_1 \ldots \textnormal{ else } b_n(s_n) \textnormal{ then } s'_n$$
+\noindent
+where $b_0, b_1, \ldots, b_n$ are arbitrary Boolean combinations of the predicates in $B$.
 \end{definition}
+
+Our use of ASMs to improve traceability is even more restrictive.
+We construct an ASM $\mathcal{M}$ that is synchronously composed with the system specification $\mathcal{S}$.
+Each state of the ASM represents a ``configurations'' of $S$ and is associated with some predicate over the states of $\mathcal{S}$.
 
@@
 % ------------------------------------------------------------
-\section{Extended Example: OM(1)}\label{sec:byz}
+\section{Putting it Together: Compositional Modeling and Verification}\label{sec:byz}
 

### 2016-11-19 doc/om-paper/paper.tex
paper: more updates.

@@
 Two examples of model-specific abstractions include the following: (1) Suppose we model a system in which a process that broadcasts the values $\{0, 1\}$ in the nominal case, but when the process suffers a fault, it broadcasts no value.
-To model the system, it is typical to define an enumerated type of broadcast values as $$\{0, 1, NONE\}$.
-However, the enumeration conflates the environment (i.e., faults) and system (i.e., the non-faulty broadcast values).
+To model the system, it is typical to define an enumerated type of broadcast values as $$\{0, 1, NONE\}$$ However, the enumeration conflates the environment (i.e., faults) and system (i.e., the non-faulty broadcast values).
 (2) Another example is to model message passing as shared state between processes to reduce the number of state variables in the model, since the message passing mechanism is elided.
@@
 Note focus is on infinite-state model-checking: will need user-provided lemmas.
-compositional approach to providing lemmas }
+compositional approach to providing lemmas.
+note verification interesting in its own right }
 
@@
 % ------------------------------------------------------------
-\section{Putting it Together: Compositional Modeling and Verification}\label{sec:byz}
+\section{Putting it Together: Compositional Modeling and Verification for Oral Messages}\label{sec:byz}
 
-We have instantiated the formal model described above to model and verify the classic OM(1) Byzantine fault-tolerant algorithm~\cite{}.
-OM(1) and its variants have been verified in other settings, allowing us to benchmark our results against others (see Section~\ref{sec:experimental} for experimental results).
+To demonstrate the model and techniques, we model the classic OM(1) Byzantine fault-tolerant algorithm~\cite{} in our framework.
+OM(1) has been modeled in a variety of formal tools~\cite{}, providing a nice comparison with our work.
+Then, we make modifications to the OM(1) model, and show the model's resilience to those changes.
+By ``resilience'', we mean that only a small, localized portion of the must be modified, and only a small set of lemmas must be modified.
+
+The dimensions along which we make changes are the following: the fault model, topology, local node behavior, and synchronization. \lee{Did we complete these?}
 
 The verification is carried out in the Symbolic Analysis Laboratory~\cite{}, which contains a suite of model-checkers and a high-level specification language.
-In our work, we use the infinite-state model-checker using $k$-induction~\cite{}.
+In our work, we use infinite-state (SMT-based) $k$-induction~\cite{}.
 
-\subsection{OM(1) Overview}
-Our first case is the ``Oral Messages'' algorithm to solve the Byzantine
-Generals Problem~\cite{Lamport-OM}. The algorithm, $\OM{m}$, is parameterized by
-the number of rounds of message relay. Here, we focus on Oral Messages with one
-round of communication, $\OM{1}$, that tolerates a single Byzantine fault, so
-long as there are a sufficient number of nodes nodes.
+The verification is interesting in its own right, as it is the first $k$-induction based verification over a fully-parameterized OM(1) model (we describe this in more detail in Section~\ref{sec:related}).
 
-Figure~\ref{fig:om1} depicts the communication pattern for \OM{1}. Each of the
-$n$ nodes in the system represents a (Byzantine) General. In this formulation, the
-center node is the commanding general, while the $n-1$ outer generals are the
-lieutenants. The algorithm starts with the commanding general sending each
-lieutenant the same message $v$. Upon receipt of the commander's message, each
-lieutenant in turn sends the message received to each of the other lieutenants.
-Once a lieutenant has received all $n-1$ expected messages, a majority vote is
-taken among the $n$ messages and the lieutenant declares its output (course of
-action) to be whichever message is in the majority.
+\lee{also note that OM(1) underlies many real-world systems}
 
-\begin{figure}[ht]
-\centering
-\begin{tikzpicture}[
-ell/.style={x radius=1, y radius=0.5, draw, shape=circle},
-arr/.style={semithick}]
-\node[ell] (c)  at (0,0)   {Commander};
-\node[ell] (l1) at (0:4)   {Lieutenant 1};
-\node[ell] (l2) at (120:4) {Lieutenant 2};
-\node[ell] (l3) [fill=gray!30] at (240:4) {Lieutenant 3};
-%
-\draw[->, arr] (c) -- node[auto] {$v$} (l1);
-\draw[->, arr] (c) -- node[auto] {$v$} (l2);
-\draw[->, arr] (c) -- node[auto] {$v$} (l3);
-\draw[<->, arr] (l1) -- node[auto,swap,near start] {$v$}
-                        node[auto,swap,near end] {$v$} (l2);
-\draw[<->, arr] (l1) -- node[auto,near start,color=red] {$x$}
-                        node[auto,near end] {$v$} (l3);
-\draw[<->, arr] (l2) -- node[auto,swap,near start,color=red] {$x$}
-                        node[auto,swap,near end] {$v$} (l3);
-\end{tikzpicture}
-\caption{\OM{1} Four generals, one traitor}
-\label{fig:om1}
-\end{figure}
+\subsection{OM(1) Algorithm}
+The ``Oral Messages'' algorithm solves the Byzantine Generals
+Problem~\cite{Lamport-OM}.
+The algorithm, $\OM{m}$, is a recursive algorithm, parameterized by the number of rounds of communication, $m$.
+Consider a finite set of nodes $N$.
+Distinguish one node as the general, $g$, and the remaining nodes $L = N \setminus \{g\}$ as the lieutenants.
+The general will broadcast a value to the lieutenants.
+The goal of the algorithm is for each of the lieutenants to agree on the value sent by $g$ in the presence of a Byzantine faulty general or lieutenants.
+More precisely, we assume the identity of a sender cannot be spoofed, broadcast communication proceeds in synchronous rounds such that the absence of a message in a round can be detected by the receiver (let $None$ be a constant used when no message is received), and a Byzantine faulty sender can arbitrary corrupt or not send a message, but there are no other faults in the system.
 
-For a fault model we follow~\cite{Rushby-OM1} and assume that nodes are the
-only source of faults. Moreover, node faults are assumed to be either
-\emph{manifest}, \emph{symmetric}, or \emph{byzantine}. A fault is called manifest if
-it is detectable by the non-faulty nodes; for example a node that sends to
-another node a message that explicitly indicates there is a fault. On the
-other hand, a fault is called symmetric if its presence is not necessarily
-detectable as a fault to the other components; for example a node which sends
-wrong messages, as opposed to invalid or missing messages.
+The algorithm proceeds as follows:
 
-Given a formal model of \OM{1}, the standard target for verification is
-\emph{validity} and \emph{agreement}. If we let $l_i$ denote the output for
-lieutenant $i$ (where $1 \le i \le n-1$), then validity states:
+\begin{itemize}
+\item {\bf $OM(0)$}: $g$ broadcasts a value to each lieutenant, and the lieutenants return the value received (or $None$).
+\item {\bf $OM(m)$, $m > 0$}:
+  \begin{enumerate}
+  \item $g$ broadcasts a value to each lieutenant, $l$.
+  \item\label{om:two} Let $v$ be the value received by $l$. Then for each $l$, execute $OM(m-1)$ with general $l$ and lieutenants $L \setminus \{l\}$.
+  \item For each lieutenant $l$, output the majority value received from the other lieutenants.
+  \end{enumerate}
+\end{itemize}
+
+\noindent
+Therefore, $OM(1)$ includes two rounds of broadcast communication, one in which the general broadcasts, and one in which the lieutenants exchange their values.
+
+%% parameterized by the
+%% number of rounds of message relay. Here, we focus on Oral Messages with one
+%% round of communication, $\OM{1}$, that tolerates a single Byzantine fault \todo{true?}, so
+%% long as there are a sufficient number of nodes nodes.
+
+%% Each of the $n$ nodes in the system represents a (Byzantine) General. In this
+%% formulation, the center node is the commanding general, while the $n-1$ outer
+%% generals are the lieutenants. The algorithm starts with the commanding general
+%% sending each lieutenant the same message $v$. Upon receipt of the commander's
+%% message, each lieutenant in turn sends the message received to each of the other
+%% lieutenants.  Once a lieutenant has received all $n-1$ expected messages, a
+%% majority vote is taken among the $n$ messages and the lieutenant declares its
+%% output (course of action) to be whichever message is in the majority.
+
+%% \begin{figure}[ht]
+%% \centering
+%% \begin{tikzpicture}[
+%% ell/.style={x radius=1, y radius=0.5, draw, shape=circle},
+%% arr/.style={semithick}]
+%% \node[ell] (c)  at (0,0)   {Commander};
+%% \node[ell] (l1) at (0:4)   {Lieutenant 1};
+%% \node[ell] (l2) at (120:4) {Lieutenant 2};
+%% \node[ell] (l3) [fill=gray!30] at (240:4) {Lieutenant 3};
+%% %
+%% \draw[->, arr] (c) -- node[auto] {$v$} (l1);
+%% \draw[->, arr] (c) -- node[auto] {$v$} (l2);
+%% \draw[->, arr] (c) -- node[auto] {$v$} (l3);
+%% \draw[<->, arr] (l1) -- node[auto,swap,near start] {$v$}
+%%                         node[auto,swap,near end] {$v$} (l2);
+%% \draw[<->, arr] (l1) -- node[auto,near start,color=red] {$x$}
+%%                         node[auto,near end] {$v$} (l3);
+%% \draw[<->, arr] (l2) -- node[auto,swap,near start,color=red] {$x$}
+%%                         node[auto,swap,near end] {$v$} (l3);
+%% \end{tikzpicture}
+%% \caption{\OM{1} Four generals, one traitor}
+%% \label{fig:om1}
+%% \end{figure}
+
+$OM(1)$ is designed to preserve \emph{validity} and \emph{agreement}
+properties.
+Validity states that if the general is non-faulty, then every lieutenant outputs the value sent by the general.
+More formally, let the general broadcast $v$, and let $l_i$ denote the output for lieutenant $i$:
 %
 \begin{equation}
+  \tag{Validity}
     \forall \,i. \quad l_i = v
@@
 %
-whereas agreement states:
+Agreement states that each lieutenant outputs the same value. Let $l_i, l_j$ denote the outputs of lieutenants $i, j \in L$, respectively:
 %
 \begin{equation}
+  \tag{Agreement}
     \forall \,i, j. \quad l_i = l_j
@@
 %
-where $i,j$ range over the non-faulty nodes.
 
-It is a classical result that \OM{1} can tolerate at most 1 (byzantine) faulty
-node as long as there are at least 4 nodes in total. By adding additional
-communication rounds, \OM{1} can be extended to an algorithm \OM{m} which
-tolerates at most $m$ faults. On the other hand it is known that any system
-which tolerates $m$ byzantine faults must involve at least $3m+1$ nodes.
+$OM(m)$ was originally designed to tolerate Byzantine faults. It is classic
+result that $OM(m)$ satisifies validity and agreement if the number of
+nodes is at least $3m+1$, and there no more than $m$
+Byzantine faulty nodes~\cite{}.
 
-In the results above there are several assumptions made regarding the
-underlying computational platform which aren't obvious from the informal
-description, but become so when one considers formally modelling it. To make
-as many assumptions as possible explicit, we require:
-\begin{enumerate}
-    \item every message sent by a node is received successfully by the
-        addressed node,
-    \item when a node receives a message it may determine who sent it,
-    \item each node can detect the absence of a message.
-\end{enumerate}
-These are the assumptions made by Lamport in his constructed solution for
-$\OM{m}$ and the corresponding impossibility result (see
-\cite{Lamport-OM}~\S2). It follows that a system implementing $\OM{1}$ must
-transition synchronously. In particular, in an asynchronous system there is no
-general method for detecting when a message is absent.
+The fault model of $OM(1)$ has been extended to hybrid fault model described in Section~\ref{sec:kibitzer}.
+Under that fault model, the hypothesis for proving validity and agreement can be weakened to assuming there must be at least $2a+2s+b+1
 
@@
 \section{Related Work}\label{sec:related}
+\lee{talk about om1 verification, comparing it to others.}
 

### 2016-11-22 doc/om-paper/paper.tex
paper: some bib entries

@@
 
-\title{Systematic Models for Inductive Verification of Fault-Tolerant Distributed Systems}
+\title{Model-Checking Fault-Tolerant Distributed System Implementations}
 \author{Ben and Lee}
@@
 \begin{abstract}
-  Some abstract.
+We present a modeling framework for fault-tolerant distributed system verification.
 \end{abstract}
@@
 
-This straightforward approach introduces substantial additional state and nondeterminism, reducing the scalability of model-checking.
+%% This straightforward approach introduces additional state and nondeterminism, reducing the scalability of model-checking.
 Modeling faults is particularly expensive given that additional state and nondeterminism is added to each process.
@@
 \begin{quote}
-To achieve the full range of faulty behaviors, it seems that a faulty source should be able to send a different incorrect value to each relay, and this requires n different values.
-It might seem that we need some additional incorrect values 3 so that faulty relays can exhibit their full range of behaviors.
+To achieve the full range of faulty behaviors, it seems that a faulty source should be able to send a \emph{different} incorrect value to each relay, and this requires $n$ different values.
+It might seem that we need some additional incorrect values so that faulty relays can exhibit their full range of behaviors.
 It would certainly be safe to introduce additional values for this purpose, but the performance of model checking is very sensitive to the size of the state space, so there is a countervailing argument against introducing additional values.
@@
 
-$OM(1)$ is designed to preserve \emph{validity} and \emph{agreement}
+OM(1) is designed to preserve \emph{validity} and \emph{agreement}
 properties.
@@
 The fault model of $OM(1)$ has been extended to hybrid fault model described in Section~\ref{sec:kibitzer}.
-Under that fault model, the hypothesis for proving validity and agreement can be weakened to assuming there must be at least $2a+2s+b+1
+Under that fault model, the hypothesis for proving validity and agreement can be weakened to assuming there must be at least $2a+2s+b+1$ nodes, where $a$ is the number of Byzantine (or asymmetric) faults, $s$ is the number of symmetric faults, and $b$ is the number of benign faults.
+\lee{is the right?}
 
@@
 \section{Toward an Architectural Domain-Specific Language}\label{sec:adsl}
+Not sure if we'll have space for this. Maybe just move to the conclusions?
 
@@
 \section{Related Work}\label{sec:related}
-\lee{talk about om1 verification, comparing it to others.}
+
+There is a long history of formally verifying fault-tolerant distributed systems.
+Most previous work has used mechanical theorem-proving, given the complexity of these systems and models.
+OM(1) is a particularly good case-study that has a history of verification.
+Most
 

### 2016-11-23 doc/om-paper/paper.tex
additional references.

@@
 
-\title{Model-Checking Fault-Tolerant Distributed System Implementations}
+\title{Compositional Model-Checking for Fault-Tolerant Distributed System Implementations}
 \author{Ben and Lee}
@@
 \begin{abstract}
-We present a modeling framework for fault-tolerant distributed system verification.
+With proof techniques like property-directed reachability (PDR) and $k$-induction with user-supplied invariants, model-checking scales further than ever before.
+Still, fault-tolerant distributed systems are particularly challenging to model-check given their large state space and non-determinism.
+The typical approach to controlling complexity is to construct ad-hoc abstractions of faults, message-passing, and behavior to simplify their behavior.
+However, these abstractions come at the price of divorcing the model from its implementation and making refactoring difficult.
+In this work, we present a systematic model for fault-tolerant distributed system verification that combines calendar automata, synchronous observers, and abstract state machines specialized for model-checking.
+We demonstrate the efficacy of the framework by presenting a the first model-checked implementation of the Oral Messages algorithm then demonstrate how modifications along axes such as the fault model, topology, node behavior, and require only local modifications to the model and proofs.
 \end{abstract}
@@
 
-There is a long history of formally verifying fault-tolerant distributed systems.
-Most previous work has used mechanical theorem-proving, given the complexity of these systems and models.
-OM(1) is a particularly good case-study that has a history of verification.
-Most
+The Oral Messages algorithm was presented in the landmark paper by Lamport, Shostak, and Pease~\cite{om}.
+Later, a hybrid fault-model and modified Oral Messages algorithm was developed by Thambidurai and Park~\cite{hybrid}.
+Using the PVS interactive theorem-prover~\cite{pvs}, Lincoln and Rushby demonstrate a flaw in Thambidurai and Park's algorithm and present a fix to it~\cite{csl-93-2}.
+Gasc{\'{o}}n and Tiwari use SAT-based bounded synthesis to partially synthesize a variant of Oral Messages that tolerates transient faults~\cite{om1-synth}.
+Bokor~\emph{et~al.} describe a message-passing model for synchronous distributed algorithms that is particularly amenable to partial-order reduction for explicit-state model-checking~\cite{Bokor2010}.
 
@@
 \section{Conclusions}\label{sec:conclusions}
+
+\lee{talk about compositional verification of properties, even though SAL doesn't directly support it}
 

### 2016-11-23 doc/om-paper/paper.tex
more related work.

@@
 \begin{abstract}
-With proof techniques like property-directed reachability (PDR) and $k$-induction with user-supplied invariants, model-checking scales further than ever before.
+With proof techniques like IC3 and $k$-induction with user-supplied invariants, model-checking scales further than ever before.
 Still, fault-tolerant distributed systems are particularly challenging to model-check given their large state space and non-determinism.
@@
 The model combines several formalisms from the literature to efficiently but systematically model and verify a broad class of systems.
-The model extends \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{Dutertre-Sorea}.
+The model extends \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal}.
 A calendar automata is a model inspired by discrete-event simulation, used in testing real-time systems~\cite{}.
@@
 \subsection{OM(1) Algorithm}
+
+\lee{rewrite this to be OMH}
+
 The ``Oral Messages'' algorithm solves the Byzantine Generals
@@
 
+The Oral Messages algorithm and its variants have a long history of formal verification, making it a good test-case for comparing and contrasting verification tools~\cite{pvs}.
 The Oral Messages algorithm was presented in the landmark paper by Lamport, Shostak, and Pease~\cite{om}.
 Later, a hybrid fault-model and modified Oral Messages algorithm was developed by Thambidurai and Park~\cite{hybrid}.
+OM(1) was subsequently verified in both the PVS and ACL2 interactive theorem-provers~\cite{Young97:IC}.
+Also in ACL2, an implementation of a circuit design to implement OM(1) is given~\cite{om-acl2-impl}; the low-level model most closely relates to our level of detail.
+A refinement-based verification approach is used, and OM(1) is specialized to a four-process system.
 Using the PVS interactive theorem-prover~\cite{pvs}, Lincoln and Rushby demonstrate a flaw in Thambidurai and Park's algorithm and present a fix to it~\cite{csl-93-2}.
-Gasc{\'{o}}n and Tiwari use SAT-based bounded synthesis to partially synthesize a variant of Oral Messages that tolerates transient faults~\cite{om1-synth}.
+
+Using automated techniques, Gasc{\'{o}}n and Tiwari use SAT-based bounded synthesis to partially synthesize a variant of Oral Messages that tolerates transient faults~\cite{om1-synth}.
 Bokor~\emph{et~al.} describe a message-passing model for synchronous distributed algorithms that is particularly amenable to partial-order reduction for explicit-state model-checking~\cite{Bokor2010}.
+That said, their model of OM(1) (not OMH(1)), timeouts after 10~hours, and we assume a voting algorithm is not explicitly specified.
+Very recently, Jovanovi{\'{c}} and Dutertre use a ``flattened'' high-level model of OM(1) as a benchmark for IC3 augmented with $k$-induction~\cite{pdr-kind}.
 
@@
 \lee{talk about compositional verification of properties, even though SAL doesn't directly support it}
+\lee{talk about k-induction vs. PDR}
+\lee{talk about lack of axiomatization in model-checking (multiple rushby bugs), but  tradeoff of deadlock. see proglema paper}
 

### 2016-11-23 source/experiments/SAL/OM/runsanity.sh
Add script for convenient checking of sanity properties

@@ -1,0 +1,9 @@
+#!/bin/bash
+#
+# Run sanity checks for the om1_cal.sal model (INFINITE version).
+
+set -e
+
+for i in $(seq 0 4); do
+    sal-inf-bmc --disable-traceability -d 25 'om1_cal{2,2}' "sanity$i"
+done

### 2016-11-24 doc/om-paper/paper.tex
mucked with intro.

@@
 In this work, we present a systematic model for fault-tolerant distributed system verification that combines calendar automata, synchronous observers, and abstract state machines specialized for model-checking.
-We demonstrate the efficacy of the framework by presenting a the first model-checked implementation of the Oral Messages algorithm then demonstrate how modifications along axes such as the fault model, topology, node behavior, and require only local modifications to the model and proofs.
+We demonstrate the efficacy of the framework by presenting the first model-checked \emph{implementation} of the Hybrid Oral Messages algorithm, using a fast majority vote, then demonstrate how modifications along axes such as the fault model, topology, node behavior, and require only local modifications to the model and proofs.
 \end{abstract}
@@
 
-\lee{needs to be redone.}
-Fault-tolerant distributed systems are famously complex to design and build, yet are the  backbone of life-critical systems.
-Consequently, this class of systems demand high-assurance of correct design and implementation.
-Formal verification can be used to prove the correctness of their designs.
+\lee{talk about Amazon/TLA? Paxos made live?}
 
-There is substantial literature on verifying fault-tolerant distributed systems.
-Most of this work uses interactive theorem-proving~\cite{}.
-Interactive theorem-proving is useful in the domain because of the complexity of the models and because distributed system verification is typically parameterized by the number of nodes; automated methods perform poorly along these two metrics.
-However, two problems with interactive verification are that it requires specialized expertise and it is labor intensive.
+Fault-tolerant distributed systems are famously complex, yet are the backbone of life-critical systems, such as commercial avionics.
+Consequently, this class of systems demands high-assurance of correct design and implementation.
+Formal verification can help provide that assurance.
 
-In contrast, model-checking for this class of systems is automated~\cite{}, but has limited scalability, particularly in the context of modeling the nondetermism of faults and distributed computation.
-The way that scalability is usually addressed is by specialization and ad-hoc model-specific abstractions.
-Specialization includes fixing a (small number) of nodes and topology for distributed protocols that are parameterized by the number of nodes and network.
+The verification of this class of systems has usually been at the algorithmic level, eliding details about a concrete implementation, and historically, has relied on formal models verified by interactive theorem-proving~\cite{om-acl2-impl,Young97:IC,csl-93-2,pvs}.
+If formal verification is to be introduced into the workflow of system designers, though, we need automated methods and programmatic verification of \emph{implementations}.
+(Mostly) automated proof techniques are required to reduce the need for specialized verification expertise.
+System designers create software and hardware implementations to test, simulate, and deploy.
+Discrepancies between implementations can formal models can easily arise if the latter is abstracted too much from the former, particularly if those abstractions are ad-hoc and system specific.
+Furthermore, as an implementations are tweaked and modified to explore the design space, it is easy for the formal models to fall out of synch.
 
-Two examples of model-specific abstractions include the following: (1) Suppose we model a system in which a process that broadcasts the values $\{0, 1\}$ in the nominal case, but when the process suffers a fault, it broadcasts no value.
+To address this problem, we envision building a domain-specific language (DSL) for specifying and exploring distributed fault-tolerant system implementations.
+We envision a DSL that generates both code for simulation, testing, and deployment, as well as formal models for verification (we say more about the DSL in Section~\ref{sec:conclusions}).
+
+The first step in realizing this vision, however, and the one we address in this paper, is to develop an intermediate representation (IR) for the DSL to target when generating formal models.
+Our target for the IR is one supported by infinite-state model-checking.
+With proof techniques like IC3~\cite{} and $k$-induction~\cite{}, model-checking scales further than ever before.
+We envision an IR that allows for specifying system behavior, assumptions, and requirements, but also includes invariants that can be systematically generated by the DSL.
+Finally, the IR must support easy refactoring and modifications to the design.
+Making small changes to system specification---for example, modifying the fault model---should require only localized changes to the IR representation, including proofs.
+In short, the IR should be compositional.
+
+In this work, we combine a number of ideas known in the literature and folklore to build scalable and compositional formal models of fault-tolerant distributed systems.
+We describe a generic formal model for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}.
+There are three aspects of the model: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) \emph{synchronous observers}, originally from the synchronous languages community~\cite{} and more recently popularized by Rushby for model-checking; and (3) \emph{abstract state machines} for abstracting complex systems.
+
+To demonstrate the model, we describe a formally-verified implementation of the Oral Messages Hybrid algorithm (OMH)~\cite{}, implementing the Boyer-Moore Fast Majority Vote algorithm~\cite{}.
+The verification is interesting in its own right, as it is the first fully parametric (on the number of nodes) model-checked implementation of the classic algorithm.
+Then, we modify the algorithm along axes such as the fault model, topology, node behavior, and show that only local modifications are required to the model and proofs.
+
+%% Two examples of model-specific abstractions include the following: (1) Suppose we model a system in which a process that broadcasts the values $\{0, 1\}$ in the nominal case, but when the process suffers a fault, it broadcasts no value.
 To model the system, it is typical to define an enumerated type of broadcast values as $$\{0, 1, NONE\}$$ However, the enumeration conflates the environment (i.e., faults) and system (i.e., the non-faulty broadcast values).
@@
 While these optimizations might be safe under particular assumptions, they require careful meta-level reasoning and justification and divorce the model from any realization of it.
-
-However, recent advances in model-checking allow for better scalability and model fidelity, and they begin to blur the lines between model-checking and interactive theorem-proving.
-In particular, $k$-induction for infinite-state systems allows the specifier to assert $k$-inductive lemmas
-
-  (We discuss the applicability of property-directed reachability in Section~\ref{sec:related}.)
-
-Our contributions combine a number of ideas in the formal verification literature to build scalable, generic formal models of fault-tolerant distributed systems in which verification is more automated than in traditional interactive theorem-proving.
-We describe a generic formal model for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}.
-The model combines several formalisms from the literature to efficiently but systematically model and verify a broad class of systems.
-The model extends \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal}.
-A calendar automata is a model inspired by discrete-event simulation, used in testing real-time systems~\cite{}.
-Calendar automata have previously been used to verify bounded asynchronous systems~\cite{}; here we verify a synchronous system.
-
-The model also employs \emph{synchronous observers}~\cite{}.
-Synchronous observers are a modeling abstraction in which an ``observer'' actor is synchronously composed with the system under verification.
-We use synchronous observers to compose distinct transition systems modeling the system, environment, assumptions, and requirements.
-Finally, we use an \emph{abstract state machine}~\cite{}, also composed with the system under verification using synchronous observation.
-The abstract state machine simplifies discovering invariants and debugging counterexamples.
-
-We validate the model with a number of small examples \lee{todo}, as well as an extended case-study, verifying the Oral Messages (1) protocol (OM(1))~\cite{}.
-While our model is more detailed---it explicitly models channels---and also more general---it allows bounded asynchrony between processes---it is the most scalable model-checking verification of the protocol.
 
@@
 
+\lee{talk about atom-sally}
+
+\section*{Acknowledgments}
+\lee{NASA contract, honeywell ppl, rushby's model?}
+
 \bibliographystyle{IEEEtran}

### 2016-11-24 doc/om-paper/paper.tex
compositional --> modular.

@@
 
-\title{Compositional Model-Checking for Fault-Tolerant Distributed System Implementations}
+\title{Modular Model-Checking for Fault-Tolerant Distributed System Implementations}
 \author{Ben and Lee}
@@
 Making small changes to system specification---for example, modifying the fault model---should require only localized changes to the IR representation, including proofs.
-In short, the IR should be compositional.
+In short, the IR should be modular.
 
-In this work, we combine a number of ideas known in the literature and folklore to build scalable and compositional formal models of fault-tolerant distributed systems.
+In this work, we combine a number of ideas known in the literature and folklore to build scalable and modular formal models of fault-tolerant distributed systems.
 We describe a generic formal model for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}.

### 2016-11-25 doc/om-paper/paper.tex
paper reminders
added ben's benchmark results.
add dashes for no data.
reworked intro.

@@
 The verification of this class of systems has usually been at the algorithmic level, eliding details about a concrete implementation, and historically, has relied on formal models verified by interactive theorem-proving~\cite{om-acl2-impl,Young97:IC,csl-93-2,pvs}.
-If formal verification is to be introduced into the workflow of system designers, though, we need automated methods and programmatic verification of \emph{implementations}.
+If formal verification is to be introduced into the workflow of system designers, though, we need more automated methods.
 (Mostly) automated proof techniques are required to reduce the need for specialized verification expertise.
+We also need programmatic verification of \emph{implementations}.
 System designers create software and hardware implementations to test, simulate, and deploy.
-Discrepancies between implementations can formal models can easily arise if the latter is abstracted too much from the former, particularly if those abstractions are ad-hoc and system specific.
-Furthermore, as an implementations are tweaked and modified to explore the design space, it is easy for the formal models to fall out of synch.
+Discrepancies between implementations and algorithmic models can arise if the latter is abstracted too much from the former, particularly if those abstractions are ad-hoc and system specific.
+Furthermore, as an implementations are tweaked and modified to explore the design space, it is easy for the formal models to become inconsistent.
 
-To address this problem, we envision building a domain-specific language (DSL) for specifying and exploring distributed fault-tolerant system implementations.
-We envision a DSL that generates both code for simulation, testing, and deployment, as well as formal models for verification (we say more about the DSL in Section~\ref{sec:conclusions}).
+There are at least two classes of abstractions that separate traditional models of fault-tolerant distributed algorithms from their implementations.
+One is to intertwine the environmental model with the system description.
+For example, the behavior of nodes is naturally specified as a transition system in which transitions are guarded by the node's fault state.
+But faults are part of the environment; in an implementation, does not change behavior based on its faults.
+Another class of abstractions is used to simplify models.
+For example, message passing might be abstracted with shared state, or a node's local behavior---e.g., a majority vote---is elided and instead, the output is constrained by a \emph{specification} of a majority vote.
 
-The first step in realizing this vision, however, and the one we address in this paper, is to develop an intermediate representation (IR) for the DSL to target when generating formal models.
-Our target for the IR is one supported by infinite-state model-checking.
-With proof techniques like IC3~\cite{} and $k$-induction~\cite{}, model-checking scales further than ever before.
-We envision an IR that allows for specifying system behavior, assumptions, and requirements, but also includes invariants that can be systematically generated by the DSL.
-Finally, the IR must support easy refactoring and modifications to the design.
-Making small changes to system specification---for example, modifying the fault model---should require only localized changes to the IR representation, including proofs.
-In short, the IR should be modular.
+In this paper, we develop a formal model to address these concerns.
+We combine ideas from the literature and folklore to build scalable and modular formal models of fault-tolerant distributed systems.
+We describe a generic formal model for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}.
+There are three aspects of the model: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) \emph{synchronous observers}, originally from the synchronous languages community~\cite{} and popularized by Rushby for model-checking; and (3) \emph{abstract state machines}, useful for abstracting counterexamples.
+The model is suitable for infinite-state model-checking.
+With proof techniques like IC3 and $k$-induction, model-checking scales further than ever before~\cite{pdr-kind}.
 
-In this work, we combine a number of ideas known in the literature and folklore to build scalable and modular formal models of fault-tolerant distributed systems.
-We describe a generic formal model for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}.
-There are three aspects of the model: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) \emph{synchronous observers}, originally from the synchronous languages community~\cite{} and more recently popularized by Rushby for model-checking; and (3) \emph{abstract state machines} for abstracting complex systems.
+To demonstrate the model in Section~\ref{sec:byz}, we present a formally-verified implementation of the Oral Messages Hybrid algorithm (OMH)~\cite{}, implementing the Boyer-Moore Fast Majority Vote algorithm for algorithm~\cite{}.
+The verification is interesting in its own right, as it is the first fully parametric (on the number of nodes) model-checked \emph{implementation} of the algorithm.
+In Section~\ref{sec:experimental}, we give performance metrics for the model, and we modify the implementation by changing the fault model, topology, node behavior, and show that only local modifications are required to the model and proofs.
 
-To demonstrate the model, we describe a formally-verified implementation of the Oral Messages Hybrid algorithm (OMH)~\cite{}, implementing the Boyer-Moore Fast Majority Vote algorithm~\cite{}.
-The verification is interesting in its own right, as it is the first fully parametric (on the number of nodes) model-checked implementation of the classic algorithm.
-Then, we modify the algorithm along axes such as the fault model, topology, node behavior, and show that only local modifications are required to the model and proofs.
+%% To address this problem, we envision building a domain-specific language (DSL) for specifying and exploring distributed fault-tolerant system implementations.
+We envision a DSL that generates both code for simulation, testing, and deployment, as well as formal models for verification.
+That way, the formal models verified and code executed stay in synch, and abstractions are systematically made.
+We say more about the DSL in Section~\ref{sec:conclusions}.
 
@@
   \centering
-  \includegraphics[width=8cm]{figs/benchmark}
-  \caption{Comparison of number of nodes vs. verification time.}
-  \label{fig:compare}
+  \begin{tabular}{r||c|c|c|c|c|c|c|c|c|c|}
+      \multicolumn{11}{c}{Receivers} \\
+      \cline{2-11}
+   Relays &   1 &   2 &   3 &   4 &   5 &   6 &   7 &   8 &   9 &   10 \\
+      \hline \hline
+      1   & 13  & 18  & 23  & 29  & 35  & 43  & 54  & 70  & 90  & 129 \\
+      \hline
+      2   & 17  & 26  & 38  & 52  & 68  & 89  & 115 & 159 & 227 & -   \\
+      \hline
+      3   & 21  & 40  & 61  & 85  & 120 & 170 & 246 & 410 & -   & -   \\
+      \hline
+      4   & 27  & 58  & 98  & 195 & 213 & 628 & 855 & -   & -   & -   \\
+      \hline
+      5   & 36  & 93  & 292 & 389 & 1940& TO  & -   & -   & -   & -   \\
+      \hline
+      6   & 56  & 303 & 976 & TO  & TO  & -   & -   & -   & -   & -   \\
+      \hline
+      7   & 114 & 1993& TO  & TO  & -   & -   & -   & -   & -   & -   \\
+      \hline
+      8   & 338 & TO  & TO  & -   & -   & -   & -   & -   & -   & -   \\
+      \hline
+      9   & 895 & TO  & -   & -   & -   & -   & -   & -   & -   & -   \\
+      \hline
+     10   & TO  & -   & -   & -   & -   & -   & -   & -   & -   & -   \\
+      \hline
+  \end{tabular}
+  \label{fig:benchmark}
+  \caption{Benchmark of full proof computation time for OMH(1) implementation. Times are in seconds with a timeout (TO) limit of one hour. Dashes ('-') denote no benchmark was run.}
 \end{figure}
-
-% ------------------------------------------------------------
-\section{Toward an Architectural Domain-Specific Language}\label{sec:adsl}
-Not sure if we'll have space for this. Maybe just move to the conclusions?
 
@@
 
+\url{https://people.eecs.berkeley.edu/~apanda/assets/papers/pldi16.pdf}
+
 % ------------------------------------------------------------

### 2016-11-27 doc/om-paper/paper.tex
updated omh description.

@@
 
+\lee{note that we use calendar automata for non real-time system}
+
 \subsection{Synchronous Observers}\label{sec:sync}
@@
 
-To demonstrate the model and techniques, we model the classic OM(1) Byzantine fault-tolerant algorithm~\cite{} in our framework.
-OM(1) has been modeled in a variety of formal tools~\cite{}, providing a nice comparison with our work.
-Then, we make modifications to the OM(1) model, and show the model's resilience to those changes.
-By ``resilience'', we mean that only a small, localized portion of the must be modified, and only a small set of lemmas must be modified.
+To demonstrate the model and techniques, we model the Oral Messages Hybrid (OMH) algorithm~\cite{}, a more sophisticated version of the class Oral Messages (OM) algorithm.
+OM was first formulated by Lamport, Shostak, and Pease~\cite{om} to achieve agreement in a distributed system while tolerating Byzantine faults.
+OMH was originally developed by Thambidurai and Park~\cite{hybrid} to achieve the same result in the presence of a more sophisticated fault model that includes symmetric and benign faults.
+However, OMH had a bug, as originally formulated, which was corrected and formally verified by Lincoln and Rushby~\cite{csl-93-2}.
 
-The dimensions along which we make changes are the following: the fault model, topology, local node behavior, and synchronization. \lee{Did we complete these?}
+First, we briefly describe the algorithm, sketch our model of the algorithm within the model presented in Section~\ref{sec:model}, then describe it's invariants.
 
-The verification is carried out in the Symbolic Analysis Laboratory~\cite{}, which contains a suite of model-checkers and a high-level specification language.
-In our work, we use infinite-state (SMT-based) $k$-induction~\cite{}.
+\subsection{OMH(1) Algorithm}
+OMH is a recursive algorithm that proceeds in rounds of communication.
+Let $OMH(m)$ be a recursive function implementing the algorithm, parameterized by the number of rounds, $m$.
+Consider a finite set of nodes $N$.
+Distinguish one node as the \emph{general}, $g$, and the remaining nodes $L = N \setminus \{g\}$ as the \emph{lieutenants}.
+We assume the identity of any general or lieutenant cannot be spoofed.
+Broadcast communication proceeds in rounds.
+Denote any message that is detectably faulty (e.g., fails a CRC) or is absent, by $ERR$.
+Additionally, in the algorithm, nodes report on values they have previously received.
+In doing so, nodes must differentiate \emph{reporting} $ERR$ from an $ERR$ itself.
+Let $R$ denote that an error is being reported.
+Finally, let $V$ be a special, designated value.
 
-The verification is interesting in its own right, as it is the first $k$-induction based verification over a fully-parameterized OM(1) model (we describe this in more detail in Section~\ref{sec:related}).
+%% The goal of the algorithm is for each of the lieutenants to agree on the value sent by $g$ in the presence of a Byzantine faulty general or lieutenants.
+More precisely, we assume the identity of a sender cannot be spoofed, broadcast communication proceeds in synchronous rounds such that the absence of a message in a round can be detected by the receiver (let $None$ be a constant used when a message is detectably incorrect, including a lack of message), and a Byzantine faulty sender can arbitrary corrupt or not send a message, but there are no other faults in the system.
 
-\lee{also note that OM(1) underlies many real-world systems}
-
-\subsection{OM(1) Algorithm}
-
-\lee{rewrite this to be OMH}
-
-The ``Oral Messages'' algorithm solves the Byzantine Generals
-Problem~\cite{Lamport-OM}.
-The algorithm, $\OM{m}$, is a recursive algorithm, parameterized by the number of rounds of communication, $m$.
-Consider a finite set of nodes $N$.
-Distinguish one node as the general, $g$, and the remaining nodes $L = N \setminus \{g\}$ as the lieutenants.
-The general will broadcast a value to the lieutenants.
-The goal of the algorithm is for each of the lieutenants to agree on the value sent by $g$ in the presence of a Byzantine faulty general or lieutenants.
-More precisely, we assume the identity of a sender cannot be spoofed, broadcast communication proceeds in synchronous rounds such that the absence of a message in a round can be detected by the receiver (let $None$ be a constant used when no message is received), and a Byzantine faulty sender can arbitrary corrupt or not send a message, but there are no other faults in the system.
-
-The algorithm proceeds as follows:
+The algorithm is recursively defined:
 
 \begin{itemize}
-\item {\bf $OM(0)$}: $g$ broadcasts a value to each lieutenant, and the lieutenants return the value received (or $None$).
+\item {\bf $OM(0)$}: $g$ broadcasts a value to each lieutenant, and the lieutenants return the value received (or $ERR$).
 \item {\bf $OM(m)$, $m > 0$}:
@@
   \item $g$ broadcasts a value to each lieutenant, $l$.
-  \item\label{om:two} Let $v$ be the value received by $l$. Then for each $l$, execute $OM(m-1)$ with general $l$ and lieutenants $L \setminus \{l\}$.
-  \item For each lieutenant $l$, output the majority value received from the other lieutenants.
+  \item\label{om:two} Let $l_v$ be the value received by $l \in L$ from $g$.
+Then for each $l$, execute $OMH(m-1)$, assigning $l$ to be the general and $L \setminus \{l\}$ to be the lieutenants.
+$l$ sends $l_v$ or $R$, if $l_v = ERR$.
+  \item For each lieutenant $l \in L$, remove all $ERR$ values received in Step~\ref{om:two} from executing $OHM(m-1)$.
+Compute the majority value over the remaining values, or $V$ if there is no majority.
+If the majority value is $R$, return $E$.
   \end{enumerate}
@@
 \noindent
-Therefore, $OM(1)$ includes two rounds of broadcast communication, one in which the general broadcasts, and one in which the lieutenants exchange their values.
+In particular, $OMH(1)$ includes two rounds of broadcast communication, one in which the general broadcasts, and one in which the lieutenants exchange their values.
 
@@
 
-OM(1) is designed to preserve \emph{validity} and \emph{agreement}
+OMH is designed to preserve \emph{validity} and \emph{agreement}
 properties.
 Validity states that if the general is non-faulty, then every lieutenant outputs the value sent by the general.
-More formally, let the general broadcast $v$, and let $l_i$ denote the output for lieutenant $i$:
+More formally, let the general broadcast $v \neq ERR$, and let $l_i$ denote the output for lieutenant $i$:
 %
@@
 
-$OM(m)$ was originally designed to tolerate Byzantine faults. It is classic
-result that $OM(m)$ satisifies validity and agreement if the number of
-nodes is at least $3m+1$, and there no more than $m$
-Byzantine faulty nodes~\cite{}.
-
-The fault model of $OM(1)$ has been extended to hybrid fault model described in Section~\ref{sec:kibitzer}.
-Under that fault model, the hypothesis for proving validity and agreement can be weakened to assuming there must be at least $2a+2s+b+1$ nodes, where $a$ is the number of Byzantine (or asymmetric) faults, $s$ is the number of symmetric faults, and $b$ is the number of benign faults.
-\lee{is the right?}
+We described a hybrid fault model in Section~\ref{sec:kibitzer}.
+Under that fault model, agreement and validity hold if $2a+2s+b+1 \geq n$, where $n$ is the total number of nodes, $a$ is the number of Byzantine (or asymmetric) faults, $s$ is the number of symmetric faults, and $b$ is the number of benign faults.
+Additionally, it is required that the number of rounds $m$ be greater or equal to the number of Byzantine faults, $a$~\cite{csl-93-2,hybrid}.
 
 \subsection{Model Sketch}
+The verification is carried out in the Symbolic Analysis Laboratory~\cite{}, which contains a suite of model-checkers and a high-level specification language.
+In our work, we use infinite-state (SMT-based) $k$-induction~\cite{}.
+
+\lee{Maybe talk about MJRTY here. Maybe also talk about relaxing in synchronous communication in model.}
 
 \subsection{Invariants}
+
+\lee{make clear we're agnostic about human-provided invariants. We conjecture some can be auto-generated. Some might be inferred by IC3.}
 
@@
 
-\lee{let's also check out scalability when we ``turn off'' faults.
-Also, how hard is it to turn them off or change the fault model?
-E.g., how many lines of spec need to be changed?
-Compare to rushby's?}
+Here we present two classes of experimental results.
+First, we demonstrate the scalability results of the verification, despite the low-level modeling.
+Then we describe modularity results, demonstrated by making modifications to the model and re-validating the model.
+
+\subsection{Scalability}
 
@@
 
+\lee{let's also check out scalability when we ``turn off'' faults.
+Also, how hard is it to turn them off or change the fault model?
+E.g., how many lines of spec need to be changed?
+Compare to rushby's?}
+
+\subsection{Modular Verification}
+
 % ------------------------------------------------------------
@@
 
+\lee{mention that ASMs similar to disjunction invariants}
+
 % ------------------------------------------------------------

### 2016-11-28 doc/om-paper/paper.tex
commented stuff in the lemmas section.

@@
 \subsection{Invariants}
+To make the proof scalable, we specify inductive invariants to be used by SAL's $k$-induction engine. The invariants are modular, focused on specific aspects of the model.
 
-\lee{make clear we're agnostic about human-provided invariants. We conjecture some can be auto-generated. Some might be inferred by IC3.}
+%% cal:
+%%   cal_invt (-)
+
+%% ASM:
+%%   abstract_all (cal_invt, receiver_mode_invt)
+
+%% faults:
+%%   manifest_relay (-)
+%%   symmetric_relay (-)
+%%   fault_propagation (cal_invt)
+%%   source_to_latch (fault_propagation)
+%%   manifest_source (fault_propagation)
+
+%% receiver:
+%%   missing_buffer_size (cal_invt)
+%%   receiver_mode_invt (missing_buffer_size)
+
+%% voting:
+%%   majority_vote_invt (receiver_mode_invt)
+%%   exists_majority (receiver_mode_invt majority_vote_invt)
+
+%% protocol: (everything :))
+%%   validity
+%%   agreement
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figs/proof-structure}
-  \caption{Proof structure.}
+%%   \begin{tabular}{r|c|c|c|c|c|}
+%%              & (a) Calendar & (b) ASM & (c) Protocol & (d) Voting & (e) Fault model \\
+%%     \hline \hline
+%%       Lemmas & 2            & 1       & 3            & 4          & 5 \\
+%% Dependencies &              & (a), (d),         &              &            &   \\
+%%   \end{tabular}
+  %% \includegraphics[width=\textwidth]{figs/proof-structure}
+  \caption{Invariant classification.}
   \label{fig:proof}
 \end{figure}
+
+The lemmas fall into five categories: calendar automata lemmas, abstract state machine lemmas, protocol behavior lemmas, majority vote lemmas, and fault model lemmas.
+
+\lee{make clear we're agnostic about human-provided invariants. We conjecture some can be auto-generated. Some might be inferred by IC3.}
 
@@
 The proofs are all $k=1$ inductive.
-
-The lemmas are classified into five groups: calendar automata lemmas, abstract state machine lemmas, protocol behavior lemmas, majority vote lemmas, and fault model lemmas.
 

### 2016-11-28 doc/om-paper/paper.tex
more paper stuff.

@@
 \usepackage{tikz}
+\usetikzlibrary{arrows,automata}
+\usepackage{tikz-cd}
+\usetikzlibrary{positioning}
+
 \newcommand{\OM}[1]{\ensuremath{\mathrm{OM}(#1)}}
@@
 
+\lee{note that the ASM is not a refinement.}
+
 % ------------------------------------------------------------
@@
 \lee{Maybe talk about MJRTY here. Maybe also talk about relaxing in synchronous communication in model.}
+\lee{talk about synchrony being loosened, so the general/receivers don't all send/receive at the same time. Talk about lieutenants being ``unrolled'' into relays and receivers, following Rushby.}
 
 \subsection{Invariants}
-To make the proof scalable, we specify inductive invariants to be used by SAL's $k$-induction engine. The invariants are modular, focused on specific aspects of the model.
 
+%% Lemmas and dependencies:
 %% cal:
@@
 %%   manifest_relay (-)
-%%   symmetric_relay (-)
+%%   uninterp_msg_relay (-)
 %%   fault_propagation (cal_invt)
@@
   \centering
-%%   \begin{tabular}{r|c|c|c|c|c|}
-%%              & (a) Calendar & (b) ASM & (c) Protocol & (d) Voting & (e) Fault model \\
-%%     \hline \hline
-%%       Lemmas & 2            & 1       & 3            & 4          & 5 \\
-%% Dependencies &              & (a), (d),         &              &            &   \\
-%%   \end{tabular}
-  %% \includegraphics[width=\textwidth]{figs/proof-structure}
-  \caption{Invariant classification.}
+  \footnotesize
+  \begin{tikzpicture}[->, node distance=2.6cm, auto, shorten >=1pt, bend angle=45,thick]
+    \tikzstyle{every state}=[rectangle, rounded corners]
+
+    \node[state] (calendar) {calendar (5)};
+    \node[state] (lieutenant) [right=1.75cm of calendar] {\begin{tabular}{c}lieutenant\\state-machine (2)\end{tabular}};
+    \node[state] (faults) [below=0.5cm of calendar] {faults (5)};
+    \node[state] (asm) [right=0.5cm of faults] {ASM (1)};
+    \node[state] (voting) [below=0.5cm of lieutenant] {voting (2)};
+
+    \tikzstyle{every node}=[]
+
+    \path
+    (faults) edge [] node {} (calendar)
+    (asm)    edge [] node {} (calendar)
+    (asm)    edge [] node {} (lieutenant)
+    (voting) edge [] node {} (lieutenant);
+
+  \end{tikzpicture}
+  \caption{Invariant classification and dependencies.}
   \label{fig:proof}
@@
 
-The lemmas fall into five categories: calendar automata lemmas, abstract state machine lemmas, protocol behavior lemmas, majority vote lemmas, and fault model lemmas.
+To make the proof scalable, we specify inductive invariants to be used by SAL's $k$-induction engine. The invariants are modular. There are 11 invariants, that fall into five categories:
+\begin{enumerate}
+  \item \emph{Calendar automata}: Lemmas relating to the calendar automata model.
+These include lemmas such as time being monotonic, channels missing messages if there is no calendar event, and only nodes associated with a calendar event may execute their local state machines.
+  \item \emph{Abstract state machine (ASM)}: Lemmas relating the ASM synchronous observer states to the implementation states.
+  \item \emph{Lieutenant local state-machine (SM) behavior}: Lemmas describing the modes of behavior of the lieutenants.
+The major modes of the state machine are receiving messages, then once it has filled its buffer, it votes, and after voting returns the result.
+One additional lemma notes that the messages currently received plus missing messages equals the total number of expected messages.
+  \item \emph{Faults}: Lemmas characterizing the effect of a fault in a single broadcast. For example, a lieutenant reports that it ``received an error'' when it receives a benignly-faulty message.
+  \item \emph{Voting}: Lemmas proving that the Fast Majority Vote algorithm implements a majority vote, if one exists.
+These lemmas are nearly verbatim transcriptions from the journal proofs~\cite{mjrty}.
+\end{enumerate}
+\noindent
+The proof structure is shown in Figure~\ref{fig:proof}.
+The number of lemmas per category are shown in parentheses.
+Arrows denote dependencies.
+For example, the ASM lemmas depend on both the calendar automata and lieutenant state-machine lemmas.
 
@@
 
-The proof structure is shown in Figure~\ref{fig:proof}.
  The two graphs show the structure to prove validity and agreement, respectively.
@@
 
+\lee{note some details still elided for implementation ---e.g., CRCs, etc.}
+
 \section*{Acknowledgments}

### 2016-11-28 doc/om-paper/paper.tex
Minor edits on the abstract

--- 2016-11-28 doc/om-paper/paper.tex
@@
 \begin{abstract}
-With proof techniques like IC3 and $k$-induction with user-supplied invariants, model-checking scales further than ever before.
-Still, fault-tolerant distributed systems are particularly challenging to model-check given their large state space and non-determinism.
-The typical approach to controlling complexity is to construct ad-hoc abstractions of faults, message-passing, and behavior to simplify their behavior.
-However, these abstractions come at the price of divorcing the model from its implementation and making refactoring difficult.
-In this work, we present a systematic model for fault-tolerant distributed system verification that combines calendar automata, synchronous observers, and abstract state machines specialized for model-checking.
-We demonstrate the efficacy of the framework by presenting the first model-checked \emph{implementation} of the Hybrid Oral Messages algorithm, using a fast majority vote, then demonstrate how modifications along axes such as the fault model, topology, node behavior, and require only local modifications to the model and proofs.
+With proof techniques like IC3 and $k$-induction with user-supplied
+invariants, model-checking scales further than ever before.  Still,
+fault-tolerant distributed systems are particularly challenging to
+model-check given their large state space and non-determinism. The
+typical approach to controlling complexity is to construct ad-hoc
+abstractions of faults, message-passing, and behaviors.  However,
+these abstractions come at the price of divorcing the model from its
+implementation and make refactoring difficult. In this work, we
+present a systematic model for fault-tolerant distributed system
+verification that combines calendar automata, synchronous observers,
+and abstract state machines specialized for model-checking. We
+demonstrate the efficacy of the framework by presenting the first
+model-checked \emph{implementation} of the Hybrid Oral Messages
+algorithm, using a fast majority vote, then demonstrate how
+modifications along axes such as the fault model, topology, and node
+behavior require only local modifications to the model and proofs.
 \end{abstract}

### 2016-11-29 source/experiments/SAL/OM/runproof.sh
Improve error handling in runproof.sh script; script now returns error codes if proofs fail and counts "invalid." as a failure

@@ -173,6 +173,7 @@ function cmd {
     for ((i=0; i < ${#lines[@]}; i++))
     do
         # insert SAL options
+        # XXX make 'sal-inf-bmc' generic here
         local salcmd=${lines[$i]/sal-inf-bmc/sal-inf-bmc $SALOPT}
         # replace SAL file name with the base name followed by any parameters
         if [ -n "${SAL_PARAMS}" ]; then
@@ -191,11 +192,13 @@ function cmd {
         local tmp=$(mktemp runproof.XXXXXXXX -ut)
         exec $salcmd_params 2>&1 |\
             tee $tmp |\
-            awk '/(failed|Error|induction rule failed)/ \
+            # XXX DRY with the regex, but needs to be quoted two different
+            # ways
+            awk '/(invalid|failed|Error|induction rule failed)/ \
                      {print "'$CS'" $0 "'$CE'"} \
                  /(proved|no counterexample)/ \
                      {print "'$CG'" $0 "'$CE'"}'
-        egrep -E -q '(failed|Error|induction rule failed)' $tmp
+        egrep -E -i -q '(invalid|failed|error|induction rule failed)' $tmp
         if [ $? -eq 0 ]; then
             fail=$(($fail+1))
         else
@@ -211,6 +214,12 @@ function cmd {
 
     local t=$(($end-$start))
     printf "total time ellapsed: $t seconds"
+
+    if [ $fail -eq 0 ]; then
+        return 0
+    else
+        return 1
+    fi
 }
 
 # Write out to standard out and err to standard error.

### 2016-11-29 source/experiments/SAL/OM/bench/time_om1.sh
Return meaningful exit statuses in jobs; use echo instead of printf to fd 3

@@ -38,10 +38,10 @@ diag_params=("1,1" "2,2" "3,3" "4,4" "5,5" "6,6" "7,7" "8,8" "9,9" "10,10")
 ###  EDIT ME  ##############################################
 
 # Specify which parameter set to use
-param_set="${mixed_params[@]}"
+param_set="${small_params[@]}"
 
 # Specify how many concurrent jobs to run
-CONCURRENT_LIMIT=${CONCURRENT_LIMIT:-8}
+CONCURRENT_LIMIT=${CONCURRENT_LIMIT:-1}
 export CONCURRENT_LIMIT  # exported for concurrent.lib.sh
 
 # specify timeout
@@ -57,12 +57,17 @@ source "$(cd "$(dirname "${BASH_SOURCE[0]}")" && pwd)/common.lib.sh"
 SAL_SMC=$(resolve_prog sal-smc)          # SAL symbolic model checker
 SAL_BMC=$(resolve_prog sal-inf-bmc)  # SAL inf-state bounded model checker
 
+CS="\033[0;31m"  # red
+CG="\033[0;32m"  # green
+CE="\033[0m"     # no color
+
 # run SAL symbolic model checker with specific options, redirect output to a
 # temporary file, match the "total execution time" line and extract run time
 sal_smc_cmd() {
     local model=$1
     local par=$2
     local tmp=$(mktemp sal_cmd.XXXXXX -ut)
+    local output=""
 
     timeout $TIME_LIMIT \
         ${SAL_SMC} -v 1 \
@@ -72,42 +77,51 @@ sal_smc_cmd() {
         --enable-dynamic-reorder \
         "${model}{$par}" vaa \
         2>&1 | tee "$tmp"
+    local code=${PIPESTATUS[0]}
 
     # output run time next to job status
-    if [ $? -eq 124 ]; then
+    if [ $code -eq 124 ]; then
         # timeout
         echo "RESULT $par, timeout"
         echo "timeout" >&3
+    elif [ $code -ne 0 ]; then
+        output="fail: exit status $code"
+        echo $output
+        echo $output >&3
     else
         # normal return
         cat "$tmp" | awk "/total exec/ {print \"RESULT $par,\", \$4}"
-        cat "$tmp" | awk "/total exec/ {print \"proved\", \$4}" >&3
+        cat "$tmp" | awk "/total exec/ {print \"proved in \", \$4, \"s\"}" >&3
     fi
     rm "$tmp"
+    return $code
 }
 
 sal_bmc_cmd() {
[…]

### 2016-11-29 doc/om-paper/paper.tex
Update benchmark numbers for OMH(1)
more paper stuff on proofs.

@@
 compositional approach to providing lemmas.
-note verification interesting in its own right }
+note verification interesting in its own right.
+verification goes beyond what has been done in thm provers }
 
@@
 
-%% parameterized by the
-%% number of rounds of message relay. Here, we focus on Oral Messages with one
-%% round of communication, $\OM{1}$, that tolerates a single Byzantine fault \todo{true?}, so
-%% long as there are a sufficient number of nodes nodes.
-
-%% Each of the $n$ nodes in the system represents a (Byzantine) General. In this
-%% formulation, the center node is the commanding general, while the $n-1$ outer
-%% generals are the lieutenants. The algorithm starts with the commanding general
-%% sending each lieutenant the same message $v$. Upon receipt of the commander's
-%% message, each lieutenant in turn sends the message received to each of the other
-%% lieutenants.  Once a lieutenant has received all $n-1$ expected messages, a
-%% majority vote is taken among the $n$ messages and the lieutenant declares its
-%% output (course of action) to be whichever message is in the majority.
-
-%% \begin{figure}[ht]
-%% \centering
-%% \begin{tikzpicture}[
-%% ell/.style={x radius=1, y radius=0.5, draw, shape=circle},
-%% arr/.style={semithick}]
-%% \node[ell] (c)  at (0,0)   {Commander};
-%% \node[ell] (l1) at (0:4)   {Lieutenant 1};
-%% \node[ell] (l2) at (120:4) {Lieutenant 2};
-%% \node[ell] (l3) [fill=gray!30] at (240:4) {Lieutenant 3};
-%% %
-%% \draw[->, arr] (c) -- node[auto] {$v$} (l1);
-%% \draw[->, arr] (c) -- node[auto] {$v$} (l2);
-%% \draw[->, arr] (c) -- node[auto] {$v$} (l3);
-%% \draw[<->, arr] (l1) -- node[auto,swap,near start] {$v$}
-%%                         node[auto,swap,near end] {$v$} (l2);
-%% \draw[<->, arr] (l1) -- node[auto,near start,color=red] {$x$}
-%%                         node[auto,near end] {$v$} (l3);
-%% \draw[<->, arr] (l2) -- node[auto,swap,near start,color=red] {$x$}
-%%                         node[auto,swap,near end] {$v$} (l3);
-%% \end{tikzpicture}
-%% \caption{\OM{1} Four generals, one traitor}
-%% \label{fig:om1}
-%% \end{figure}
-
 OMH is designed to preserve \emph{validity} and \emph{agreement}
@@
 
-\subsection{Invariants}
+\subsection{Invariants}\label{sec:invariants}
 
@@
 
-To make the proof scalable, we specify inductive invariants to be used by SAL's $k$-induction engine. The invariants are modular. There are 11 invariants, that fall into five categories:
+To make the proof scalable, we specify inductive invariants to be used by SAL's $k$-induction engine. The invariants are modular. There are 11 invariants, falling into five categories:
 \begin{enumerate}
@@
 One additional lemma notes that the messages currently received plus missing messages equals the total number of expected messages.
-  \item \emph{Faults}: Lemmas characterizing the effect of a fault in a single broadcast. For example, a lieutenant reports that it ``received an error'' when it receives a benignly-faulty message.
+  \item \emph{Faults}: Lemmas characterizing the effect of a fault in a single broadcast.
+Examples include lemmas stating that if a node receives a faulty message, some ``upstream'' node in the communication path was faulty.
+Another example is that the faults of messages latched by a node in its buffer match the faults ascribed to the sender in the calendar event.
   \item \emph{Voting}: Lemmas proving that the Fast Majority Vote algorithm implements a majority vote, if one exists.
@@
 For example, the ASM lemmas depend on both the calendar automata and lieutenant state-machine lemmas.
-
-\lee{make clear we're agnostic about human-provided invariants. We conjecture some can be auto-generated. Some might be inferred by IC3.}
-
- The two graphs show the structure to prove validity and agreement, respectively.
-The nodes group related lemmas, and a directed edge from node $a$ to $b$ means lemmas of $b$ depend on $a$.
-For readability, we have replicated lemmas shared in the proofs of validity and agreement.
-Additionally, we have elided the dependencies between the lemmas within each node.
-The proofs are all $k=1$ inductive.
-
-The calendar automata lemmas do not concern the system.
-The \emph{cal\_invt} lemma is a conjunction of basic properties about the calendar automata, essentially characterizing the axiomatization described in Section~\ref{sec:calendar}.
-Additionally, the \emph{missing\_buffer\_size} lemma relates the calendar to a node's message buffer: the messages received added to the empty messages equals the total number of possible messages to receive.
-
-The \emph{abstract\_all} lemmas characterizes the transition between states of the abstract state machine.
-Technically, \emph{abstract\_all} also depends on ``boilerplate'' lemmas relating each state of the abstract state machine to a state variable representing it, which are elided for space.
-
-The system itself includes both the lemmas characterizing the global protocol behavior as well as majority vote\lee{have we talked about the voting algorithm earlier?}.
-The global behavior concerns the entire distributed system, whereas the lemmas regarding the majority vote algorithm only concern the behavior of a single node.
-The proofs mechanize closely the proofs found in Boyer and Moore's original treatment~\cite{}.
-
-Finally, the fault model lemmas relate the fault model to the system.
-The \emph{fault\_propagation} lemma connects the presence of faulty messages in the calendar with the fault status of nodes that receive those messages in subsequent rounds.
-The \emph{symmetric\_relay} and \emph{manifest\_relay} lemmas characterize the effect of symmetric and manifest faults, respectively, on the kinds of messages a receiver obtains from a faulty node.
-All the fault model lemmas mentioned so far are necessary for the agreement proof, but for validity, we need only \emph{candidate\_good}, proving that if the maximum fault assumption holds and the sender is nonfaulty, then no modified messages are received.
+As can be seen, the proof structure is quite modular.
+The calendar lemmas are general and independent of any particular protocol or fault model.
+Similarly, lemmas about the internal behavior of a lieutenant is independent of the global protocol behavior.
+It is also independent of the effect of faults on the system---the only ``knowledge'' of faults that lieutenant has is whether a fault is benign or not.
+Lemmas about the behavior of faults in the system are also independent of the particular protocol being modeled.
+Likewise, lemmas about the particular voting algorithm used depend only on the lieutenant's internal behavior.
+Only the ASM depends on both calendar-specific and local state-machine results, since it is an abstraction of the entire system implementation.
+Recall, however, that the ASM is a convenience for debugging and can be elided.
 
@@
       \hline \hline
-      1   & 13  & 18  & 23  & 29  & 35  & 43  & 54  & 70  & 90  & 129 \\
+      1   & 7    & 9   & 12   & 15  & 21   & 25   & 32   & 40  & 54  & 74  \\
       \hline
-      2   & 17  & 26  & 38  & 52  & 68  & 89  & 115 & 159 & 227 & -   \\
+      2   & 17   & 14  & 21   & 30  & 42   & 53   & 74   & 99  & 144 & -   \\
       \hline
-      3   & 21  & 40  & 61  & 85  & 120 & 170 & 246 & 410 & -   & -   \\
+      3   & 21   & 22  & 40   & 50  & 81   & 102  & 155  & 279 & -   & -   \\
       \hline
-      4   & 27  & 58  & 98  & 195 & 213 & 628 & 855 & -   & -   & -   \\
+      4   & 27   & 34  & 59   & 99  & 141  & 237  & 1114 & -   & -   & -   \\
       \hline
-      5   & 36  & 93  & 292 & 389 & 1940& TO  & -   & -   & -   & -   \\
+      5   & 22   & 94  & 125  & 335 & TO   & 1406 & -    & -   & -   & -   \\
       \hline
-      6   & 56  & 303 & 976 & TO  & TO  & -   & -   & -   & -   & -   \\
+      6   & 36   & 132 & 2966 & 844 & 2457 & -    & -    & -   & -   & -   \\
       \hline
-      7   & 114 & 1993& TO  & TO  & -   & -   & -   & -   & -   & -   \\
+      7   & 83   & 487 & TO   & TO  & -    & -    & -    & -   & -   & -   \\
       \hline
-      8   & 338 & TO  & TO  & -   & -   & -   & -   & -   & -   & -   \\
+      8   & 298  & TO  & TO   & -   & -    & -    & -    & -   & -   & -   \\
       \hline
-      9   & 895 & TO  & -   & -   & -   & -   & -   & -   & -   & -   \\
+      9   & 1428 & TO  & -    & -   & -    & -    & -    & -   & -   & -   \\
       \hline
@@
 Compare to rushby's?}
+\lee{what are the machine specs for running the benchmark?}
 
-\subsection{Modular Verification}
+\subsection{Modular Verification}\label{sec:modular}
+To demonstrate the modularity of the model and verification approach, in this section, we explore variants to the model and discover the effort required to implement the modifications and repair the proofs.
+
+%% \begin{figure}
+%%   \centering
+%%   \begin{tabular}{r|l|l|l|}
+%%                    & Modules modified    & Lemmas modified & Lemma classes modified\\
+%%   \hline \hline
+%%   Faults           & 
+%%   \hline
+%%   Time-triggered   &
+%%   \hline
+%%   \end{tabular}
+%%   \caption{Refactoring effort for protocol modifications.}
+%%   \label{fig:effort}
+%% \end{figure}
+
+\paragraph{Modified fault model.}
+
+\paragraph{Time-triggered messaging.}
+
+\subsection{Proof Effort Remarks}
+The lemmas described in Section~\ref{sec:invariants} are constructed by-hand and represent multiple days of effort, but that effort includes both model and protocol construction and generalization as well as verification.
+The counterexamples returned by SAL are very useful for strengthening invariants, but tedious to analyze---a model with five relays and two receivers contains 90 state variables, and there are known counterexamples to models that size~\cite{csl-93-2}.
+Once we developed the ASM synchronously-composed observer, the verification effort was sped up considerably.
+
+As shown in Section~\ref{sec:modular}, modifications to the implementation are modular.
+In particular, any protocol specified in the model can reuse the lemmas associated with the calendar automata and fault-model (insofar as they use the same fault model).
+
+Moreover, we are agnostic about how lemmas are discovered.
+As techniques like IC3 scale, they may be discovered automatically.
+$k$-induction in infinite-state model-checking blurs the lines between interactive and automated theorem proving.
+IC3 can even be strengthened using $k$-induciton~\ref{pdr-kind}.
+Tools like Ivy can present graphical counterexamples to help interactively develop invariants~\cite{ivy}.
 
@@
 
-\url{https://people.eecs.berkeley.edu/~apanda/assets/papers/pldi16.pdf}
-
 \lee{mention that ASMs similar to disjunction invariants}
@@
 
+\lee{link to all results in github}
 \lee{talk about compositional verification of properties, even though SAL doesn't directly support it}
@@
 \lee{talk about lack of axiomatization in model-checking (multiple rushby bugs), but  tradeoff of deadlock. see proglema paper}
-
 \lee{talk about atom-sally}
-
 \lee{note some details still elided for implementation ---e.g., CRCs, etc.}

### 2016-11-30 doc/om-paper/paper.tex
Answer questions at end of scalability subsecion

@@
 
-\lee{let's also check out scalability when we ``turn off'' faults.
-Also, how hard is it to turn them off or change the fault model?
-E.g., how many lines of spec need to be changed?
-Compare to rushby's?}
-\lee{what are the machine specs for running the benchmark?}
+The benchmarks were performed on a server with Intel Xeon E312xx (Sandy
+Bridge) CPUs (mid 2011).
+
+For comparison, when the fault model is ``turned off'' by defining the fault
+type to contain a single 'nonfaulty' value and removing fault specific lemmas,
+model-checking scales further. For example when there are 3 relays and 3
+receivers the time is 25s (compared to 40s above), and when where are 5 relays
+and 5 receivers the time is 315s (compared to timeout above). See Figure
+\ref{fig:effort} for more comparisons.
 

### 2016-11-30 doc/om-paper/paper.tex
added table for extensions.

@@
 \subsection{Modular Verification}\label{sec:modular}
+
+\begin{figure}
+  \centering
+  \begin{tabular}{p{3cm}|p{3cm}|l|l|}
+                                  & Modules modified    & Lemmas modified & Lemma classes modified\\
+  \hline \hline
+  Byzantine-only fault model      & 0                   & 0               & 0 \\
+  \hline
+  Time-triggered message-passing  & \texttt{clock}, \texttt{relay}
+                                  , \texttt{receiver}   & 0               & 0 \\
+  \hline
+  \end{tabular}
+  \caption{Refactoring effort for protocol modifications.}
+  \label{fig:effort}
+\end{figure}
 To demonstrate the modularity of the model and verification approach, in this section, we explore variants to the model and discover the effort required to implement the modifications and repair the proofs.
+The results are summarized in Figure~\ref{sec:effort} and sketched below.
 
-%% \begin{figure}
-%%   \centering
-%%   \begin{tabular}{r|l|l|l|}
-%%                    & Modules modified    & Lemmas modified & Lemma classes modified\\
-%%   \hline \hline
-%%   Faults           & 
-%%   \hline
-%%   Time-triggered   &
-%%   \hline
-%%   \end{tabular}
-%%   \caption{Refactoring effort for protocol modifications.}
-%%   \label{fig:effort}
-%% \end{figure}
-
-\paragraph{Modified fault model.}
+\paragraph{Fault model.}
+Modifying the fault model is easy. In involves modifying the definition of a single function, \texttt{mfa}, that takes as an argument the static assignment of node identifiers to faults. Moreover
 

### 2016-11-30 doc/om-paper/paper.tex
pervasive paper edits.

@@
 \begin{abstract}
-With proof techniques like IC3 and $k$-induction with user-supplied
-invariants, model-checking scales further than ever before.  Still,
-fault-tolerant distributed systems are particularly challenging to
-model-check given their large state space and non-determinism. The
-typical approach to controlling complexity is to construct ad-hoc
-abstractions of faults, message-passing, and behaviors.  However,
-these abstractions come at the price of divorcing the model from its
-implementation and make refactoring difficult. In this work, we
-present a systematic model for fault-tolerant distributed system
-verification that combines calendar automata, synchronous observers,
-and abstract state machines specialized for model-checking. We
-demonstrate the efficacy of the framework by presenting the first
-model-checked \emph{implementation} of the Hybrid Oral Messages
-algorithm, using a fast majority vote, then demonstrate how
-modifications along axes such as the fault model, topology, and node
-behavior require only local modifications to the model and proofs.
+With proof techniques like IC3 and $k$-induction, model-checking scales further
+than ever before.  Still, fault-tolerant distributed systems are particularly
+challenging to model-check given their large state space and
+non-determinism. The typical approach to controlling complexity is to construct
+ad-hoc abstractions of faults, message-passing, and behaviors.  However, these
+abstractions come at the price of divorcing the model from its implementation
+and make refactoring difficult. In this work, we present a systematic model for
+fault-tolerant distributed system verification that combines calendar automata,
+synchronous observers, and abstract state machines specialized for
+model-checking. We demonstrate the efficacy of the framework by presenting the
+first model-checked \emph{implementation} of the Hybrid Oral Messages algorithm,
+using a fast majority vote, then demonstrate how modifications along axes such
+as the fault model, timing constraints, and local node behavior require only local modifications
+to the model and proofs. This work is carried out in SAL.
 \end{abstract}
@@
 For example, the behavior of nodes is naturally specified as a transition system in which transitions are guarded by the node's fault state.
-But faults are part of the environment; in an implementation, does not change behavior based on its faults.
+But faults are part of the environment; an implementation does not use its own fault status to choose actions!
 Another class of abstractions is used to simplify models.
-For example, message passing might be abstracted with shared state, or a node's local behavior---e.g., a majority vote---is elided and instead, the output is constrained by a \emph{specification} of a majority vote.
+For example, message passing might be abstracted with shared state, or a node's local behavior is elided and instead, the output is constrained by a \emph{specification} of the behavior.
 
@@
 We describe a generic formal model for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}.
-There are three aspects of the model: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) \emph{synchronous observers}, originally from the synchronous languages community~\cite{} and popularized by Rushby for model-checking; and (3) \emph{abstract state machines}, useful for abstracting counterexamples.
+There are three aspects of the model: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) \emph{synchronous observers}, originally from the synchronous languages community~\cite{} and popularized by Rushby for model-checking~\cite{}; and (3) \emph{abstract state machines}, useful for abstracting counterexamples.
 The model is suitable for infinite-state model-checking.
-With proof techniques like IC3 and $k$-induction, model-checking scales further than ever before~\cite{pdr-kind}.
 
@@
 The verification is interesting in its own right, as it is the first fully parametric (on the number of nodes) model-checked \emph{implementation} of the algorithm.
-In Section~\ref{sec:experimental}, we give performance metrics for the model, and we modify the implementation by changing the fault model, topology, node behavior, and show that only local modifications are required to the model and proofs.
+In Section~\ref{sec:experimental}, we give performance metrics for the model.
+First, we benchmark the model, showing that despite the additional complexity, it scales farther than other model-checking verifications at the algorithmic level.
+The scalability is due principally to specifying invariants and using $k$-induction in SRI's SAL model-checker~\cite{}.
+Developing invariants requires some user guidance.
+A user should not have to discover all new invariants for new protocols or implementations.
+We show that most of the lemmas required are not protocol specific and are modular, concerning only specific aspects of the system.
+To demonstrate this, we modify the OHM implementation in various ways to show that only local modifications are required to the model and proofs.
+
+Finally, in Section~\ref{sec:related} we describe related work, and we make concluding remarks in Section~\ref{sec:conclusions}.
 
@@
 
-\lee{finish intro...
-One more contribution is that we abstract faults using partially interpreted functions in model-checking, which is new?
-Note focus is on infinite-state model-checking: will need user-provided lemmas.
-compositional approach to providing lemmas.
-note verification interesting in its own right.
-verification goes beyond what has been done in thm provers }
-
  %% Timed automata are an alternative to the \emph{timed automata} formalism~\cite{}, as implemented in model-checkers like UPPAAL~\cite{} and Kronos~\cite{}.
@@
 Timed automata assume the existence of continuously-varying real-valued variables, known as \emph{real-time clocks} in a solver.
-Timed automata model-checkers are specialized for verifying real-time systems in which the only infinite-valued variables are the real-time clocks.
-Modern general-purpose model-checkers like \lee{name some} use SMT solvers for general infinite-state system verification that is not limited to real-time clocks.
+Model-checkers for timed-automata are specialized for verifying real-time systems in which the only infinite-valued variables are the real-time clocks.
 
@@
 While synchronous observers have been part of model-checking folklore for some time, Rushby systematically describes their application to the domain~\cite{}.
-The general form for specifying a system using synchronous observers is the following:
-$$\mathcal{S} | \mathcal{E} | \mathcal{A} | \mathcal{R}$$
+In Section~\ref{sec:sync-model} we define synchronous observers and then describe their application to modeling faults in Section~\ref{sec:kibitzer}.
+
+\subsubsection{Model}
+A synchronous observer is of the form
+$$\mathcal{S} | \mathcal{O}$$
 \noindent
 where `$|$' denotes synchronous composition of transition systems (defined formally below).
-$\mathcal{S}$ is the \emph{system}, or machine to be realized in an implementation.
-$\mathcal{E}$ is the \emph{environment} in which the system operates.
-For example, the environment can be used to specify the behavior of faults in the system.
-$\mathcal{A}$ are the \emph{assumptions} on the system, and $\mathcal{R}$ are the \emph{requirements}.
-Synchronous observers can be thought of as a generalization of state predicates,like Hoare triples~\cite{}, to transitions systems.
-
+$\mathcal{S}$ is the \emph{system}, and $\mathcal{O}$ is the \emph{observer}.
 We define synchronous composition more precisely below, beginning with some preliminary definitions:
@@
 be transitions systems.
-
-\begin{definition}[Deadlock]
-$M$ is deadlocked iff either the set of initial states $I$ is empty, or there exists a state $s_0 \in S$ and state $s_1 \in I$ such that $s_0$ is reachable from $s_1$ and there exists no $s_2 \in S$ such that $s_0 \rightarrow s_2$.
-\end{definition}
 
@@
 
-\begin{definition}[Transition system refinement]
-$M \subseteq M'$ if and only if $S \subseteq S'$, $I \subseteq I'$, and $\rightarrow \subseteq \rightarrow'$.
-\end{definition}
+%% \begin{definition}[Transition system refinement]
+%% $M \subseteq M'$ if and only if $S \subseteq S'$, $I \subseteq I'$, and $\rightarrow \subseteq \rightarrow'$.
+%% \end{definition}
 
-By convention, the assumptions $\mathcal{A}$ should be a refinement of the system $\mathcal{S}$ and environment $\mathcal{E}$.
+%% By convention, the assumptions $\mathcal{A}$ should be a refinement of the system $\mathcal{S}$ and environment $\mathcal{E}$.
 The system is \emph{realizable}~\cite{} under its assumptions if $\mathcal{S} | \mathcal{E} | \mathcal{A}$ is not deadlocked.
 A system is \emph{correct} if $\mathcal{S} | \mathcal{E} | \mathcal{A}$ refines the requirements $\mathcal{R}$.
+
+%% \begin{definition}[Deadlock]
+%% $M$ is deadlocked iff either the set of initial states $I$ is empty, or there exists a state $s_0 \in S$ and state $s_1 \in I$ such that $s_0$ is reachable from $s_1$ and there exists no $s_2 \in S$ such that $s_0 \rightarrow s_2$.
+%% \end{definition}
 
@@
 Second, as described by Rushby~\cite{}, requirements can be specified using transition systems rather than temporal logic formula, which is often more natural.
+We use synchronous observers for both purposes.
 
-\subsection{The Synchronous Kibitzer}\label{sec:kibitzer}
-\lee{We use synchronous composition to implement fault injection. Mention in the intro. This is fundamental to the model, so we have section devoted to it.}
-In Section~\ref{sec:sync}, we mentioned an environment transition system synchronously composed with the requirements, system, and assumptions.
-A particularly important part of the environment is the model of faults, which we address here.
-The typical approach to injecting faults into a formal model of a distributed system is to add new state variables (or increase the domain of existing state variables) to each process representing its fault state.
+\subsubsection{The Synchronous Kibitzer}\label{sec:kibitzer}
+Synchronous observers can be used to model the environment, and in particular, to compositionally inject faults.
+The typical approach to injecting faults into a formal model of a distributed system is to add new state variables to each process representing its fault state.
 Then additional state variables are added to represent the nondeterminism introduced by faults into the state or messages by a process.
@@
 
-Not only is modeling faults expensive, but it is generally non-compositional insofar as the fault model pervades the entire model; for example, consider the model of Oral Messages presented by Rushby~\cite{}.
+Not only is modeling faults expensive, but it is generally non-compositional insofar as the fault model pervades the entire model.
 
@@
 
-To address these problems, we introduce a synchronous observer that injects faults that we call a \emph{synchronous kibitzer}.
+To address these issues, we introduce a synchronous observer that injects faults that we call a \emph{synchronous kibitzer}.
 The kibitzer decomposes the state and transitions associated with the fault model from the system.
@@
 
-In specifying a fault-tolerant system, we must also specify a \emph{maximum fault assumption} (MFA), which specifies the maximum number of of each kind of fault that may be present in the system for some system property to hold.
-Define a enumeration of fault types $$\textnormal{FAULT\_TYPE} = \{none, byz, sym, man\}$$ As in the previous section, let $Id \subset \mathbb{N}$ be a finite set of indices, and let the variable $$faults: Id \rightarrow \textnormal{FAULT\_TYPE}$$ range over possible mappings from process indices to faults.
+%% In specifying a fault-tolerant system, we must also specify a \emph{maximum fault assumption} (MFA), which specifies the maximum number of of each kind of fault that may be present in the system for some system property to hold.
+
+Define an enumeration of fault types $$\textnormal{FAULT\_TYPE} = \{none, byz, sym, man\}$$ As in the previous section, let $Id \subset \mathbb{N}$ be a finite set of indices, and let the variable $$faults: Id \rightarrow \textnormal{FAULT\_TYPE}$$ range over possible mappings from process indices to faults.
 
@@
 
-\lee{Talk about how to define an MFA in the environment}
+The upshot is that the fault model and fault-injection is completely modular to the specification of the protocol and node behavior.
 
@@
 A state is a collection of variable values, and in large models, there are a lot.
-A single state includes assignments to each of the variables, and counterexample trace may be many 10s of states in length.
-In practice, counterexamples depend on a small number of state variables.
+A single state includes assignments to each of the variables, and the counterexample traces we dealt with (see Section~\ref{sec:experimental}) can contain hundreds of state variables.
+In practice, counterexamples depend on a small subset of these.
 The user has to manually separate the state variables that matter from those that do not for a specific property.
@@
 
-Systems are usually designed in modes such that predicates on state variables are parameterized on which mode the system is in.
+%% Systems are usually designed in modes such that predicates on state variables are parameterized on which mode the system is in.
 The transition between modes implements a finite state-machine.
@@
 
-Our solution to the problem of constructing a mode-based state machine uses abstract state machines (ASMs).
+But the relationship between many of the state variables have an intuitive structure that can be characterized by a state machine imposed over it.
+The state machine describes discrete modes or configurations.
+Each mode captures a relation on a subset of the state variables, as well as the set of possible modes that follow.
+
+The idea is an implementation of abstract state machines (ASMs).
 ASMs are a generalization of finite state machines originally developed by Gurevich to generalize state machines to operate over arbitrary data structures~\cite{asm}.
-
 In our context, we require only a restricted form of ASMs known as a \emph{sequential small-step ASM}:
@@
 
-\lee{note that the ASM is not a refinement.}
+\lee{probably need an example here.}
 
@@
 
-To demonstrate the model and techniques, we model the Oral Messages Hybrid (OMH) algorithm~\cite{}, a more sophisticated version of the class Oral Messages (OM) algorithm.
+To demonstrate the model and techniques, we model an implementation of the Oral Messages Hybrid (OMH) algorithm~\cite{}, a more sophisticated version of the class Oral Messages (OM) algorithm.
 OM was first formulated by Lamport, Shostak, and Pease~\cite{om} to achieve agreement in a distributed system while tolerating Byzantine faults.
 OMH was originally developed by Thambidurai and Park~\cite{hybrid} to achieve the same result in the presence of a more sophisticated fault model that includes symmetric and benign faults.
-However, OMH had a bug, as originally formulated, which was corrected and formally verified by Lincoln and Rushby~\cite{csl-93-2}.
+However, OMH had a bug, as originally formulated, which was corrected and formally verified by Lincoln and Rushby using interactive theorem-proving~\cite{csl-93-2}.
 
@@
 
-To make the proof scalable, we specify inductive invariants to be used by SAL's $k$-induction engine. The invariants are modular. There are 11 invariants, falling into five categories:
+To make the proof scalable, we specify inductive invariants to be used by SAL's $k$-induction engine. There are 11 invariants, falling into five categories:
 \begin{enumerate}
@@
 
+\lee{discuss results}
+
 \subsection{Modular Verification}\label{sec:modular}
@@
 That said, their model of OM(1) (not OMH(1)), timeouts after 10~hours, and we assume a voting algorithm is not explicitly specified.
-Very recently, Jovanovi{\'{c}} and Dutertre use a ``flattened'' high-level model of OM(1) as a benchmark for IC3 augmented with $k$-induction~\cite{pdr-kind}.
+\lee{for what size?
+Also, Rushby's model?} Very recently, Jovanovi{\'{c}} and Dutertre use a ``flattened'' high-level model of OM(1) as a benchmark for IC3 augmented with $k$-induction~\cite{pdr-kind}.
 

### 2016-12-01 doc/om-paper/paper.tex
started a conclusion.

@@
 \section{Conclusions}\label{sec:conclusions}
+This work fits within a larger project, in collaboration with Honeywell Labs, to build an \emph{architectural domain-specific language} (ADSL) for specifying and verifying distributed fault-tolerant systems.
+The goal of the language is to synthesize both software and/or hardware implementations as well as formal models for verification.
+Before building such an ADSL, we needed a scalable general formal model to which to compile, leading to the work presented.
+An ADSL will make refactoring even easier.
+We have developed a preliminary ADSL that generates C code as well as formal models SRI's Sally~\cite{}, a model-checking descendent to SAL that will be described in a subsequent paper.\footnote{\url{https://github.com/GaloisInc/atom-sally}}
 
@@
 \lee{talk about lack of axiomatization in model-checking (multiple rushby bugs), but  tradeoff of deadlock. see proglema paper}
-\lee{talk about atom-sally}
 \lee{note some details still elided for implementation ---e.g., CRCs, etc.}
@@
 \section*{Acknowledgments}
-\lee{NASA contract, honeywell ppl, rushby's model?}
+This work is partially supported by NASA contract \#NNL14AA08C.
+We are indebted to our collaborators Brendan Hall and Srivatsan Varadarajan at Honeywell Labs, and to Wilfredo Torres-Pomales at NASA Langley for their discussions and insights.
 

### 2016-12-02 doc/om-paper/paper.tex
Updates to section 1, including several filled in citations

@@
+% vim: wrap linebreak nolist textwidth=0 wrapmargin=0
 %\documentclass{IEEEtran/IEEEtran}
@@
 \title{Modular Model-Checking for Fault-Tolerant Distributed System Implementations}
-\author{Ben and Lee}
+\author{Benjamin F Jones and Lee Pike}
+\institute{Galois, Inc.}
 
@@
 System designers create software and hardware implementations to test, simulate, and deploy.
-Discrepancies between implementations and algorithmic models can arise if the latter is abstracted too much from the former, particularly if those abstractions are ad-hoc and system specific.
-Furthermore, as an implementations are tweaked and modified to explore the design space, it is easy for the formal models to become inconsistent.
+Discrepancies between implementations and algorithmic models can arise if the latter is abstracted too much from the former, particularly if those abstractions are ad-hoc and system specific \ben{for example?}.
+Furthermore, as implementations are modified to explore the design space, it is easy for the formal model and the implementation to become inconsistent.
 
@@
 One is to intertwine the environmental model with the system description.
-For example, the behavior of nodes is naturally specified as a transition system in which transitions are guarded by the node's fault state.
+For example, the behaviors of nodes are naturally specified as a transition system in which transitions are guarded by the node's fault state.
 But faults are part of the environment; an implementation does not use its own fault status to choose actions!
 Another class of abstractions is used to simplify models.
-For example, message passing might be abstracted with shared state, or a node's local behavior is elided and instead, the output is constrained by a \emph{specification} of the behavior.
+For example, message passing might be abstracted with shared state, or a node's local behavior is elided and instead, the output is constrained by a \emph{specification} of the behavior, e.g.
+pre- and post-conditions.
 
@@
 We describe a generic formal model for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}.
-There are three aspects of the model: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) \emph{synchronous observers}, originally from the synchronous languages community~\cite{} and popularized by Rushby for model-checking~\cite{}; and (3) \emph{abstract state machines}, useful for abstracting counterexamples.
+There are three aspects of the model: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) \emph{synchronous observers}, originally from the synchronous languages community~\cite{Halbwachs-Lagnier:92,Halbwachs-Lagnier:94} and popularized by Rushby for model-checking~\cite{Rushby:SAS14}; and (3) \emph{abstract state machines}, useful for abstracting counterexamples.
 The model is suitable for infinite-state model-checking.
 
-To demonstrate the model in Section~\ref{sec:byz}, we present a formally-verified implementation of the Oral Messages Hybrid algorithm (OMH)~\cite{}, implementing the Boyer-Moore Fast Majority Vote algorithm for algorithm~\cite{}.
+To demonstrate the modeling technique, in Section~\ref{sec:byz} we present a formally-verified implementation of the Oral Messages Hybrid algorithm (OMH)~\cite{Lincoln-Rushby}, implementing the Boyer-Moore Fast Majority Vote algorithm for algorithm~\cite{Boyer-Moore:91}.
 The verification is interesting in its own right, as it is the first fully parametric (on the number of nodes) model-checked \emph{implementation} of the algorithm.
@@
 First, we benchmark the model, showing that despite the additional complexity, it scales farther than other model-checking verifications at the algorithmic level.
-The scalability is due principally to specifying invariants and using $k$-induction in SRI's SAL model-checker~\cite{}.
+The scalability is due principally to specifying invariants and using $k$-induction in SRI's SAL model-checker~\cite{sal}.
 Developing invariants requires some user guidance.
@@
 We envision a DSL that generates both code for simulation, testing, and deployment, as well as formal models for verification.
-That way, the formal models verified and code executed stay in synch, and abstractions are systematically made.
+That way, the formal models verified and code executed stay in sync, and abstractions are systematically made.
 We say more about the DSL in Section~\ref{sec:conclusions}.
@@
 \emph{Timed automata} are specialized formalisms for verifying real-time systems~\cite{}.
-Timed automata assume the existence of continuously-varying real-valued variables, known as \emph{real-time clocks} in a solver.
+Timed automata assume the existence of continuously-varying real-valued variables, known as \emph{real-time clocks}.
 Model-checkers for timed-automata are specialized for verifying real-time systems in which the only infinite-valued variables are the real-time clocks.
@@
 
-Calendar automata avoid the problems of encoding timed automata.
+The calendar automaton avoids the problems of encoding timed automata.
 Rather than encoding ``how much time has passed since the last event'', it encodes ``how far into the future is the next scheduled event'', and a real-valued variable representing the current time is updated to the next event time.
@@
 
-\item \label{cal:d} In all states, if the current time equals a timeout, then the only transition enabled is a calendar event update associated with the timeout: $\forall \sigma \in S$, $\exists (e_i, t_i) \in \sigma(cal)$ such that $\sigma(now) = t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma'(now) = \sigma(now)$, $\sigma'(c_j) = \sigma(c_j)$ for all $c_j \in \sigma(c)$ such that $c_j \neq c_i$, and $\sigma'(t_i) > \sigma(t_i)$.
+\item \label{cal:d} In all states, if the current time equals a timeout, then the only transitions enabled are calendar event updates associated with the timeout: $\forall \sigma \in S$, $\exists (e_i, t_i) \in \sigma(cal)$ such that $\sigma(now) = t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma'(now) = \sigma(now)$, $\sigma'(c_j) = \sigma(c_j)$ for all $c_j \in \sigma(c)$ such that $c_j \neq c_i$, and $\sigma'(t_i) > \sigma(t_i)$.
 \end{enumerate}

### 2016-12-02 doc/om-paper/paper.tex
Updates to section 2.1 - 2.2 plus more filled in refs

@@
 \newcommand{\OM}[1]{\ensuremath{\mathrm{OM}(#1)}}
+\newcommand{\Id}{\ensuremath{\mathrm{Id}}}
+\newcommand{\Msg}{\ensuremath{\mathrm{Msg}}}
 
@@
 \end{enumerate}
+\ben{Do we also require that $c_i \notin \sigma'(c)$ ?}
 
@@
 \end{lemma}
-\begin{proof}
-By induction on states.
+% \begin{proof}
+% By induction on states.
 In an initial state, the property holds immediately (Law~\ref{cal:a}).
@@
 Otherwise, if $\sigma(now) = t_i$, for some $t_i$, then the only possible transition is a calendar event update, incrementing $t_i$ but not $\sigma(now)$ (Law~\ref{cal:d}).
-\end{proof}
+% \end{proof}
 
@@
 \end{lemma}
-\begin{proof}
-From Lemma~\ref{lem:ft}, in $\sigma(now) \leq t_i$, for all $(e_i, t_i) \in \sigma(cal)$.
+% \begin{proof}
+% From Lemma~\ref{lem:ft}, in $\sigma(now) \leq t_i$, for all $(e_i, t_i) \in \sigma(cal)$.
 For any $\sigma'$ such that $\sigma \rightarrow \sigma'$, either there is a time progress update (Law~\ref{cal:c}) or a calendar event update (Law~\ref{cal:d}).
 In the first case, $\sigma(now) < \sigma'(now)$, and in the second case, $\sigma(now) = \sigma'(now)$.
-\end{proof}
+% \end{proof}
+
+Proofs of these two lemmas are straightforward so we omit them here.
 
@@
 
-Again following Dutetre and Sorea, we associate calendar events with channels in a distributed system~\cite{dutetre}.
+Again, following Dutetre and Sorea, we associate calendar events with channels in a distributed system~\cite{dutetre}.
 Specializing calendars to message passing does not lose generality since all external communication from an individual process can be abstracted as message passing.
@@
 
-Assume processes are indexed from a finite set $Id \subset \mathbb{N}$.
+Assume processes are indexed from a finite set $\Id \subset \mathbb{N}$.
 A \emph{channel} from process $i$ to $j$ is an ordered pair $(i,j)$.
-Define a set of messages $Msg$.
+Define a set of messages $\Msg$.
 Given a channel and a timeout, let $send$ be a relation on messages sent on a channel at a given time:
-$$send \subseteq Id \times Id \times \mathbb{R} \times Msg$$
+$$send \subseteq \Id \times \Id \times \mathbb{R} \times \Msg$$
 So $send(i, j, t, m)$ holds iff $i$ sends to $j$ message $m$ at time $t$. Likewise, let
-$$recv \subseteq Id \times Id \times \mathbb{R} \times Msg$$
+$$recv \subseteq \Id \times \Id \times \mathbb{R} \times \Msg$$
 be a relation on messages received on a channel at a time, so that $recv(i, j, t, m)$ holds iff the message $m$ received by $j$ from $i$ at time $t$.
 
-In the non-faulty case, we require that messages received were previously sent and not previously received: if $recv(i, j, t, m)$, then $\exists t'$ such that $send(i, j, t', m)$ where $t' < t$, and $\neg\exists t''$ such that $t' < t'' < t$ and $recv(i, j, t'', m) = recv(i, j, t, m)$.
+In the absence of faults, we require that messages received were previously sent and not previously received: if $recv(i, j, t, m)$, then $\exists t'$ such that $send(i, j, t', m)$ where $t' < t$, and $\neg\exists t''$ such that $t' < t'' < t$ and $recv(i, j, t'', m) = recv(i, j, t, m)$.
 (We address faults in Section~\ref{sec:kibitzer}.)
@@
 \subsection{Synchronous Observers}\label{sec:sync}
-Synchronous observers are state machines that are synchronously composed with a system specification to check properties over the system's state variables.
-Synchronous observers are a method for composing specifications with a program.
-Synchronous observers were first introduced in the context of the synchronous programming languages, such as Lustre~\cite{}.
-While synchronous observers have been part of model-checking folklore for some time, Rushby systematically describes their application to the domain~\cite{}.
+Synchronous observers are state machines that are synchronously composed with a system specification to check properties over the system's state variables, i.e.
+they are a method for composing specifications with a program.
+Synchronous observers were first introduced in the context of the synchronous programming languages, such as Lustre~\cite{lustre:ieee}.
+While synchronous observers have been part of model-checking folklore for some time, Rushby systematically describes their application to the present domain in~\cite{Rushby:SAS14}.
 In Section~\ref{sec:sync-model} we define synchronous observers and then describe their application to modeling faults in Section~\ref{sec:kibitzer}.
@@
 \begin{definition}[Reachable states]
-For a transition relation $\rightarrow$ over the states of $S$, for states $s_0, s_1, \ldots, s_k \in S$, $s_k$ is reachable from $s_0$ iff
-$$s_0 \rightarrow s_1 \rightarrow \ldots \rightarrow s_k$$
+For a transition relation $\rightarrow$ over the states of $S$, for states $s_0,  s \in S$, we say $s$ is reachable from $s_0$ iff there exist states $s_1, \ldots, s_k \in S$ such that
+$$s_0 \rightarrow s_1 \rightarrow \ldots \rightarrow s_k \rightarrow s$$
 \end{definition}
 \noindent
-In particular, a state $s$ is reachable from $s$.
+By convention, every state $s$ is reachable from itself.
 
@@
 
-Define an enumeration of fault types $$\textnormal{FAULT\_TYPE} = \{none, byz, sym, man\}$$ As in the previous section, let $Id \subset \mathbb{N}$ be a finite set of indices, and let the variable $$faults: Id \rightarrow \textnormal{FAULT\_TYPE}$$ range over possible mappings from process indices to faults.
+Define an enumeration of fault types $$\textnormal{FAULT\_TYPE} = \{none, byz, sym, man\}$$ As in the previous section, let $\Id \subset \mathbb{N}$ be a finite set of indices, and let the variable $$faults: \Id \rightarrow \textnormal{FAULT\_TYPE}$$ range over possible mappings from process indices to faults.
 
@@
 % Define $rnd: \mathbb{R} \rightarrow \mathbb{N}$ such that if $rnd(t_0) < rnd(t_1)$, then $t_0 < t_1$.
-A $broadcast: Id \rightarrow 2^{Id} \rightarrow \mathbb{R} \rightarrow Msg \rightarrow 2^{E}$ takes a sender, a set of receivers, a real-time, and a message to send each receiver, and returns a set of calendar events:
+A $broadcast: \Id \rightarrow 2^{\Id} \rightarrow \mathbb{R} \rightarrow \Msg \rightarrow 2^{E}$ takes a sender, a set of receivers, a real-time, and a message to send each receiver, and returns a set of calendar events:
 $$broadcast(i, R, t, m) = \{m | j \in R \textnormal{ and } send(i, j, t) = m\}$$
@@
 &nonfaulty\_constraint =\\
-  &\quad \forall i, j \in Id, t \in \mathbb{R}.\\
+  &\quad \forall i, j \in \Id, t \in \mathbb{R}.\\
   &\quad\quad faults(i) = none\\
@@
 &sym\_constraint =\\
-  &\quad \forall i, j, k \in Id, t \in \mathbb{R}.\\
+  &\quad \forall i, j, k \in \Id, t \in \mathbb{R}.\\
   &\quad\quad (faults(i) = sym \textnormal{ and } broadcast(i, \{j, k\}, t, m))\\

### 2016-12-02 doc/om-paper/paper.tex
Updates to section 4, including new example for ASM subsection

@@
 \title{Modular Model-Checking for Fault-Tolerant Distributed System Implementations}
-\author{Benjamin F Jones and Lee Pike}
-\institute{Galois, Inc.}
+
+%\author{Benjamin F Jones and Lee Pike}
+\author{Benjamin F Jones\inst{1} \and Lee Pike\inst{2}}
+
+\institute{Galois, Inc., Portland OR 97204,\\
+\email{bjones@galois.com}
+\and
+Galois, Inc., Portland OR 97204,\\
+\email{leepike@galois.com}}
 
@@
 %% Two examples of model-specific abstractions include the following: (1) Suppose we model a system in which a process that broadcasts the values $\{0, 1\}$ in the nominal case, but when the process suffers a fault, it broadcasts no value.
-To model the system, it is typical to define an enumerated type of broadcast values as $$\{0, 1, NONE\}$$ However, the enumeration conflates the environment (i.e., faults) and system (i.e., the non-faulty broadcast values).
+To model the system, it is typical to define an enumerated type of broadcast values as $$\{0, 1, NONE\}$$ However, the enumeration conflates the environment (i.e., faults) and system (i.e., the nonfaulty broadcast values).
 (2) Another example is to model message passing as shared state between processes to reduce the number of state variables in the model, since the message passing mechanism is elided.
@@
 \subsubsection{The Synchronous Kibitzer}\label{sec:kibitzer}
-Synchronous observers can be used to model the environment, and in particular, to compositionally inject faults.
+Synchronous observers can be used to model the environment, and in particular, inject faults in a compositional way.
 The typical approach to injecting faults into a formal model of a distributed system is to add new state variables to each process representing its fault state.
-Then additional state variables are added to represent the nondeterminism introduced by faults into the state or messages by a process.
+Then additional state variables are added to represent the non-determinism introduced by faults into the state or messages by a process.
 
@@
 
-Not only is modeling faults expensive, but it is generally non-compositional insofar as the fault model pervades the entire model.
+Not only is modeling faults generally expensive, but in typical practice it is generally non-compositional insofar as the fault model pervades the entire model.
 
-Another difficulty with adding faults to a model is that the number of states that must be introduced may be dependent non-obviously on other aspects of the fault model, specific protocol, and system size.
+Another difficulty with adding faults to a model is that the number of states that must be introduced may depend non-obviously on other aspects of the fault model, specific protocol, and system size.
 Furthermore, since SAT-based model-checking is NP-complete, it is important not to introduce any state beyond the minimal required.
@@
 A little thought will show that the way faulty relays have their most significant impact is by tipping the majority vote in a receiver one way or the other, and this can only be achieved if they use the same values as nonfaulty relays.
-Hence, we decide against further extension to the range of values~\cite{}.
+Hence, we decide against further extension to the range of values~\cite{Rushby:OM1}.
 \end{quote}
@@
 The kibitzer decomposes the state and transitions associated with the fault model from the system.
-For the sake of concreteness, we introduce a particular fault model, the hybrid fault model of Thambidurai and Park~\cite{}.
-The fault model distinguishes Byzantine (or arbitrary), symmetric, and manifest faults.
-This fault model applies to broadcast systems, in which a process is expected to broad the same value to multiple receivers.
+For the sake of concreteness, we introduce a particular fault model, the hybrid fault model of Thambidurai and Park~\cite{hybrid}.
+This fault model distinguishes Byzantine (or arbitrary), symmetric, and manifest faults.
+It  applies to broadcast systems in which a process is expected to broadcast the same value to multiple receivers.
 A \emph{Byzantine fault} is one in which a process that is intended to broadcast the same value to other processes may instead broadcast arbitrary values to different receivers (including no value or the correct value).
@@
 
-%% In specifying a fault-tolerant system, we must also specify a \emph{maximum fault assumption} (MFA), which specifies the maximum number of of each kind of fault that may be present in the system for some system property to hold.
+%% In specifying a fault-tolerant system, we must also specify a \emph{maximum fault assumption} (MFA), which specifies the maximum number of each kind of fault that may be present in the system for some system property to hold.
 
-Define an enumeration of fault types $$\textnormal{FAULT\_TYPE} = \{none, byz, sym, man\}$$ As in the previous section, let $\Id \subset \mathbb{N}$ be a finite set of indices, and let the variable $$faults: \Id \rightarrow \textnormal{FAULT\_TYPE}$$ range over possible mappings from process indices to faults.
+Define an enumeration of fault types $$\textnormal{FAULT\_TYPE} = \{none, byz, sym, man\}.$$ As in the previous section, let $\Id \subset \mathbb{N}$ be a finite set of process indices, and let the variable $$faults: \Id \rightarrow \textnormal{FAULT\_TYPE}$$ range over possible mappings from processes to faults.
 
@@
 \noindent
-For symmetric faults, we there is no requirement that the messages sent are the ones received, only that every recipient receives the same value.
+For symmetric faults, there is no requirement that the messages sent are the ones received, only that every recipient receives the same value.
 \begin{align*}
@@
 In model-checking, counterexamples to properties are traces of states.
-A state is a collection of variable values, and in large models, there are a lot.
 A single state includes assignments to each of the variables, and the counterexample traces we dealt with (see Section~\ref{sec:experimental}) can contain hundreds of state variables.
-In practice, counterexamples depend on a small subset of these.
+In practice, the essence of a counterexample depends on a small subset of these.
 The user has to manually separate the state variables that matter from those that do not for a specific property.
-It is not enough prevent variables that stutter from displaying since some state variables might be updated without affecting the property.
-The problem is particularly pertinent when model-checking models that have sufficient detail for implementation synthesis, as is the goal in our work.
+Moreover, It is not enough to prevent variables that stutter from displaying (a feature that many off-the-shelf model-checkers provide) since some state variables might be updated without affecting the property.
+The problem is particularly pertinent when formal models have sufficient detail for implementation synthesis, as is the goal in our work.
 
@@
 
-But the relationship between many of the state variables have an intuitive structure that can be characterized by a state machine imposed over it.
+But the relationship between many of the state variables often has an intuitive structure, one familiar to the designer, that can be characterized by a state machine imposed over it.
 The state machine describes discrete modes or configurations.
@@
 
-The idea is an implementation of abstract state machines (ASMs).
+This idea is an implementation of abstract state machines (ASMs).
 ASMs are a generalization of finite state machines originally developed by Gurevich to generalize state machines to operate over arbitrary data structures~\cite{asm}.
@@
 
+\ben{I'm confused by this definition, but I haven't looked at the reference, so maybe it would be clear if I did.
+My intuitive notion of ASM involves a set of predicates B which *cover* the reachable states and a transition relation over B that abstracts the one over S.}
+
 Our use of ASMs to improve traceability is even more restrictive.
@@
 
-\lee{probably need an example here.}
+\begin{example}
+Consider a simple distributed system involving 3 processors: $S$, $L_1$, and $L_2$.
+The purpose of the system is for $S$ to broadcast a message to $L_1$ and $L_2$ after which they perform a computation on the message and update their local state with the result.
+One might model this system as a transition system whose state is the product of the local processor states along with the state of the environment and any globally important phenomena such as time.
+
+\begin{figure}
+\centering
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3cm,
+  thick,main node/.style={circle,fill=blue!20,draw,
+  font=\sffamily\Large\bfseries,minimum size=8mm}]
+
+  \node[main node] (S) at (0,0) {$S$};
+  \node[main node] (L1) at (2,1) {$L_1$};
+  \node[main node] (L2) at (2,-1) {$L_2$};
+
+  \path[every node/.style={font=\sffamily\small,
+          fill=white,inner sep=1pt}]
+    (S)  edge (L1)
+         edge (L2);
+
+\end{tikzpicture}
+\label{fig:exa-trans}
+\caption{Simple message passing system}
+\end{figure}
+
+Intuitively, there are subsets of the overall state space which are important to understanding the evolution of such a system.
+For example, let $A_1$ denote the subset of states in which the processor $S$ has not sent a message, let $A_2$ be the subset in which $S$ has broadcast a message, and let $A_3$ be the subset in which $L_1$ and $L_2$ have each received and processed the message.
+The union $A_1 \cup A_2 \cup A_3$ covers the set of reachable states of the system and defines an ASM that is useful for reasoning about its dynamics and understanding counter-example traces.
+\end{example}
 
@@
 properties.
-Validity states that if the general is non-faulty, then every lieutenant outputs the value sent by the general.
+Validity states that if the general is nonfaulty, then every lieutenant outputs the value sent by the general.
 More formally, let the general broadcast $v \neq ERR$, and let $l_i$ denote the output for lieutenant $i$:

### 2016-12-02 doc/om-paper/paper.tex
unpushed changes to paper.

@@
 \end{abstract}
+
+\lee{grep to get consistent manifest/benign, also lieutenant vs relay/receiver.}
 
@@
 
+\lee{what we have defined are actually strictly stronger than synchronous observers since they can modify state.}
+
 \subsubsection{Model}
@@
 Then additional state variables are added to represent the non-determinism introduced by faults into the state or messages by a process.
+
+\lee{cite abstractions paper and point out that we can model faults on transmission rather than node}
 
@@
   \centering
-  \begin{tabular}{p{3cm}|p{3cm}|l|l|}
-                                  & Modules modified    & Lemmas modified & Lemma classes modified\\
+  \begin{tabular}{|p{3cm}|p{3cm}|l|l|}
+  \hline
+                                  & Transition systems & Definitions & Invariants & Invariant classes \\
   \hline \hline
-  Byzantine-only fault model      & 0                   & 0               & 0 \\
+  Omissive-asymetric faults       & 0                   & 4          & 2          & 1                 \\
   \hline
-  Time-triggered message-passing  & \texttt{clock}, \texttt{relay}
-                                  , \texttt{receiver}   & 0               & 0 \\
+  Time-triggered messaging        & 3                   & \lee{todo} & 0          &                   \\
+  Mid-value selection             &\lee{todo}           &  \lee{todo}&\lee{todo}  & \lee{todo}        \\
   \hline
@@
 \end{figure}
-To demonstrate the modularity of the model and verification approach, in this section, we explore variants to the model and discover the effort required to implement the modifications and repair the proofs.
+To demonstrate the modularity of the model and verification approach, in this section, we explore variants to the model and report the effort required to implement the modifications and repair the proofs.
 The results are summarized in Figure~\ref{sec:effort} and sketched below.
+In the table, for each modification, we report how much of the model must be modified.
+Specifically, we report on the number of synchronously-composed transition systems, function definitions, invariants, and invariants classes (as defined in Section~\ref{sec:invariants}) that must be modified.
 
-\paragraph{Fault model.}
-Modifying the fault model is easy. In involves modifying the definition of a single function, \texttt{mfa}, that takes as an argument the static assignment of node identifiers to faults. Moreover
+\paragraph{Omissive asymetric faults.}
+Remove faults already described by the fault model is easy.
+Recall that in our model faults do not appear in the system specification and only operate on the calendar.
+Removing a fault from the system requires only setting the number of a particular kind of faults to zero in the maximum fault assumption.
+
+Adding new kinds of faults requires more work but is modular.
+Consider adding \emph{omissive asymmetric faults}, a restriction of Byzantine faults in which a broadcaster either sends the correct value or a manifest fault~\cite{}, to the fault model.
+Doing so requires modifying none of transition systems.
+However, we modify four definitions (\texttt{msg()}, characterizing the result on the calendar due to a faulty message; \texttt{omissive_asymmetric_fault_constraint()}, defining the fault, and two functions that characterize the overall maximum fault assumption).
+Two invariants, both in the class of invariants cover faults, are extended to cover the cases where a sender is omissive asymmetric.
 
 \paragraph{Time-triggered messaging.}
+Oral Messages and its variants are synchronous protocols, in which communication occurs in rounds. In our original model, the nodes do not explicitly depend on a clock.
+
+\lee{todo}
+
+\paragraph{Mid-value selection.}
+
+\lee{todo}
+
+\lee{any other modifications?}
 

### 2016-12-02 doc/om-paper/paper.tex
Paper updates, fixes

@@
   \centering
-  \begin{tabular}{|p{3cm}|p{3cm}|l|l|}
+  \begin{tabular}{|p{3cm}|l|l|l|l|}
   \hline
-                                  & Transition systems & Definitions & Invariants & Invariant classes \\
+                                  & Transition systems  &Definitions & Invariants & Invariant classes \\
   \hline \hline
@@
   \hline
-  Time-triggered messaging        & 3                   & \lee{todo} & 0          &                   \\
+  Time-triggered messaging        & 3                   &            &            &                   \\
+  \hline
   Mid-value selection             &\lee{todo}           &  \lee{todo}&\lee{todo}  & \lee{todo}        \\
@@
 Doing so requires modifying none of transition systems.
-However, we modify four definitions (\texttt{msg()}, characterizing the result on the calendar due to a faulty message; \texttt{omissive_asymmetric_fault_constraint()}, defining the fault, and two functions that characterize the overall maximum fault assumption).
+However, we modify four definitions (\texttt{msg()}, characterizing the result on the calendar due to a faulty message; \texttt{omissive\_asymmetric\_fault\_constraint()}, defining the fault, and two functions that characterize the overall maximum fault assumption).
 Two invariants, both in the class of invariants cover faults, are extended to cover the cases where a sender is omissive asymmetric.
@@
 \paragraph{Time-triggered messaging.}
-Oral Messages and its variants are synchronous protocols, in which communication occurs in rounds. In our original model, the nodes do not explicitly depend on a clock.
+A \emph{time-triggered} distributed system is one in which nodes are independently clocked, but clock constraints (ensured by periodic clock synchronization) allow the model to appear as if it is executing synchronously.
+
+Oral Messages and its variants are synchronous protocols, in which communication occurs in rounds.
+Typically, this is modeled by synchronously composing lieutenants.
+Our calendar-based model is more general, in which all nodes are asynchronously composed.
+That said,
 

### 2016-12-02 doc/om-paper/paper.tex
Add \OMH macro

--- 2016-12-02 doc/om-paper/paper.tex
@@
 \newcommand{\OM}[1]{\ensuremath{\mathrm{OM}(#1)}}
+\newcommand{\OMH}{\ensuremath{\mathrm{OMH}}}
 \newcommand{\Id}{\ensuremath{\mathrm{Id}}}
@@
 
-To demonstrate the modeling technique, in Section~\ref{sec:byz} we present a formally-verified implementation of the Oral Messages Hybrid algorithm (OMH)~\cite{Lincoln-Rushby}, implementing the Boyer-Moore Fast Majority Vote algorithm for algorithm~\cite{Boyer-Moore:91}.
+To demonstrate the modeling technique, in Section~\ref{sec:byz} we present a formally-verified implementation of the Oral Messages Hybrid algorithm (\OMH)~\cite{Lincoln-Rushby}, implementing the Boyer-Moore Fast Majority Vote algorithm for algorithm~\cite{Boyer-Moore:91}.
 The verification is interesting in its own right, as it is the first fully parametric (on the number of nodes) model-checked \emph{implementation} of the algorithm.
@@
 
-To demonstrate the model and techniques, we model an implementation of the Oral Messages Hybrid (OMH) algorithm~\cite{}, a more sophisticated version of the class Oral Messages (OM) algorithm.
-OM was first formulated by Lamport, Shostak, and Pease~\cite{om} to achieve agreement in a distributed system while tolerating Byzantine faults.
-OMH was originally developed by Thambidurai and Park~\cite{hybrid} to achieve the same result in the presence of a more sophisticated fault model that includes symmetric and benign faults.
-However, OMH had a bug, as originally formulated, which was corrected and formally verified by Lincoln and Rushby using interactive theorem-proving~\cite{csl-93-2}.
+To demonstrate the model and techniques, we model an implementation of the Oral Messages Hybrid ($\OMH$) algorithm~\cite{csl-93-2}, a more sophisticated version of the classic Oral Messages ($\OM$) algorithm.
+$\OM$ was first formulated by Lamport, Shostak, and Pease~\cite{om} to achieve agreement in a distributed system while tolerating Byzantine faults.
+$\OMH$ was originally developed by Thambidurai and Park~\cite{hybrid} to achieve the same result in the presence of a more sophisticated fault model that includes symmetric and benign faults.
+However, $\OMH$ had a bug, as originally formulated, which was corrected and formally verified by Lincoln and Rushby using interactive theorem-proving~\cite{csl-93-2}.
 
@@
 
-\subsection{OMH(1) Algorithm}
-OMH is a recursive algorithm that proceeds in rounds of communication.
-Let $OMH(m)$ be a recursive function implementing the algorithm, parameterized by the number of rounds, $m$.
+\ben{The word "model" is very overloaded in the paper so far.
+I wonder if we could use a different word like "modeling framework", or something more concise, for the one conotation and just "model" for the concrete mathematical description of a particular system?}
+
+\subsection{$\OMH(m)$ Algorithm}
+$\OMH$ is a recursive algorithm that proceeds in rounds of communication.
+Let $\OMH(m)$ be a recursive function implementing the algorithm, parameterized by the number of rounds, $m$.
 Consider a finite set of nodes $N$.
@@
   \item\label{om:two} Let $l_v$ be the value received by $l \in L$ from $g$.
-Then for each $l$, execute $OMH(m-1)$, assigning $l$ to be the general and $L \setminus \{l\}$ to be the lieutenants.
+Then for each $l$, execute $\OMH(m-1)$, assigning $l$ to be the general and $L \setminus \{l\}$ to be the lieutenants.
 $l$ sends $l_v$ or $R$, if $l_v = ERR$.
@@
 \noindent
-In particular, $OMH(1)$ includes two rounds of broadcast communication, one in which the general broadcasts, and one in which the lieutenants exchange their values.
+In particular, $\OMH(1)$ includes two rounds of broadcast communication, one in which the general broadcasts, and one in which the lieutenants exchange their values.
 
-OMH is designed to preserve \emph{validity} and \emph{agreement}
+$\OMH$ is designed to preserve \emph{validity} and \emph{agreement}
 properties.
@@
   \label{fig:benchmark}
-  \caption{Benchmark of full proof computation time for OMH(1) implementation. Times are in seconds with a timeout (TO) limit of one hour. Dashes ('-') denote no benchmark was run.}
+  \caption{Benchmark of full proof computation time for $\OMH(1)$ implementation. Times are in seconds with a timeout (TO) limit of one hour. Dashes ('-') denote no benchmark was run.}
 \end{figure}
@@
 Bokor~\emph{et~al.} describe a message-passing model for synchronous distributed algorithms that is particularly amenable to partial-order reduction for explicit-state model-checking~\cite{Bokor2010}.
-That said, their model of OM(1) (not OMH(1)), timeouts after 10~hours, and we assume a voting algorithm is not explicitly specified.
+That said, their model of $\OM(1)$ (not $\OMH(1)$), timeouts after 10~hours, and we assume a voting algorithm is not explicitly specified.
 \lee{for what size?

### 2016-12-02 doc/om-paper/paper.tex

@@
   \centering
-  \begin{tabular}{p{3cm}|p{3cm}|l|l|}
-                                  & Modules modified    & Lemmas modified & Lemma classes modified\\
+  \begin{tabular}{|p{3cm}|l|l|l|l|}
+  \hline
+                                  & Transition systems  &Definitions & Invariants & Invariant classes \\
   \hline \hline
-  Byzantine-only fault model      & 0                   & 0               & 0 \\
+  Omissive-asymetric faults       & 0                   & 4          & 2          & 1                 \\
   \hline
-  Time-triggered message-passing  & \texttt{clock}, \texttt{relay}
-                                  , \texttt{receiver}   & 0               & 0 \\
+  Time-triggered messaging        & 3                   &            &            &                   \\
+  \hline
+  Mid-value selection             &\lee{todo}           &  \lee{todo}&\lee{todo}  & \lee{todo}        \\
   \hline
@@
 \end{figure}
-To demonstrate the modularity of the model and verification approach, in this section, we explore variants to the model and discover the effort required to implement the modifications and repair the proofs.
+To demonstrate the modularity of the model and verification approach, in this section, we explore variants to the model and report the effort required to implement the modifications and repair the proofs.
 The results are summarized in Figure~\ref{sec:effort} and sketched below.
+In the table, for each modification, we report how much of the model must be modified.
+Specifically, we report on the number of synchronously-composed transition systems, function definitions, invariants, and invariants classes (as defined in Section~\ref{sec:invariants}) that must be modified.
 
-\paragraph{Fault model.}
-Modifying the fault model is easy. In involves modifying the definition of a single function, \texttt{mfa}, that takes as an argument the static assignment of node identifiers to faults. Moreover
+\paragraph{Omissive asymetric faults.}
+Remove faults already described by the fault model is easy.
+Recall that in our model faults do not appear in the system specification and only operate on the calendar.
+Removing a fault from the system requires only setting the number of a particular kind of faults to zero in the maximum fault assumption.
+
+Adding new kinds of faults requires more work but is modular.
+Consider adding \emph{omissive asymmetric faults}, a restriction of Byzantine faults in which a broadcaster either sends the correct value or a manifest fault~\cite{}, to the fault model.
+Doing so requires modifying none of transition systems.
+However, we modify four definitions (\texttt{msg()}, characterizing the result on the calendar due to a faulty message; \texttt{omissive\_asymmetric\_fault\_constraint()}, defining the fault, and two functions that characterize the overall maximum fault assumption).
+Two invariants, both in the class of invariants cover faults, are extended to cover the cases where a sender is omissive asymmetric.
 
 \paragraph{Time-triggered messaging.}
+A \emph{time-triggered} distributed system is one in which nodes are independently clocked, but clock constraints (ensured by periodic clock synchronization) allow the model to appear as if it is executing synchronously.
+
+Oral Messages and its variants are synchronous protocols, in which communication occurs in rounds.
+Typically, this is modeled by synchronously composing lieutenants.
+Our calendar-based model is more general, in which all nodes are asynchronously composed.
+That said,
+
+\lee{todo}
+
+\paragraph{Mid-value selection.}
+
+\lee{todo}
+
+\lee{any other modifications?}
 

### 2016-12-02 doc/om-paper/paper.tex
paper edits with Ben.

@@
 \subsection{Synchronous Observers}\label{sec:sync}
+
+\lee{Lee will rework this section; update references to it.}
+
 Synchronous observers are state machines that are synchronously composed with a system specification to check properties over the system's state variables, i.e.
@@
 \begin{definition}[Abstract state machine (ASM)]
+
+\lee{Ben will redo. I'd just remove the Gurevich reference}
+
 Let $\mathcal{M} = (S, I, \rightarrow)$ be a transition system and let $B \subseteq 2^S$ be a set of predicates over $S$.
@@
   \hline
-  Time-triggered messaging        & 3                   &            &            &                   \\
+  Time-triggered messaging        & 3                   & 3          & 5          & 2                 \\
   \hline
@@
 \end{figure}
+
+\lee{Ben: add fault-model removed plus mid-value selection models}
+
 To demonstrate the modularity of the model and verification approach, in this section, we explore variants to the model and report the effort required to implement the modifications and repair the proofs.
@@
 \paragraph{Time-triggered messaging.}
-A \emph{time-triggered} distributed system is one in which nodes are independently clocked, but clock constraints (ensured by periodic clock synchronization) allow the model to appear as if it is executing synchronously.
+A \emph{time-triggered} distributed system is one in which nodes are independently clocked, but clock constraints (ensured by periodic clock synchronization) allow the model to appear as if it is executing synchronously~\cite{}.
+
+We create three definitions: encoding nondeterministic message delay, and two definitions that are helpers in invariants.
+The guards in the relays and receivers are modified to latch messages received outside the specified windows as being manifest faults.
+Additionally, the ASM definition is modified to track the times in the calendar, not just the messages.
+Two new invariants calendar invariants are introduced, stating that the calendar messages are either empty, or their time-stamps fall within the respective message windows.
+Then, three invariants classifying faults are relaxed to allow for the possibility of faulty nodes sending manifestly-faulty messages.
 

### 2016-12-04 doc/om-paper/paper.tex
started paper edits...

@@
 \lee{grep to get consistent manifest/benign, also lieutenant vs relay/receiver.}
+\lee{grep for synchronous observers section references}
 
@@
 System designers create software and hardware implementations to test, simulate, and deploy.
-Discrepancies between implementations and algorithmic models can arise if the latter is abstracted too much from the former, particularly if those abstractions are ad-hoc and system specific \ben{for example?}.
+Discrepancies between implementations and algorithmic models can arise if the latter is abstracted too much from the former~\cite{paxos}, particularly if those abstractions are ad-hoc and system specific.
 Furthermore, as implementations are modified to explore the design space, it is easy for the formal model and the implementation to become inconsistent.
@@
 
-In this paper, we develop a formal model to address these concerns.
+In this paper, we develop a formal framework to address these concerns.
 We combine ideas from the literature and folklore to build scalable and modular formal models of fault-tolerant distributed systems.
-We describe a generic formal model for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}.
-There are three aspects of the model: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) \emph{synchronous observers}, originally from the synchronous languages community~\cite{Halbwachs-Lagnier:92,Halbwachs-Lagnier:94} and popularized by Rushby for model-checking~\cite{Rushby:SAS14}; and (3) \emph{abstract state machines}, useful for abstracting counterexamples.
-The model is suitable for infinite-state model-checking.
+We describe a generic formal framework for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}.
+There are three aspects of the framework: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) \emph{synchronous observers}, originally from the synchronous languages community~\cite{Halbwachs-Lagnier:92,Halbwachs-Lagnier:94} and popularized by Rushby for model-checking~\cite{Rushby:SAS14}; and (3) \emph{abstract state machines}, useful for abstracting counterexamples.
+The framework is suitable for infinite-state model-checking.
 
-To demonstrate the modeling technique, in Section~\ref{sec:byz} we present a formally-verified implementation of the Oral Messages Hybrid algorithm (\OMH)~\cite{Lincoln-Rushby}, implementing the Boyer-Moore Fast Majority Vote algorithm for algorithm~\cite{Boyer-Moore:91}.
+To demonstrate the framework, in Section~\ref{sec:byz} we present a formally-verified implementation of the Oral Messages Hybrid algorithm (\OMH)~\cite{Lincoln-Rushby}, implementing the Boyer-Moore Fast Majority Vote algorithm for algorithm~\cite{Boyer-Moore:91}.
 The verification is interesting in its own right, as it is the first fully parametric (on the number of nodes) model-checked \emph{implementation} of the algorithm.
@@
 % ------------------------------------------------------------
-\section{Formal Model}\label{sec:model}
-Here we describe our formal model specialized for fault-tolerant distributed systems.
-The model draws on three principal abstractions: calendar automata, synchronous observers, and abstract state machines; we describe each below.
+\section{Formal Framework}\label{sec:model}
+Here we describe our formal framework specialized for fault-tolerant distributed systems.
+The framework draws on three principal abstractions: calendar automata, synchronous observers, and abstract state machines; we describe each below.
 
@@
 To avoid this problem, Dutetre and Sorea developed \emph{calendar automata}~\cite{bruno}, which is itself inspired by event calendars used in discrete-event simulation.
+Indeed, while calendar automata have historically been used for verifying real-time systems, they can be used to verify non real-time systems, too, specified in a discrete-event formulation.
+We discuss taking a non real-time model and easily transforming it to a real-time time-triggered model in Section~\ref{sec:modular}.
 
@@
 
-\item \label{cal:d} In all states, if the current time equals a timeout, then the only transitions enabled are calendar event updates associated with the timeout: $\forall \sigma \in S$, $\exists (e_i, t_i) \in \sigma(cal)$ such that $\sigma(now) = t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma'(now) = \sigma(now)$, $\sigma'(c_j) = \sigma(c_j)$ for all $c_j \in \sigma(c)$ such that $c_j \neq c_i$, and $\sigma'(t_i) > \sigma(t_i)$.
+\item \label{cal:d} In all states, if the current time equals a timeout, then the only transitions enabled are calendar event updates associated with the timeout: $\forall \sigma \in S$, $\exists (e_i, t_i) \in \sigma(cal)$ such that $\sigma(now) = t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma'(now) = \sigma(now)$, $\sigma'(c_j) = \sigma(c_j)$ for all $c_j \in \sigma(c)$ such that $c_j \neq c_i$, and $c_i \notin \sigma'(c)$.
+
 \end{enumerate}
-\ben{Do we also require that $c_i \notin \sigma'(c)$ ?}
 
@@
 
-\lee{technically, atomic time is just a convenience.}
 The event of receiving a message initiates a process to update its local state machine and generate additional messages to send.
@@
 
-\lee{note that we don't need full generality of send and receive events in OM(1)}
+%% \subsection{Synchronous Observers and }\label{sec:sync}
 
-\lee{note that we use calendar automata for non real-time system}
+\lee{Cutting synchronous observer section; need to talk about use for lemmas elsewhere.}
 
-\subsection{Synchronous Observers}\label{sec:sync}
-
-\lee{Lee will rework this section; update references to it.}
-
-Synchronous observers are state machines that are synchronously composed with a system specification to check properties over the system's state variables, i.e.
+%% Synchronous observers are state machines that are synchronously composed with a system specification to check properties over the system's state variables, i.e.
 they are a method for composing specifications with a program.
@@
 
-\lee{what we have defined are actually strictly stronger than synchronous observers since they can modify state.}
+%% \lee{what we have defined are actually strictly stronger than synchronous observers since they can modify state.}
 
-\subsubsection{Model}
-A synchronous observer is of the form
-$$\mathcal{S} | \mathcal{O}$$
-\noindent
-where `$|$' denotes synchronous composition of transition systems (defined formally below).
+%% \subsubsection{Model}
+%% A synchronous observer is of the form
+%% $$\mathcal{S} | \mathcal{O}$$
+%% \noindent
+%% where `$|$' denotes synchronous composition of transition systems (defined formally below).
 $\mathcal{S}$ is the \emph{system}, and $\mathcal{O}$ is the \emph{observer}.
@@
 
-\begin{definition}[Consistent states]
-  States $s, s'$ are consistent iff for every assignment $(x = v) \in s$ and $(x' =
-  v') \in s'$, $x \neq x'$ or $v = v'$.
-\end{definition}
-
-\begin{definition}[Reachable states]
-For a transition relation $\rightarrow$ over the states of $S$, for states $s_0,  s \in S$, we say $s$ is reachable from $s_0$ iff there exist states $s_1, \ldots, s_k \in S$ such that
-$$s_0 \rightarrow s_1 \rightarrow \ldots \rightarrow s_k \rightarrow s$$
-\end{definition}
-\noindent
-By convention, every state $s$ is reachable from itself.
-
-In the following definitions, let
-$M = (S, I, \rightarrow)$
-and
-$M' = (S', I', \rightarrow')$
-be transitions systems.
-
-%% \begin{definition}[Transition system union]
-%% %% $M \cup M' = (S \cup S', I \cup I', \rightarrow_{M|M'})$ where $\rightarrow_{M|M'}$ is the smallest relation such that $(s,s') \in \rightarrow_{M|M'}$ iff 
+%% \begin{definition}[Consistent states]
+%%   States $s, s'$ are consistent iff for every assignment $(x = v) \in s$ and $(x' =
+%%   v') \in s'$, $x \neq x'$ or $v = v'$.
 %% \end{definition}
 
-In defining the synchronous composition of two transition systems, we allow the possibility that composing two states can lead to inconsistent states (as opposed to the semantics in which composition of states $s_0$ and $s_1$ is the ordered pair $(s_0, s_1)$).
+%% \begin{definition}[Reachable states]
+%% For a transition relation $\rightarrow$ over the states of $S$, for states $s_0,  s \in S$, we say $s$ is reachable from $s_0$ iff there exist states $s_1, \ldots, s_k \in S$ such that
+%% $$s_0 \rightarrow s_1 \rightarrow \ldots \rightarrow s_k \rightarrow s$$
+%% \end{definition}
+%% \noindent
+%% By convention, every state $s$ is reachable from itself.
 
-\begin{definition}[Synchronous composition]
-The synchronous composition of $M$ and $M'$ is a transition system
-$$M | M' = (S_{M | M'}, I_{M | M'}, \rightarrow_{M | M'})$$
-\noindent
-such that
-\begin{itemize}
-\item $S_{M | M'} = S \cup S'$,
-\item $I_{M | M'} = \{ s_I \cup s_{I'} | s_I \in I\textnormal{, }s_{I'} \in I'\textnormal{, and } s_I \textnormal{ and } s_{I'}\textnormal{ are consistent}\}$
-\item and for $s, s' \in S_{M | M'}$, $s \rightarrow_{M | M'} s'$ iff
-  \begin{itemize}
-  \item there exist states $s_M, s'_M \in S$ such that $s_M \rightarrow s'_M$, $s_M \subseteq s$, and $s'_M \subseteq s'$,
-  \item there exist states $s_{M'}, s'_{M'} \in S'$ such that $s_{M'} \rightarrow s'_{M'}$, $s_{M'} \subseteq s$, and $s'_{M'} \subseteq s'$,
-  \item $s_M$ and $s_{M'}$ are consistent,
-  \item and $s'_M$ and $s'_{M'}$ are consistent.
-  \end{itemize}
-\end{itemize}
-\end{definition}
+%% In the following definitions, let
+%% $M = (S, I, \rightarrow)$
+%% and
+%% $M' = (S', I', \rightarrow')$
+%% be transitions systems.
+
+%% %% \begin{definition}[Transition system union]
+%% %% %% $M \cup M' = (S \cup S', I \cup I', \rightarrow_{M|M'})$ where $\rightarrow_{M|M'}$ is the smallest relation such that $(s,s') \in \rightarrow_{M|M'}$ iff 
+%% %% \end{definition}
+
+%% In defining the synchronous composition of two transition systems, we allow the possibility that composing two states can lead to inconsistent states (as opposed to the semantics in which composition of states $s_0$ and $s_1$ is the ordered pair $(s_0, s_1)$).
+
+%% \begin{definition}[Synchronous composition]
+%% The synchronous composition of $M$ and $M'$ is a transition system
+%% $$M | M' = (S_{M | M'}, I_{M | M'}, \rightarrow_{M | M'})$$
+%% \noindent
+%% such that
+%% \begin{itemize}
+%% \item $S_{M | M'} = S \cup S'$,
+%% \item $I_{M | M'} = \{ s_I \cup s_{I'} | s_I \in I\textnormal{, }s_{I'} \in I'\textnormal{, and } s_I \textnormal{ and } s_{I'}\textnormal{ are consistent}\}$
+%% \item and for $s, s' \in S_{M | M'}$, $s \rightarrow_{M | M'} s'$ iff
+%%   \begin{itemize}
+%%   \item there exist states $s_M, s'_M \in S$ such that $s_M \rightarrow s'_M$, $s_M \subseteq s$, and $s'_M \subseteq s'$,
+%%   \item there exist states $s_{M'}, s'_{M'} \in S'$ such that $s_{M'} \rightarrow s'_{M'}$, $s_{M'} \subseteq s$, and $s'_{M'} \subseteq s'$,
+%%   \item $s_M$ and $s_{M'}$ are consistent,
+%%   \item and $s'_M$ and $s'_{M'}$ are consistent.
+%%   \end{itemize}
+%% \end{itemize}
+%% \end{definition}
 
@@
 
-There are two practical benefits of the synchronous observer approach.
+%% There are two practical benefits of the synchronous observer approach.
 First, the system $\mathcal{S}$ contains no extraneous modeling context about the environment, assumptions, or requirements, making it easier to decompose $\mathcal{S}$ and to synthesize an implementation directly from it.
@@
 
-\subsubsection{The Synchronous Kibitzer}\label{sec:kibitzer}
-Synchronous observers can be used to model the environment, and in particular, inject faults in a compositional way.
-The typical approach to injecting faults into a formal model of a distributed system is to add new state variables to each process representing its fault state.
+\subsubsection{A Synchronous Kibitzer}\label{sec:kibitzer}
+The typical approach to modeling faults is to add new state variables to each process representing its fault state.
 Then additional state variables are added to represent the non-determinism introduced by faults into the state or messages by a process.
+As a simple example, we might define a node that sends a good message if it is non-faulty and a bad message otherwise.
+In pseudo-code using guarded commands, its definition might look like the following:
 
-\lee{cite abstractions paper and point out that we can model faults on transmission rather than node}
+\begin{verbatim}
+node:
+  health: Fault\_Type;
+
+  faulty(health)      --> send(bad\_msg);
+  non\_faulty(health) --> send(good\_msg);
+\end{verbatim}
 
@@
 
-Not only is modeling faults generally expensive, but in typical practice it is generally non-compositional insofar as the fault model pervades the entire model.
+\noindent
+But this typical approach mixes the specification of a node's behavior with the fault model, an aspect of the environment. Doing so divorces a node specification from its implementation.
 
-Another difficulty with adding faults to a model is that the number of states that must be introduced may depend non-obviously on other aspects of the fault model, specific protocol, and system size.
-Furthermore, since SAT-based model-checking is NP-complete, it is important not to introduce any state beyond the minimal required.
+A second difficulty with model-checking fault-tolerant systems in general is that modeling faults requires adding state and nondeterminism.
+The minimum number of additional states that must be introduced may depend non-obviously on other aspects of the fault model, specific protocol, and system size.
 Such constraints lead to ``meta-model'' reasoning, such as the following, as described by Rushby:
@@
 It would certainly be safe to introduce additional values for this purpose, but the performance of model checking is very sensitive to the size of the state space, so there is a countervailing argument against introducing additional values.
-A little thought will show that the way faulty relays have their most significant impact is by tipping the majority vote in a receiver one way or the other, and this can only be achieved if they use the same values as nonfaulty relays.
+A little thought will show that $\ldots$.
 Hence, we decide against further extension to the range of values~\cite{Rushby:OM1}.
@@
 
-\subsection{Model Sketch}
+\subsection{Model Sketch}\label{sec:sketch}
+\lee{Ben: can you introduce the relay/receiver unrolling of lieutenants here, since I refer to them later?}
+
 The verification is carried out in the Symbolic Analysis Laboratory~\cite{}, which contains a suite of model-checkers and a high-level specification language.

### 2016-12-04 doc/om-paper/paper.tex
fixed text in synch observer section.

@@
 \usepackage{graphicx}
-
+\usepackage{alltt}
 \usepackage{amsmath}
@@
 There are three aspects of the framework: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) \emph{synchronous observers}, originally from the synchronous languages community~\cite{Halbwachs-Lagnier:92,Halbwachs-Lagnier:94} and popularized by Rushby for model-checking~\cite{Rushby:SAS14}; and (3) \emph{abstract state machines}, useful for abstracting counterexamples.
-The framework is suitable for infinite-state model-checking.
+The framework is suitable for infinite-state model-checking supported by satisfiability modulo theories (SMT).
 
@@
 
+In specifying a system, a \emph{clock} module is synchronously composed with the system. The clock module updates time to the next calendar event (and stutters otherwise.
+
 %% \subsection{Synchronous Observers and }\label{sec:sync}
@@
 
-\subsubsection{A Synchronous Kibitzer}\label{sec:kibitzer}
+\subsection{A Synchronous Kibitzer}\label{sec:kibitzer}
 The typical approach to modeling faults is to add new state variables to each process representing its fault state.
@@
 node:
-  health: Fault\_Type;
-
-  faulty(health)      --> send(bad\_msg);
-  non\_faulty(health) --> send(good\_msg);
+  health: Fault_Type;
+  faulty(health)      --> send(bad_msg);
+  non_faulty(health) --> send(good_msg);
 \end{verbatim}
@@
 \noindent
-But this typical approach mixes the specification of a node's behavior with the fault model, an aspect of the environment. Doing so divorces a node specification from its implementation.
+But this typical approach mixes the specification of a node's behavior with the fault model, an aspect of the environment.
+Generally, nodes do not contain state variables assigned to their faults, or use their fault-status to determine their behavior!\footnote{There are exceptions.
+Benign faults may be detected by a node itself (e.g., in a built-in-test), an distributed diagnosis can let a node discover that others believe it is faulty~\cite{}.} The upshot is that combining faults and node state divorces the specification from its implementation.
 
@@
 The minimum number of additional states that must be introduced may depend non-obviously on other aspects of the fault model, specific protocol, and system size.
-Such constraints lead to ``meta-model'' reasoning, such as the following, as described by Rushby:
+Such constraints lead to ``meta-model'' reasoning, such as the following, as described by Rushby in describing the number of data values that a model must include to model the full range of Byzantine faults (defined later in this section):
 
@@
 
-To address these issues, we introduce a synchronous observer that injects faults that we call a \emph{synchronous kibitzer}.
-The kibitzer decomposes the state and transitions associated with the fault model from the system.
-For the sake of concreteness, we introduce a particular fault model, the hybrid fault model of Thambidurai and Park~\cite{hybrid}.
-This fault model distinguishes Byzantine (or arbitrary), symmetric, and manifest faults.
+The second problem is the most straightforward to solve.
+In infinite-state model-checking, we can use either the integers or the reals as the datatype for values.
+Fault-tolerant voting schemes, such as a majority vote or mid-value selection (see Section~\ref{sec:byz}), require only equality, or a total order, respectively, to be defined for the data.
+
+The solution to the first problem is more involved.
+Our solution is to introduce what we call a \emph{synchronous kibitzer} that symbolically injects faults into the model.
+The kibitzer decomposes the state and transitions associated with the fault model from the system itself.
+For the sake of concreteness in describing the synchronous kibitzer, we introduce a particular fault model, the hybrid fault model of Thambidurai and Park~\cite{hybrid}.
+This fault model distinguishes Byzantine, symmetric, and manifest faults.
 It  applies to broadcast systems in which a process is expected to broadcast the same value to multiple receivers.
-A \emph{Byzantine fault} is one in which a process that is intended to broadcast the same value to other processes may instead broadcast arbitrary values to different receivers (including no value or the correct value).
-A \emph{symmetric fault} is one in which a process may broadcast the same, but incorrect, value to other processes, and a \emph{manifest fault} is one in which a process's broadcast fault is detectable by the receivers; e.g., by performing a cyclic redundancy check (CRC) or because the value arrives outside of a predetermined window.
+A \emph{Byzantine} (or \emph{arbitrary} fault is one in which a process that is intended to broadcast the same value to other processes may instead broadcast arbitrary values to different receivers (including no value or the correct value).
+A \emph{symmetric} fault is one in which a process may broadcast the same, but incorrect, value to other processes.
+Finally, a \emph{manifest} (or \emph{benign} fault is one in which a process's broadcast fault is detectable by the receivers; e.g., by performing a cyclic redundancy check (CRC) or because the valuep arrives outside of a predetermined window.
 
@@
 
-Define an enumeration of fault types $$\textnormal{FAULT\_TYPE} = \{none, byz, sym, man\}.$$ As in the previous section, let $\Id \subset \mathbb{N}$ be a finite set of process indices, and let the variable $$faults: \Id \rightarrow \textnormal{FAULT\_TYPE}$$ range over possible mappings from processes to faults.
+Define a set of fault types $$\textnormal{Faults} = \{none, byz, sym, man\}.$$ As in the previous section, let $\Id \subset \mathbb{N}$ be a finite set of process indices, and let the variable $$faults: \Id \rightarrow \textnormal{Faults}$$ range over possible mappings from processes to faults.
 
@@
 $$broadcast(i, R, t, m) = \{m | j \in R \textnormal{ and } send(i, j, t) = m\}$$
-
-\lee{msg in SAL is recv here}
 
@@
 
-The upshot is that the fault model and fault-injection is completely modular to the specification of the protocol and node behavior.
+\paragraph{Transient faults.} If the $faults$ mapping is a constant, then faults are permanently but nondeterministically assigned to nodes.
+However, we can easily model \emph{transient faults}~\cite{} in which nodes are faulty temporarily by making $faults$ a state variable that is updated nondeterministically.
+Whether we model permanent or transient faults, a \emph{maximum fault assumption} (MFA) describes the maximum number of faults permitted in the system.
+The $faults$ mapping can be nondeterministically updated while satisfying the MFA using a constraint such as $faults \in \{f \, | \, mfa(f) \}$, where the MFA is defined by $mfa()$.
 

### 2016-12-06 doc/om-paper/paper.tex
paper: edits throughout.

@@
 abstractions come at the price of divorcing the model from its implementation
-and make refactoring difficult. In this work, we present a systematic model for
+and making refactoring difficult. In this work, we present a systematic model for
 fault-tolerant distributed system verification that combines calendar automata,
-synchronous observers, and abstract state machines specialized for
-model-checking. We demonstrate the efficacy of the framework by presenting the
+symbolic fault injection, and abstract state machines observers specialized for
+infinite-state model-checking. We demonstrate the efficacy of the framework by presenting the
 first model-checked \emph{implementation} of the Hybrid Oral Messages algorithm,
@@
 as the fault model, timing constraints, and local node behavior require only local modifications
-to the model and proofs. This work is carried out in SAL.
+to the model and proofs. This work is carried out in SAL model-checker.
 \end{abstract}
@@
 \lee{grep for synchronous observers section references}
+\lee{grep for first use of ``synchronous composition'' and give pointer to SAL semantics}
 
 \section{Introduction}
-
-\lee{talk about Amazon/TLA? Paxos made live?}
 
@@
 The verification of this class of systems has usually been at the algorithmic level, eliding details about a concrete implementation, and historically, has relied on formal models verified by interactive theorem-proving~\cite{om-acl2-impl,Young97:IC,csl-93-2,pvs}.
-If formal verification is to be introduced into the workflow of system designers, though, we need more automated methods.
+If formal verification is to be introduced into the workflow of system designers, though, we need more automated methods that scale for implementation-level models.
 (Mostly) automated proof techniques are required to reduce the need for specialized verification expertise.
@@
 Discrepancies between implementations and algorithmic models can arise if the latter is abstracted too much from the former~\cite{paxos}, particularly if those abstractions are ad-hoc and system specific.
-Furthermore, as implementations are modified to explore the design space, it is easy for the formal model and the implementation to become inconsistent.
+Furthermore, as implementations are modified to explore the design space, it is easy for the formal model and the implementation to become inconsistent, so the verification is no longer about the system deployed.
 
@@
 For example, the behaviors of nodes are naturally specified as a transition system in which transitions are guarded by the node's fault state.
-But faults are part of the environment; an implementation does not use its own fault status to choose actions!
+But faults are part of the environment; an implementation does not typically use its own fault status to choose actions!
 Another class of abstractions is used to simplify models.
@@
 We describe a generic formal framework for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}.
-There are three aspects of the framework: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) \emph{synchronous observers}, originally from the synchronous languages community~\cite{Halbwachs-Lagnier:92,Halbwachs-Lagnier:94} and popularized by Rushby for model-checking~\cite{Rushby:SAS14}; and (3) \emph{abstract state machines}, useful for abstracting counterexamples.
+There are three aspects of the framework: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) symbolic fault injection for decomposing the effects of faults from the system specification, and (3) \emph{abstract state machines observers}, useful for abstracting counterexamples.
 The framework is suitable for infinite-state model-checking supported by satisfiability modulo theories (SMT).
 
-To demonstrate the framework, in Section~\ref{sec:byz} we present a formally-verified implementation of the Oral Messages Hybrid algorithm (\OMH)~\cite{Lincoln-Rushby}, implementing the Boyer-Moore Fast Majority Vote algorithm for algorithm~\cite{Boyer-Moore:91}.
+To demonstrate the framework, in Section~\ref{sec:byz} focus on a particular representative algorithm as a case-study.
+We present a formally-verified implementation of the Oral Messages Hybrid algorithm (\OMH)~\cite{Lincoln-Rushby}, implementing the Boyer-Moore Fast Majority Vote algorithm for voting~\cite{Boyer-Moore:91}.
 The verification is interesting in its own right, as it is the first fully parametric (on the number of nodes) model-checked \emph{implementation} of the algorithm.
-In Section~\ref{sec:experimental}, we give performance metrics for the model.
+
+In Section~\ref{sec:experimental}, we give describe two sets of experiments on the model.
 First, we benchmark the model, showing that despite the additional complexity, it scales farther than other model-checking verifications at the algorithmic level.
@@
 A user should not have to discover all new invariants for new protocols or implementations.
-We show that most of the lemmas required are not protocol specific and are modular, concerning only specific aspects of the system.
-To demonstrate this, we modify the OHM implementation in various ways to show that only local modifications are required to the model and proofs.
+We show that most of the lemmas required are not protocol specific and are modular, concerning only specific aspects of the system, such as the timing, faults, or local node behavior.
+To demonstrate this, we modify the OHM implementation and show that in each case, only a small number of localized modifications are required to repair the model and proofs.
+The modifications we study are to add omissive-asymetric faults, make the implementation time-triggered, and replace the majority vote with a mid-value selection.
 
@@
 Here we describe our formal framework specialized for fault-tolerant distributed systems.
-The framework draws on three principal abstractions: calendar automata, synchronous observers, and abstract state machines; we describe each below.
+The framework draws on three principal abstractions: calendar automata, symbolic fault injection, and abstract state machines observers; we describe each below.
 
 \subsection{Calendar Automata}\label{sec:calendar}
-\emph{Timed automata} are specialized formalisms for verifying real-time systems~\cite{}.
+\emph{Timed automata} are specialized formalisms for verifying real-time systems~\cite{alur}.
 Timed automata assume the existence of continuously-varying real-valued variables, known as \emph{real-time clocks}.
@@
 However, real-time system verification in general-purpose model-checkers requires an explicit formalism of real-time progression.
-Trying to encode real-time clocks directly is difficult; in particular, one must avoid Zeno's paradox in which no progress is made because state transitions simply update real-valued variables by an infinitesimally small amount~\cite{bruno,lamport}.
-To avoid this problem, Dutetre and Sorea developed \emph{calendar automata}~\cite{bruno}, which is itself inspired by event calendars used in discrete-event simulation.
-Indeed, while calendar automata have historically been used for verifying real-time systems, they can be used to verify non real-time systems, too, specified in a discrete-event formulation.
+Trying to encode real-time clocks directly is difficult; in particular, one must avoid Zeno's paradox in which no progress is made because state transitions simply update real-valued variables by an infinitesimally small amount.
+To avoid this problem, Dutetre and Sorea developed \emph{calendar automata}~\cite{cal}, which is itself inspired by event calendars used in discrete-event simulation.
+
+%% Indeed, while calendar automata have historically been used for verifying real-time systems, they can be used to verify non real-time systems, too, specified in a discrete-event formulation.
 We discuss taking a non real-time model and easily transforming it to a real-time time-triggered model in Section~\ref{sec:modular}.
@@
 
-Again, following Dutetre and Sorea, we associate calendar events with channels in a distributed system~\cite{dutetre}.
+Again, following Dutetre and Sorea, we associate calendar events with channels in a distributed system~\cite{cal}.
 Specializing calendars to message passing does not lose generality since all external communication from an individual process can be abstracted as message passing.
@@
 
-In specifying a system, a \emph{clock} module is synchronously composed with the system. The clock module updates time to the next calendar event (and stutters otherwise.
+In specifying a system, a \emph{clock} module is synchronously composed with the system. The clock module updates time to the next calendar event, and stutters otherwise.
 
@@
 
-\subsection{A Synchronous Kibitzer}\label{sec:kibitzer}
+\subsection{Symbolic Fault Injection: a Synchronous Kibitzer}\label{sec:kibitzer}
 The typical approach to modeling faults is to add new state variables to each process representing its fault state.
@@
 
+\small
 \begin{verbatim}
@@
 \end{verbatim}
+\normalsize
 
@@
 The minimum number of additional states that must be introduced may depend non-obviously on other aspects of the fault model, specific protocol, and system size.
-Such constraints lead to ``meta-model'' reasoning, such as the following, as described by Rushby in describing the number of data values that a model must include to model the full range of Byzantine faults (defined later in this section):
+Such constraints lead to ``meta-model'' reasoning, such as the following, in which Rushby describes the number of data values that a particular protocol model must include to model the full range of Byzantine faults (defined later in this section):
 
@@
 A \emph{symmetric} fault is one in which a process may broadcast the same, but incorrect, value to other processes.
-Finally, a \emph{manifest} (or \emph{benign} fault is one in which a process's broadcast fault is detectable by the receivers; e.g., by performing a cyclic redundancy check (CRC) or because the valuep arrives outside of a predetermined window.
+Finally, a \emph{manifest} (or \emph{benign} fault is one in which a process's broadcast fault is detectable by the receivers; e.g., by performing a cyclic redundancy check (CRC) or because the value arrives outside of a predetermined window.
 
@@
 
+Thus, faults can be modeled solely in terms of their effects on sending and receiving messages. A node's specification does not have to depend on its fault status directly.
+
 \paragraph{Transient faults.} If the $faults$ mapping is a constant, then faults are permanently but nondeterministically assigned to nodes.
@@
 Whether we model permanent or transient faults, a \emph{maximum fault assumption} (MFA) describes the maximum number of faults permitted in the system.
-The $faults$ mapping can be nondeterministically updated while satisfying the MFA using a constraint such as $faults \in \{f \, | \, mfa(f) \}$, where the MFA is defined by $mfa()$.
+The $faults$ mapping can be nondeterministically updated during execution while satisfying the MFA using a constraint such as $faults \in \{f \, | \, mfa(f) \}$, where the MFA is defined by $mfa()$.
 
 % ------------------------------------------------------------
-\subsection{Abstract State Machines}\label{sec:abstract}\label{sec:asms}
+\subsection{Abstract State Machines Observers}\label{sec:abstract}\label{sec:asms}
+\lee{These are really observers. Let's call them abstract state machine observers and add a bit of the synchronous observers text here. Also, cite disjunctive invariants paper?}
+
 In model-checking, counterexamples to properties are traces of states.
-A single state includes assignments to each of the variables, and the counterexample traces we dealt with (see Section~\ref{sec:experimental}) can contain hundreds of state variables.
+A single state includes assignments to each of the variables, and the counterexample traces we examine (see Section~\ref{sec:experimental}) can contain hundreds of state variables.
 In practice, the essence of a counterexample depends on a small subset of these.
 The user has to manually separate the state variables that matter from those that do not for a specific property.
-Moreover, It is not enough to prevent variables that stutter from displaying (a feature that many off-the-shelf model-checkers provide) since some state variables might be updated without affecting the property.
+Moreover, It is not enough to prevent variables that stutter from displaying since some state variables might be updated without affecting the property.
 The problem is particularly pertinent when formal models have sufficient detail for implementation synthesis, as is the goal in our work.
@@
 $\OMH$ was originally developed by Thambidurai and Park~\cite{hybrid} to achieve the same result in the presence of a more sophisticated fault model that includes symmetric and benign faults.
-However, $\OMH$ had a bug, as originally formulated, which was corrected and formally verified by Lincoln and Rushby using interactive theorem-proving~\cite{csl-93-2}.
+However, $\OMH$ had a bug, as originally formulated, which was corrected and the mended algorithm was formally verified by Lincoln and Rushby using interactive theorem-proving~\cite{csl-93-2}.
 
@@
 \lee{Ben: can you introduce the relay/receiver unrolling of lieutenants here, since I refer to them later?}
+\lee{explain why there are 7 transition systems defined (source, recv, relay, abstractor, abstractorMonitor, clock, observer }
 
@@
 
-    \node[state] (calendar) {calendar (5)};
+    \node[state] (calendar) {calendar (1)};
     \node[state] (lieutenant) [right=1.75cm of calendar] {\begin{tabular}{c}lieutenant\\state-machine (2)\end{tabular}};
@@
 For example, the ASM lemmas depend on both the calendar automata and lieutenant state-machine lemmas.
-As can be seen, the proof structure is quite modular.
+As can be seen, the proof structure is modular.
 The calendar lemmas are general and independent of any particular protocol or fault model.
@@
 
-The benchmarks were performed on a server with Intel Xeon E312xx (Sandy
-Bridge) CPUs (mid 2011).
+We present benchmarks in Figure~\ref{fig:benchmark}. The benchmarks were
+performed on a server with Intel Xeon E312xx (Sandy Bridge) CPUs (mid 2011).
+The table provides execution times in seconds, with a timeout limit of one hour, for verifying the model, given a selected number of relays and receivers.
+The voting logic is in the receivers, so they have substantially more state than the relays, and dominate the execution time.
+The execution times sums the execution times for verifying each of the eleven lemmas individually, as well as the final agreement and validity theorems.
+Each proof incurs the full startup, parsing, type-checking, and model-generation time of SAL.
 
-For comparison, when the fault model is ``turned off'' by defining the fault
-type to contain a single 'nonfaulty' value and removing fault specific lemmas,
-model-checking scales further. For example when there are 3 relays and 3
-receivers the time is 25s (compared to 40s above), and when where are 5 relays
-and 5 receivers the time is 315s (compared to timeout above). See Figure
-\ref{fig:effort} for more comparisons.
+Note that the theorems hold even in the degenerate cases of one relay or one receiver.
 
-\lee{discuss results}
+\lee{Ben: can you run a couple of representative examples of Rushby's model, say for 5 and 6 receivers?}
+
+%% For comparison, when the fault model is ``turned off'' by defining the fault
+%% type to contain a single 'nonfaulty' value and removing fault specific lemmas,
+%% model-checking scales further. For example when there are 3 relays and 3
+%% receivers the time is 25s (compared to 40s above), and when where are 5 relays
+%% and 5 receivers the time is 315s (compared to timeout above). See Figure
+%% \ref{fig:effort} for more comparisons.
+
+%% \lee{discuss results}
 
@@
   \centering
-  \begin{tabular}{|p{3cm}|l|l|l|l|}
+  \begin{tabular}{|p{2.25cm}|p{2.25cm}|p{2.25cm}|p{2.25cm}|p{2.25cm}|}
   \hline
-                                  & Transition systems  &Definitions & Invariants & Invariant classes \\
+                                  & Transition systems (7 total)  &Definitions\linebreak (58 total) & Invariants\linebreak (11 total) & Invariant classes (5 total) \\
   \hline \hline
-  Omissive-asymetric faults       & 0                   & 4          & 2          & 1                 \\
+  Omissive-asymetric faults       & None                & 4 modified & 2 modified & faults             \\
   \hline
-  Time-triggered messaging        & 3                   & 3          & 5          & 2                 \\
+  Time-triggered messaging        & source, relays, receivers & 3 new      & 2 new, 3 modified & calendar, faults \\
   \hline
@@
   \end{tabular}
-  \caption{Refactoring effort for protocol modifications.}
+  \caption{Refactoring effort for protocol modifications, measured by which portions of the model have to be modified.}
   \label{fig:effort}
@@
 To demonstrate the modularity of the model and verification approach, in this section, we explore variants to the model and report the effort required to implement the modifications and repair the proofs.
-The results are summarized in Figure~\ref{sec:effort} and sketched below.
+The results are summarized in Figure~\ref{fig:effort} and sketched below.
 In the table, for each modification, we report how much of the model must be modified.
-Specifically, we report on the number of synchronously-composed transition systems, function definitions, invariants, and invariants classes (as defined in Section~\ref{sec:invariants}) that must be modified.
+We report on four aspects of the system: which transition systems making up the full model are modified (as described in Section~\ref{sec:model}), how many definitions have to be added or modified, the number of invariants that have to be added or modified, and which invariant classes (as defined in Section~\ref{sec:invariants}) those lemmas belong to.
 
 \paragraph{Omissive asymetric faults.}
-Remove faults already described by the fault model is easy.
+Removing faults already described by the fault model is easy.
 Recall that in our model faults do not appear in the system specification and only operate on the calendar.
@@
 
-Adding new kinds of faults requires more work but is modular.
+Adding new kinds of faults requires more work but is still modular.
 Consider adding \emph{omissive asymmetric faults}, a restriction of Byzantine faults in which a broadcaster either sends the correct value or a manifest fault~\cite{}, to the fault model.
@@
 
-We create three definitions: encoding nondeterministic message delay, and two definitions that are helpers in invariants.
+Changing the model to be time-triggered principally requires making the source, relays, and receivers driven explicitly by the passage of time.
+As well, a ``receive window'' is defined at which messages from non-faulty nodes should be received.
+Messages received outside the window are marked as coming from manifest-faulty senders.
+To implement a time-triggered model, we create three definitions: encoding nondeterministic message delay, and two definitions that are helpers in invariants.
 The guards in the relays and receivers are modified to latch messages received outside the specified windows as being manifest faults.
@@
 Then, three invariants classifying faults are relaxed to allow for the possibility of faulty nodes sending manifestly-faulty messages.
-
-Oral Messages and its variants are synchronous protocols, in which communication occurs in rounds.
-Typically, this is modeled by synchronously composing lieutenants.
-Our calendar-based model is more general, in which all nodes are asynchronously composed.
-That said,
-
-\lee{todo}
 
@@
 $k$-induction in infinite-state model-checking blurs the lines between interactive and automated theorem proving.
-IC3 can even be strengthened using $k$-induciton~\ref{pdr-kind}.
+IC3 can even be strengthened using $k$-induciton~\cite{pdr-kind}.
 Tools like Ivy can present graphical counterexamples to help interactively develop invariants~\cite{ivy}.
@@
 The Oral Messages algorithm was presented in the landmark paper by Lamport, Shostak, and Pease~\cite{om}.
-Later, a hybrid fault-model and modified Oral Messages algorithm was developed by Thambidurai and Park~\cite{hybrid}.
 OM(1) was subsequently verified in both the PVS and ACL2 interactive theorem-provers~\cite{Young97:IC}.
 Also in ACL2, an implementation of a circuit design to implement OM(1) is given~\cite{om-acl2-impl}; the low-level model most closely relates to our level of detail.
-A refinement-based verification approach is used, and OM(1) is specialized to a four-process system.
+A refinement-based verification approach is used, and OM(1) is specialized to a four-node system.
+It appears that a general purpose voting algorithm (like Fast MJRTY) is not verified as part of the work.
+A hybrid fault-model and modified Oral Messages algorithm was developed by Thambidurai and Park~\cite{hybrid}.
 Using the PVS interactive theorem-prover~\cite{pvs}, Lincoln and Rushby demonstrate a flaw in Thambidurai and Park's algorithm and present a fix to it~\cite{csl-93-2}.
 
-Using automated techniques, Gasc{\'{o}}n and Tiwari use SAT-based bounded synthesis to partially synthesize a variant of Oral Messages that tolerates transient faults~\cite{om1-synth}.
+Using automated techniques, Gasc{\'{o}}n and Tiwari use SAT-based bounded synthesis to partially synthesize a variant of Oral Messages that tolerates transient faults, specialized for three or four nodes~\cite{om1-synth}.
 Bokor~\emph{et~al.} describe a message-passing model for synchronous distributed algorithms that is particularly amenable to partial-order reduction for explicit-state model-checking~\cite{Bokor2010}.
-That said, their model of $\OM(1)$ (not $\OMH(1)$), timeouts after 10~hours, and we assume a voting algorithm is not explicitly specified.
-\lee{for what size?
-Also, Rushby's model?} Very recently, Jovanovi{\'{c}} and Dutertre use a ``flattened'' high-level model of OM(1) as a benchmark for IC3 augmented with $k$-induction~\cite{pdr-kind}.
-
-\lee{mention that ASMs similar to disjunction invariants}
+The model is efficient for up to five nodes, but results are not presented beyond that.
+Very recently, Jovanovi{\'{c}} and Dutertre use a ``flattened'' high-level model of OM(1) as a benchmark for IC3 augmented with $k$-induction~\cite{pdr-kind}.
 
@@
 The goal of the language is to synthesize both software and/or hardware implementations as well as formal models for verification.
-Before building such an ADSL, we needed a scalable general formal model to which to compile, leading to the work presented.
-An ADSL will make refactoring even easier.
-We have developed a preliminary ADSL that generates C code as well as formal models SRI's Sally~\cite{}, a model-checking descendent to SAL that will be described in a subsequent paper.\footnote{\url{https://github.com/GaloisInc/atom-sally}}
+Before building such an ADSL, we needed a scalable general formal model to which to compile, leading to the work presented in this paper.
+We hypothesize that an ADSL will make refactoring even easier, and we can generate invariants or invariant templates useful for verification.
+Indeed, we have developed a preliminary ADSL that generates C code as well as formal models SRI's Sally~\cite{}, a model-checking descendent to SAL that will be described in future work.\footnote{\url{https://github.com/GaloisInc/atom-sally}}
+
+Beyond building an ADSL, another avenue of research is producing a formal proof that a software implementation satisfies the node specification in our framework.
+While our model of node behavior is low-level, there are gaps.
+For example, our work is in SAL's language of guarded commands~\cite{} and needs to be either refined or verified to be equivalent to a software implementation's semantics.
+Another aspect is that behavior related to networking and benign fault diagnosis (e.g., computing a CRC) is are left abstract.
+In a software implementation, the \texttt{send()} and \texttt{recv()} functions would have to implement these behaviors.
+Finally, we have modeled messages as integers; but real data may be more complex.
+A synthesis approach may be suitable and has been used in the past for fault-tolerant clock-synchronization~\cite{ddd}.
 
 \lee{link to all results in github}
-\lee{talk about compositional verification of properties, even though SAL doesn't directly support it}
-\lee{talk about k-induction vs. PDR}
+\lee{talk about compositional verification of properties, even though SAL doesn't directly support it---maybe not so important}
+\lee{talk about k-induction vs. PDR---actually I don't think we have much to say}
 \lee{talk about lack of axiomatization in model-checking (multiple rushby bugs), but  tradeoff of deadlock. see proglema paper}
-\lee{note some details still elided for implementation ---e.g., CRCs, etc.}
 

### 2016-12-06 doc/om-paper/paper.tex
Update the experiments table with results for mid-value select
Add description of mid-value select variant

@@
   \hline \hline
+  No Fault Model                  & receivers & 8 modified & 2 modified & ASM, faults \\
+  \hline
   Omissive-asymetric faults       & None      & 4 modified & 2 modified & faults             \\
@@
   \hline
-  Mid-value selection             &\lee{todo}           &  \lee{todo}&\lee{todo}  & \lee{todo}        \\
+  Mid-value selection             & receivers & 4 new, 3 modified & 2 modified & ASM, voting \\
   \hline
@@
 
-\lee{Ben: add fault-model removed plus mid-value selection models}
-
 To demonstrate the modularity of the model and verification approach, in this section, we explore variants to the model and report the effort required to implement the modifications and repair the proofs.
@@
 
-\paragraph{Omissive asymetric faults.}
+\paragraph{Omissive asymmetric faults.}
 Removing faults already described by the fault model is easy.
@@
 \paragraph{Mid-value selection.}
-
-\lee{todo}
-
-\lee{any other modifications?}
+Our $\OMH(1)$ model leverages majority vote in order to tolerate faults.
+Another choice for the fault masking algorithm used here is mid-value select (or median).
+This choice is more common in applications involving hardware, signal selection (where signal values cannot be relied on to exactly match even in the absence of faults), or cases where information about congruence is useful.
+To implement mid-value select in our model, we allow messages sent to take values in $\mathbb{R}$ and the receiver transition system is modified in two ways.
+First, a second buffer is introduced which will hold the sorted contents of the main buffer once voting has commenced.
+Second, a mid-value select function is called on the sorted buffer and the result is stored as the receiver's vote.
+The only invariants needing modification were the ASM definition (to account for the values stored by the new buffer and the relation between it and the main buffer) and the voting invariant.
 

### 2016-12-07 doc/om-paper/paper.tex
Add model sketch section, futz with experiments table, add Rushby benchmark data

@@
 \usepackage{amsfonts}
+\usepackage{bm}
+\usepackage{multirow}
 
@@
 \newcommand{\Msg}{\ensuremath{\mathrm{Msg}}}
+\newcommand{\ERR}{\ensuremath{\mathrm{ERR}}}
+\newcommand{\TO}{\ensuremath{\bm{\mathrm{T}}}}
 
@@
 First, we benchmark the model, showing that despite the additional complexity, it scales farther than other model-checking verifications at the algorithmic level.
-The scalability is due principally to specifying invariants and using $k$-induction in SRI's SAL model-checker~\cite{sal}.
+The scalability is due principally to specifying invariants and using $k$-induction in SRI's SAL model-checker~\cite{SAL}.
 Developing invariants requires some user guidance.
@@
 
-First, we briefly describe the algorithm, sketch our model of the algorithm within the model presented in Section~\ref{sec:model}, then describe it's invariants.
-
-\ben{The word "model" is very overloaded in the paper so far.
-I wonder if we could use a different word like "modeling framework", or something more concise, for the one conotation and just "model" for the concrete mathematical description of a particular system?}
+First, we briefly describe the algorithm, sketch our model of the algorithm within the framework presented in Section~\ref{sec:model}, then describe it's invariants.
 
@@
 $\OMH$ is a recursive algorithm that proceeds in rounds of communication.
-Let $\OMH(m)$ be a recursive function implementing the algorithm, parameterized by the number of rounds, $m$.
+Here we give a recursive specification for $\OMH(m)$, parameterized by the number of rounds, $m$.
 Consider a finite set of nodes $N$.
@@
 Broadcast communication proceeds in rounds.
-Denote any message that is detectably faulty (e.g., fails a CRC) or is absent, by $ERR$.
+Denote any message that is detectably faulty (e.g., fails a CRC) or is absent, by $\ERR$.
 Additionally, in the algorithm, nodes report on values they have previously received.
-In doing so, nodes must differentiate \emph{reporting} $ERR$ from an $ERR$ itself.
+In doing so, nodes must differentiate \emph{reporting} $\ERR$ from an $\ERR$ itself.
 Let $R$ denote that an error is being reported.
@@
 \begin{itemize}
-\item {\bf $OM(0)$}: $g$ broadcasts a value to each lieutenant, and the lieutenants return the value received (or $ERR$).
-\item {\bf $OM(m)$, $m > 0$}:
+\item {\bf $\OM(0)$}: $g$ broadcasts a value to each lieutenant, and the lieutenants return the value received (or $\ERR$).
+\item {\bf $\OM(m)$, $m > 0$}:
   \begin{enumerate}
@@
 Then for each $l$, execute $\OMH(m-1)$, assigning $l$ to be the general and $L \setminus \{l\}$ to be the lieutenants.
-$l$ sends $l_v$ or $R$, if $l_v = ERR$.
-  \item For each lieutenant $l \in L$, remove all $ERR$ values received in Step~\ref{om:two} from executing $OHM(m-1)$.
+$l$ sends $l_v$ or $R$, if $l_v = \ERR$.
+  \item For each lieutenant $l \in L$, remove all $\ERR$ values received in Step~\ref{om:two} from executing $OHM(m-1)$.
 Compute the majority value over the remaining values, or $V$ if there is no majority.
@@
 
-$\OMH$ is designed to preserve \emph{validity} and \emph{agreement}
-properties.
+$\OMH$ is designed to ensure \emph{validity} and \emph{agreement} properties under suitable hypotheses on the number and type of faults in the system.
 Validity states that if the general is nonfaulty, then every lieutenant outputs the value sent by the general.
-More formally, let the general broadcast $v \neq ERR$, and let $l_i$ denote the output for lieutenant $i$:
+More formally, let the general broadcast $v \neq \ERR$, and let $l_i$ denote the output for lieutenant $i$:
 %
@@
 \subsection{Model Sketch}\label{sec:sketch}
-\lee{Ben: can you introduce the relay/receiver unrolling of lieutenants here, since I refer to them later?}
-\lee{explain why there are 7 transition systems defined (source, recv, relay, abstractor, abstractorMonitor, clock, observer }
 
-The verification is carried out in the Symbolic Analysis Laboratory~\cite{}, which contains a suite of model-checkers and a high-level specification language.
-In our work, we use infinite-state (SMT-based) $k$-induction~\cite{}.
+We have implemented $\OMH(1)$ (as well as many variants, described in Section~\ref{sec:modular}) in the Symbolic Analysis Laboratory~\cite{SAL} (SAL) language.
+The verification is also carried out in SAL, which contains a suite of model-checkers.
+In our work, we use infinite-state (SMT-based) $k$-induction~\cite{Moura_boundedmodel,Armando-et-al}.
 
-\lee{Maybe talk about MJRTY here. Maybe also talk about relaxing in synchronous communication in model.}
-\lee{talk about synchrony being loosened, so the general/receivers don't all send/receive at the same time. Talk about lieutenants being ``unrolled'' into relays and receivers, following Rushby.}
+We have chosen to follow Rushby~\cite{Rushby:OM1} in ``unrolling'' the communication among lieutenants into two sets of logical nodes: relays and receivers.
+A special node called the source takes the place of the general above.
+Thus, for example, in the case where 4 nodes execute $\OMH(1)$, our model consists of one source, three relays, and three receivers.
+We model communication through one-way, typed channels.
+The source broadcasts a message to each relay which, in turn, each broadcast their messages to all receivers.
+
+Our SAL model defines 7 transition systems in total: \texttt{clock}, {\texttt{source}, \texttt{relay} (parametrized over an ID), \texttt{receiver} (parametrized over an ID), \texttt{abstractor}, \texttt{observer}, \texttt{abstract\_monitor}.
+The first four of these are composed asynchronously, in an intermediate system we label \texttt{system}, and share access to a global calendar consisting of event slots $(\text{message}, \text{time})$, one for each channel in the system.
+\texttt{clock} is responsible for updating a global variable \texttt{t} representing time according to the rules in Section~\ref{sec:calendar}.
+The asynchronous composition gives us a relaxed form of the original specification, but it is possible without much additional work to enforce timing constraints, see Section~\ref{sec:modular} "Time-Triggered Messaging".
+The remaining three systems are observers of the previously defined asynchronous system.
+The \texttt{observer} controls the state of \texttt{validity} and \texttt{agreement} variables, setting them to \texttt{FALSE} if, when the receivers have computed their vote, the properties do not hold.
+The \texttt{abstractor} and \texttt{abstract\_monitor} monitor the state of \texttt{system} and update the values of a collection of 6 abstract state predicates.
+These predicates express different increasing levels of progress of \texttt{system}: from the source sending its message, to the relays broadcasting, to the receivers voting on received values.
 
@@
   \centering
-  \begin{tabular}{r||c|c|c|c|c|c|c|c|c|c|}
-      \multicolumn{11}{c}{Receivers} \\
-      \cline{2-11}
-   Relays &   1  &  2  &   3  &  4  &   5  &  6   &  7   &  8  &  9  &  10 \\
-      \hline \hline
+  \begin{tabular}{c|c||c|c|c|c|c|c|c|c|c|c|}
+      \multicolumn{2-12}{c}{Receivers} \\
+      \cline{2-12}
+          &     & 1  &  2  &   3  &  4  &   5  &  6   &  7   &  8  &  9  &  10 \\
+      \cline{2-12} \cline{2-12}
+      \multirow{10}{*}{Relays} &
             1   & 7    & 9   & 12   & 15  & 21   & 25   & 32   & 40  & 54  & 74  \\
-      \hline
-      2   & 17   & 14  & 21   & 30  & 42   & 53   & 74   & 99  & 144 & -   \\
-      \hline
-      3   & 21   & 22  & 40   & 50  & 81   & 102  & 155  & 279 & -   & -   \\
-      \hline
-      4   & 27   & 34  & 59   & 99  & 141  & 237  & 1114 & -   & -   & -   \\
-      \hline
-      5   & 22   & 94  & 125  & 335 & TO   & 1406 & -    & -   & -   & -   \\
-      \hline
-      6   & 36   & 132 & 2966 & 844 & 2457 & -    & -    & -   & -   & -   \\
-      \hline
-      7   & 83   & 487 & TO   & TO  & -    & -    & -    & -   & -   & -   \\
-      \hline
-      8   & 298  & TO  & TO   & -   & -    & -    & -    & -   & -   & -   \\
-      \hline
-      9   & 1428 & TO  & -    & -   & -    & -    & -    & -   & -   & -   \\
-      \hline
-     10   & TO   & -   & -    & -   & -    & -    & -    & -   & -   & -   \\
-      \hline
+      \cline{2-12}
+          & 2   & 17   & 14  & 21   & 30  & 42   & 53   & 74   & 99  & 144 & -   \\
+      \cline{2-12}
+          & 3   & 21   & 22  & 40   & 50  & 81   & 102  & 155  & 279 & -   & -   \\
+      \cline{2-12}
+          & 4   & 27   & 34  & 59   & 99  & 141  & 237  & 1114 & -   & -   & -   \\
+      \cline{2-12}
+          & 5   & 22   & 94  & 125  & 335 & \TO   & 1406 & -    & -   & -   & -   \\
+      \cline{2-12}
+          & 6   & 36   & 132 & 2966 & 844 & 2457 & -    & -    & -   & -   & -   \\
+      \cline{2-12}
+          & 7   & 83   & 487 & \TO   & \TO  & -    & -    & -    & -   & -   & -   \\
+      \cline{2-12}
+          & 8   & 298  & \TO  & \TO   & -   & -    & -    & -    & -   & -   & -   \\
+      \cline{2-12}
+          & 9   & 1428 & \TO  & -    & -   & -    & -    & -    & -   & -   & -   \\
+      \cline{2-12}
+          & 10   & \TO   & -   & -    & -   & -    & -    & -    & -   & -   & -   \\
+      \cline{2-12}
   \end{tabular}
+  \caption{Benchmark of full proof computation time for $\OMH(1)$ implementation. Times are in seconds with a timeout (\TO) limit of one hour. Dashes ('-') denote no benchmark was run.}
   \label{fig:benchmark}
-  \caption{Benchmark of full proof computation time for $\OMH(1)$ implementation. Times are in seconds with a timeout (TO) limit of one hour. Dashes ('-') denote no benchmark was run.}
 \end{figure}
@@
 
-\lee{Ben: can you run a couple of representative examples of Rushby's model, say for 5 and 6 receivers?}
+As a point of comparison, we also benchmarked Rushby's model~\cite{Rushby:OM1} on the same machine, using the same environment.
+When the number of receivers was fixed at 2, the verification time for increasing numbers of relays was: $0.12 s$, $0.20 s$, $2.56 s$, $8.24 s$, $28.0 s$, $549 s$, \TO, \TO, \TO (compare to the numbers in the second column of Figure~\ref{fig:benchmark}).
+Empirical evidence suggests that our verification, while slower for small numbers of nodes, scales much better for larger numbers of nodes.
 
@@
   \hline
-  Omissive-asymetric faults       & None      & 4 modified & 2 modified & faults             \\
+  Omissive Asymmetric Faults       & none      & 4 modified & 2 modified & faults             \\
   \hline
-  Time-triggered messaging        & source, relays, receivers & 3 new & 2 new, 3 modified & calendar, faults \\
+  Time-Triggered Messaging        & source, relays, receivers & 3 new & 2 new, 3 modified & calendar, faults \\
   \hline
-  Mid-value selection             & receivers & 4 new, 3 modified & 2 modified & ASM, voting \\
+  Mid-Value Selection & receivers & 4 new, 3 modified & 2 modified & ASM, voting \\
   \hline
@@
 
-\paragraph{Omissive asymmetric faults.}
+\paragraph{Omissive Asymmetric Faults.}
 Removing faults already described by the fault model is easy.
@@
 
-\paragraph{Time-triggered messaging.}
+\paragraph{Time-Triggered Messaging.}
 A \emph{time-triggered} distributed system is one in which nodes are independently clocked, but clock constraints (ensured by periodic clock synchronization) allow the model to appear as if it is executing synchronously~\cite{}.
@@
 
-\paragraph{Mid-value selection.}
+\paragraph{Mid-Value Selection.}
 Our $\OMH(1)$ model leverages majority vote in order to tolerate faults.
-Another choice for the fault masking algorithm used here is mid-value select (or median).
+Another choice for the fault masking algorithm used here is mid-value selection (or median).
 This choice is more common in applications involving hardware, signal selection (where signal values cannot be relied on to exactly match even in the absence of faults), or cases where information about congruence is useful.
-To implement mid-value select in our model, we allow messages sent to take values in $\mathbb{R}$ and the receiver transition system is modified in two ways.
+To implement mid-value selection in our model, we allow messages sent to take values in $\mathbb{R}$ and the receiver transition system is modified in two ways.
 First, a second buffer is introduced which will hold the sorted contents of the main buffer once voting has commenced.
@@
 \subsection{Proof Effort Remarks}
+
 The lemmas described in Section~\ref{sec:invariants} are constructed by-hand and represent multiple days of effort, but that effort includes both model and protocol construction and generalization as well as verification.

### 2016-12-07 doc/om-paper/paper.tex
merged conflict.
Fix screwup in the experiments table
Cleanup and making a little space in paper

@@
 \newboolean{submission}  %set to true for the submission version
-\setboolean{submission}{false}
-%\setboolean{submission}{true}
+%\setboolean{submission}{false}
+\setboolean{submission}{true}
 \ifthenelse
@@
 
-%\author{Benjamin F Jones and Lee Pike}
 \author{Benjamin F Jones\inst{1} \and Lee Pike\inst{2}}
@@
 \lee{Ben will redo. I'd just remove the Gurevich reference}
+\ben{I still need to do this!}
 
@@
 
-\begin{example}
-Consider a simple distributed system involving 3 processors: $S$, $L_1$, and $L_2$.
+% \begin{example}
+% Consider a simple distributed system involving 3 processors: $S$, $L_1$, and $L_2$.
 The purpose of the system is for $S$ to broadcast a message to $L_1$ and $L_2$ after which they perform a computation on the message and update their local state with the result.
 One might model this system as a transition system whose state is the product of the local processor states along with the state of the environment and any globally important phenomena such as time.
-
-\begin{figure}
-\centering
-\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3cm,
-  thick,main node/.style={circle,fill=blue!20,draw,
-  font=\sffamily\Large\bfseries,minimum size=8mm}]
-
-  \node[main node] (S) at (0,0) {$S$};
-  \node[main node] (L1) at (2,1) {$L_1$};
-  \node[main node] (L2) at (2,-1) {$L_2$};
-
-  \path[every node/.style={font=\sffamily\small,
-          fill=white,inner sep=1pt}]
-    (S)  edge (L1)
-         edge (L2);
-
-\end{tikzpicture}
-\label{fig:exa-trans}
-\caption{Simple message passing system}
-\end{figure}
+% 
+% \begin{figure}
+% \centering
+% \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3cm,
+%   thick,main node/.style={circle,fill=blue!20,draw,
+%   font=\sffamily\Large\bfseries,minimum size=8mm}]
+% 
+%   \node[main node] (S) at (0,0) {$S$};
+%   \node[main node] (L1) at (2,1) {$L_1$};
+%   \node[main node] (L2) at (2,-1) {$L_2$};
+% 
+%   \path[every node/.style={font=\sffamily\small,
+%           fill=white,inner sep=1pt}]
+%     (S)  edge (L1)
+%          edge (L2);
+% 
+% \end{tikzpicture}
+% \label{fig:exa-trans}
+% \caption{Simple message passing system}
+% \end{figure}
 
@@
 The union $A_1 \cup A_2 \cup A_3$ covers the set of reachable states of the system and defines an ASM that is useful for reasoning about its dynamics and understanding counter-example traces.
-\end{example}
 
-% ------------------------------------------------------------
+% \end{example}
 
-% ------------------------------------------------------------
 \section{Putting it Together: Compositional Modeling and Verification for Oral Messages}\label{sec:byz}
@@
   \begin{tabular}{c|c||c|c|c|c|c|c|c|c|c|c|}
-      \multicolumn{2-12}{c}{Receivers} \\
+      \multicolumn{1}{r}{} & \multicolumn{1}{r}{} & \multicolumn{10}{c}{Receivers} \\
       \cline{2-12}
@@
       \cline{2-12} \cline{2-12}
-      \multirow{10}{*}{Relays} &
+          \multirow{10}{*}{Relays\phantom{x}} &
             1   & 7    & 9   & 12   & 15  & 21   & 25   & 32   & 40  & 54  & 74  \\
@@
   \hline
-  Time-Triggered Messaging        & source, relays, receivers & 3 new & 2 new, 3 modified & calendar, faults \\
+  Time-Triggered Messaging        & source, relays, receivers, ASM & 3 new & 2 modified, 3 modified & calendar, faults \\
   \hline

### 2016-12-07 doc/om-paper/paper.tex
wibbles sitting with Ben.

@@
 \begin{itemize}
-\item {\bf $\OM(0)$}: $g$ broadcasts a value to each lieutenant, and the lieutenants return the value received (or $\ERR$).
-\item {\bf $\OM(m)$, $m > 0$}:
+\item {\bf $\OMH(0)$}: $g$ broadcasts a value to each lieutenant, and the lieutenants return the value received (or $\ERR$).
+\item {\bf $\OMH(m)$, $m > 0$}:
   \begin{enumerate}
@@
 $l$ sends $l_v$ or $R$, if $l_v = \ERR$.
-  \item For each lieutenant $l \in L$, remove all $\ERR$ values received in Step~\ref{om:two} from executing $OHM(m-1)$.
+  \item For each lieutenant $l \in L$, remove all $\ERR$ values received in Step~\ref{om:two} from executing $\OMH(m-1)$.
 Compute the majority value over the remaining values, or $V$ if there is no majority.
@@
 
-We have implemented $\OMH(1)$ (as well as many variants, described in Section~\ref{sec:modular}) in the Symbolic Analysis Laboratory~\cite{SAL} (SAL) language.
-The verification is also carried out in SAL, which contains a suite of model-checkers.
+We have implemented $\OMH(1)$ (as well as the variants, described in Section~\ref{sec:modular}) in Symbolic Analysis Laboratory (SAL)~\cite{SAL}.
+SAL contains a suite of model-checkers.
 In our work, we use infinite-state (SMT-based) $k$-induction~\cite{Moura_boundedmodel,Armando-et-al}.
 
-We have chosen to follow Rushby~\cite{Rushby:OM1} in ``unrolling'' the communication among lieutenants into two sets of logical nodes: relays and receivers.
-A special node called the source takes the place of the general above.
-Thus, for example, in the case where 4 nodes execute $\OMH(1)$, our model consists of one source, three relays, and three receivers.
+We follow Rushby~\cite{Rushby:OM1} in ``unrolling'' the communication among lieutenants into two sets of logical nodes: \emph{relays} and \emph{receivers}. Relays \lee{finish}
+
+We refer to the general as the \emph{source}.
+Thus, for example, in the case where four nodes execute $\OMH(1)$, our model consists of one source, three relays, and three receivers.
 We model communication through one-way, typed channels.

### 2016-12-07 doc/om-paper/paper.tex
down to 16 pgs.

@@
 \usepackage{multirow}
+\usepackage{multicol}
 
@@
 With this machinery, we can define the semantics of faults by constraining the relationship between a message broadcast and the values received by the recipients.
-For a nonfaulty process that broadcasts, every recipient receives the sent message.
+For a nonfaulty process that broadcasts, every recipient receives the sent message, and for symmetric faults, there is no requirement that the messages sent are the ones received, only that every recipient receives the same value:
+\noindent
+\begin{multicols}{2}
 \begin{align*}
@@
   &\quad\quad\quad \textnormal{ implies } recv(i, j, t) = send(i, j, t)
-\end{align*}
+\end{align*}\break
 \noindent
-For symmetric faults, there is no requirement that the messages sent are the ones received, only that every recipient receives the same value.
 \begin{align*}
@@
   &\quad \forall i, j, k \in \Id, t \in \mathbb{R}.\\
-  &\quad\quad (faults(i) = sym \textnormal{ and } broadcast(i, \{j, k\}, t, m))\\
-  &\quad\quad\quad \textnormal{ implies } recv(i, j, t) = recv(i, k, t)
+  &\quad\quad\quad (\phantom{\textnormal{and }} faults(i) = sym\\
+  &\quad\quad\quad \phantom{(}\textnormal{and } broadcast(i, \{j, k\}, t, m))\\
+  &\quad\quad \textnormal{ implies } recv(i, j, t) = recv(i, k, t)
 \end{align*}
+\end{multicols}
 \noindent
@@
 Validity states that if the general is nonfaulty, then every lieutenant outputs the value sent by the general.
-More formally, let the general broadcast $v \neq \ERR$, and let $l_i$ denote the output for lieutenant $i$:
-%
+Agreement states that each lieutenant outputs the same value.
+More formally, Let $l_i, l_j$ denote the outputs of lieutenants $i, j \in L$, respectively, and let $v$ be the value the general broadcasts:
+\noindent
+\begin{multicols}{2}
 \begin{equation}
   \tag{Validity}
-    \forall \,i. \quad l_i = v
-\end{equation}
-%
-Agreement states that each lieutenant outputs the same value. Let $l_i, l_j$ denote the outputs of lieutenants $i, j \in L$, respectively:
-%
+    \forall \,i. \quad v \neq \ERR \Rightarrow l_i = v
+\end{equation}\break
 \begin{equation}
@@
 \end{equation}
+\end{multicols}
+%
+
+%
 %
@@
 SAL contains a suite of model-checkers.
-In our work, we use infinite-state (SMT-based) $k$-induction~\cite{Moura_boundedmodel,Armando-et-al}.
+In our work, we use infinite-state (SMT-based) $k$-induction~\cite{}.
 

### 2016-12-07 doc/om-paper/paper.tex
various paper edits throughout.

@@
 
-\title{Modular Model-Checking for Fault-Tolerant Distributed System Implementations}
+\title{Modular Model-Checking Implementations of a Byzantine Fault-Tolerant Protocol}
 
-\author{Benjamin F Jones\inst{1} \and Lee Pike\inst{2}}
+\author{Benjamin F Jones\inst{1} \and Lee Pike\inst{1}}
 
-\institute{Galois, Inc., Portland OR 97204,\\
-\email{bjones@galois.com}
-\and
-Galois, Inc., Portland OR 97204,\\
-\email{leepike@galois.com}}
+\institute{Galois, Inc., Portland OR 97204\\
+\email{\{bjones,leepike\}@galois.com}
+%% \and
+%% Galois, Inc., Portland OR 97204,\\
+%% \email{leepike@galois.com}
+}
 
@@
 
-\lee{grep to get consistent manifest/benign, also lieutenant vs relay/receiver.}
 \lee{grep for synchronous observers section references}
-\lee{grep for first use of ``synchronous composition'' and give pointer to SAL semantics}
 
@@
 
-The verification of this class of systems has usually been at the algorithmic level, eliding details about a concrete implementation, and historically, has relied on formal models verified by interactive theorem-proving~\cite{om-acl2-impl,Young97:IC,csl-93-2,pvs}.
+The verification of this class of systems has usually been at the algorithmic level, eliding details about a concrete implementation, and historically, has relied on formal models verified by interactive theorem-proving~\cite{om-acl2-impl,Young97:IC,Lincoln-Rushby,pvs}.
 If formal verification is to be introduced into the workflow of system designers, though, we need more automated methods that scale for implementation-level models.
@@
 
-There are at least two classes of abstractions that separate traditional models of fault-tolerant distributed algorithms from their implementations.
+There are at least two classes of abstractions that separate protocol-level models of fault-tolerant distributed algorithms from their implementations.
 One is to intertwine the environmental model with the system description.
@@
 Another class of abstractions is used to simplify models.
-For example, message passing might be abstracted with shared state, or a node's local behavior is elided and instead, the output is constrained by a \emph{specification} of the behavior, e.g.
-pre- and post-conditions.
+For example, message passing might be abstracted with shared state, or a node's local behavior is elided and instead, the output is constrained by a \emph{specification} of the behavior.
 
-In this paper, we develop a formal framework to address these concerns.
+In this paper, we present a framework for modeling fault-tolerant distributed systems, and use that framework to verify Oral Messages Hybrid algorithm (\OMH)~\cite{Lincoln-Rushby}
+
+.
 We combine ideas from the literature and folklore to build scalable and modular formal models of fault-tolerant distributed systems.
@@
 To demonstrate the framework, in Section~\ref{sec:byz} focus on a particular representative algorithm as a case-study.
-We present a formally-verified implementation of the Oral Messages Hybrid algorithm (\OMH)~\cite{Lincoln-Rushby}, implementing the Boyer-Moore Fast Majority Vote algorithm for voting~\cite{Boyer-Moore:91}.
+We present a formally-verified implementation of the Oral Messages Hybrid algorithm (\OMH)~\cite{Lincoln-Rushby}, implementing the Boyer-Moore Fast Majority Vote algorithm for voting~\cite{mjrty}.
 The verification is interesting in its own right, as it is the first fully parametric (on the number of nodes) model-checked \emph{implementation} of the algorithm.
@@
 \subsection{Calendar Automata}\label{sec:calendar}
-\emph{Timed automata} are specialized formalisms for verifying real-time systems~\cite{alur}.
+%% \emph{Timed automata} are specialized formalisms for verifying real-time systems~\cite{alur}.
 Timed automata assume the existence of continuously-varying real-valued variables, known as \emph{real-time clocks}.
@@
 
-However, real-time system verification in general-purpose model-checkers requires an explicit formalism of real-time progression.
+Real-time system verification in general-purpose model-checkers requires an explicit formalism of real-time progression.
 Trying to encode real-time clocks directly is difficult; in particular, one must avoid Zeno's paradox in which no progress is made because state transitions simply update real-valued variables by an infinitesimally small amount.
 To avoid this problem, Dutetre and Sorea developed \emph{calendar automata}~\cite{cal}, which is itself inspired by event calendars used in discrete-event simulation.
+Rather than encoding ``how much time has passed since the last event'', it encodes ``how far into the future is the next scheduled event'', and a real-valued variable representing the current time is updated to the next event time.
 
@@
 
-The calendar automaton avoids the problems of encoding timed automata.
-Rather than encoding ``how much time has passed since the last event'', it encodes ``how far into the future is the next scheduled event'', and a real-valued variable representing the current time is updated to the next event time.
-Doing so avoids Zeno updates.
+%% The calendar automaton avoids the problems of encoding timed automata.
+%% Doing so avoids Zeno updates.
 
@@
 
-From the definitions, it follows that in every state, the timeouts are never in the past:
+From the definitions, it follows that in every state, the timeouts are never in the past, and that time is monotonic:
+
 \begin{lemma}[Future timeouts]\label{lem:ft}
@@
 
-\noindent
-Furthermore, time is monotonic:
 \begin{lemma}[Monotonic time]
@@
 
-In specifying a system, a \emph{clock} module is synchronously composed with the system. The clock module updates time to the next calendar event, and stutters otherwise.
+In specifying a system, a \emph{clock} module is synchronously composed (under SAL semantics for synchronous composition~\cite{SAL}) with the system.
+The clock module updates time to the next calendar event, and stutters otherwise.
 
@@
 Generally, nodes do not contain state variables assigned to their faults, or use their fault-status to determine their behavior!\footnote{There are exceptions.
-Benign faults may be detected by a node itself (e.g., in a built-in-test), an distributed diagnosis can let a node discover that others believe it is faulty~\cite{}.} The upshot is that combining faults and node state divorces the specification from its implementation.
+Benign faults may be detected by a node itself (e.g., in a built-in-test).} The upshot is that combining faults and node state divorces the specification from its implementation.
 
@@
 \paragraph{Transient faults.} If the $faults$ mapping is a constant, then faults are permanently but nondeterministically assigned to nodes.
-However, we can easily model \emph{transient faults}~\cite{} in which nodes are faulty temporarily by making $faults$ a state variable that is updated nondeterministically.
+However, we can easily model \emph{transient faults} in which nodes are faulty temporarily by making $faults$ a state variable that is updated nondeterministically.
 Whether we model permanent or transient faults, a \emph{maximum fault assumption} (MFA) describes the maximum number of faults permitted in the system.
@@
 This idea is an implementation of abstract state machines (ASMs).
-ASMs are a generalization of finite state machines originally developed by Gurevich to generalize state machines to operate over arbitrary data structures~\cite{asm}.
+ASMs are a generalization of finite state machines originally developed by Gurevich to generalize state machines to operate over arbitrary data structures~\cite{}.
 In our context, we require only a restricted form of ASMs known as a \emph{sequential small-step ASM}:
@@
 
-To demonstrate the model and techniques, we model an implementation of the Oral Messages Hybrid ($\OMH$) algorithm~\cite{csl-93-2}, a more sophisticated version of the classic Oral Messages ($\OM$) algorithm.
+To demonstrate the model and techniques, we model an implementation of the Oral Messages Hybrid ($\OMH$) algorithm~\cite{Lincoln-Rushby}, a more sophisticated version of the classic Oral Messages ($\OM$) algorithm.
 $\OM$ was first formulated by Lamport, Shostak, and Pease~\cite{om} to achieve agreement in a distributed system while tolerating Byzantine faults.
 $\OMH$ was originally developed by Thambidurai and Park~\cite{hybrid} to achieve the same result in the presence of a more sophisticated fault model that includes symmetric and benign faults.
-However, $\OMH$ had a bug, as originally formulated, which was corrected and the mended algorithm was formally verified by Lincoln and Rushby using interactive theorem-proving~\cite{csl-93-2}.
+However, $\OMH$ had a bug, as originally formulated, which was corrected and the mended algorithm was formally verified by Lincoln and Rushby using interactive theorem-proving~\cite{Lincoln-Rushby}.
 
@@
 $l$ sends $l_v$ or $R$, if $l_v = \ERR$.
-  \item For each lieutenant $l \in L$, remove all $\ERR$ values received in Step~\ref{om:two} from executing $\OMH(m-1)$.
+  \item\label{om:three} For each lieutenant $l \in L$, remove all $\ERR$ values received in Step~\ref{om:two} from executing $\OMH(m-1)$.
 Compute the majority value over the remaining values, or $V$ if there is no majority.
@@
 Under that fault model, agreement and validity hold if $2a+2s+b+1 \geq n$, where $n$ is the total number of nodes, $a$ is the number of Byzantine (or asymmetric) faults, $s$ is the number of symmetric faults, and $b$ is the number of benign faults.
-Additionally, it is required that the number of rounds $m$ be greater or equal to the number of Byzantine faults, $a$~\cite{csl-93-2,hybrid}.
+Additionally, it is required that the number of rounds $m$ be greater or equal to the number of Byzantine faults, $a$~\cite{Lincoln-Rushby,hybrid}.
 
@@
 
-We have implemented $\OMH(1)$ (as well as the variants, described in Section~\ref{sec:modular}) in Symbolic Analysis Laboratory (SAL)~\cite{SAL}.
+We have implemented $\OMH(1)$ (as well as the variants described in Section~\ref{sec:modular}) in Symbolic Analysis Laboratory (SAL)~\cite{SAL}.
 SAL contains a suite of model-checkers.
-In our work, we use infinite-state (SMT-based) $k$-induction~\cite{}.
+In our work, we use infinite-state (SMT-based) $k$-induction~\cite{cal}.
 
-We follow Rushby~\cite{Rushby:OM1} in ``unrolling'' the communication among lieutenants into two sets of logical nodes: \emph{relays} and \emph{receivers}. Relays \lee{finish}
-
+We follow Rushby~\cite{Rushby:OM1} in ``unrolling'' the communication among lieutenants into two sets of logical nodes: \emph{relays} and \emph{receivers}.
+Relays encode the lieutenants' Step~\ref{om:two} of the \OMH algorithm, in which they rebroadcast the values received from the general after filtering manifestly bad messages, while the receivers encode the voting step.
 We refer to the general as the \emph{source}.
-Thus, for example, in the case where four nodes execute $\OMH(1)$, our model consists of one source, three relays, and three receivers.
+The unrolling shows that a generalization of the original algorithm holds: the number of relays and receivers need not be the same.
 We model communication through one-way, typed channels.
@@
 
-Our SAL model defines 7 transition systems in total: \texttt{clock}, {\texttt{source}, \texttt{relay} (parametrized over an ID), \texttt{receiver} (parametrized over an ID), \texttt{abstractor}, \texttt{observer}, \texttt{abstract\_monitor}.
+Our SAL model defines seven transition systems in total: \texttt{clock}, {\texttt{source}, \texttt{relay} (parametrized over an ID), \texttt{receiver} (parametrized over an ID), \texttt{observer}, \texttt{abstractor}, and \texttt{abstract\_monitor}.
 The first four of these are composed asynchronously, in an intermediate system we label \texttt{system}, and share access to a global calendar consisting of event slots $(\text{message}, \text{time})$, one for each channel in the system.
-\texttt{clock} is responsible for updating a global variable \texttt{t} representing time according to the rules in Section~\ref{sec:calendar}.
-The asynchronous composition gives us a relaxed form of the original specification, but it is possible without much additional work to enforce timing constraints, see Section~\ref{sec:modular} "Time-Triggered Messaging".
-The remaining three systems are observers of the previously defined asynchronous system.
-The \texttt{observer} controls the state of \texttt{validity} and \texttt{agreement} variables, setting them to \texttt{FALSE} if, when the receivers have computed their vote, the properties do not hold.
+The \texttt{clock} transition system is responsible for updating a global variable \texttt{t} representing time according to the rules in Section~\ref{sec:calendar}.
+
+The asynchronous composition of the system relaxes the original specification of the algorithm considerably.
+For example, in our implementation, a receiver may receive a message from one relay before another relay has received a message from the source.
+We only require that all relays and receivers have executed before voting.
+With a general asynchronous model, it is easy to refine it further; for example, we refine it to a time-triggered model in Section~\ref{sec:modular}.
+
+The \texttt{observer} is a synchronous observer~\cite{Rushby:SAS14} that encodes the validity and agreement properties as state machines.
+State variables denoting validity and agreement are set to be false if the receivers have completed their vote but the respective properties do not hold.
+
+Finally, the \texttt{abstractor} and \texttt{abstract\_monitor} encode an abstract state machine observer for the system, as described in Section~\ref{sec:abstract}.
+
+%% controls the state of \texttt{validity} and \texttt{agreement} variables, setting them to \texttt{FALSE} if, when the receivers have computed their vote, the properties do not hold.
 The \texttt{abstractor} and \texttt{abstract\_monitor} monitor the state of \texttt{system} and update the values of a collection of 6 abstract state predicates.
@@
 Each proof incurs the full startup, parsing, type-checking, and model-generation time of SAL.
-
-Note that the theorems hold even in the degenerate cases of one relay or one receiver.
+Observe the theorems hold even in the degenerate cases of one relay or one receiver.
 
@@
   \centering
-  \begin{tabular}{|p{2.25cm}|p{2.25cm}|p{2.25cm}|p{2.25cm}|p{2.25cm}|}
+  \begin{tabular}{|p{3cm}|p{2cm}|p{2cm}|p{2cm}|p{2cm}|}
   \hline
-                                  & Transition systems (7 total)  &Definitions\linebreak (58 total) & Invariants\linebreak (11 total) & Invariant classes (5 total) \\
+                                  & Transition systems\linebreak (7 total)  &Definitions\linebreak (58 total) & Invariants\linebreak (11 total) & Invariant classes\linebreak (5 total) \\
   \hline \hline
@@
 Adding new kinds of faults requires more work but is still modular.
-Consider adding \emph{omissive asymmetric faults}, a restriction of Byzantine faults in which a broadcaster either sends the correct value or a manifest fault~\cite{}, to the fault model.
+Consider adding \emph{omissive asymmetric faults}, a restriction of Byzantine faults in which a broadcaster either sends the correct value or a benign fault~\cite{omissive}, to the fault model.
 Doing so requires modifying none of transition systems.
@@
 \paragraph{Time-Triggered Messaging.}
-A \emph{time-triggered} distributed system is one in which nodes are independently clocked, but clock constraints (ensured by periodic clock synchronization) allow the model to appear as if it is executing synchronously~\cite{}.
+A \emph{time-triggered} distributed system is one in which nodes are independently clocked, but clock constraints (ensured by periodic clock synchronization) allow the model to appear as if it is executing synchronously~\cite{kopetz}.
 
@@
 The lemmas described in Section~\ref{sec:invariants} are constructed by-hand and represent multiple days of effort, but that effort includes both model and protocol construction and generalization as well as verification.
-The counterexamples returned by SAL are very useful for strengthening invariants, but tedious to analyze---a model with five relays and two receivers contains 90 state variables, and there are known counterexamples to models that size~\cite{csl-93-2}.
+The counterexamples returned by SAL are very useful for strengthening invariants, but tedious to analyze---a model with five relays and two receivers contains 90 state variables, and there are known counterexamples to models that size~\cite{Lincoln-Rushby}.
 Once we developed the ASM synchronously-composed observer, the verification effort was sped up considerably.
@@
 A hybrid fault-model and modified Oral Messages algorithm was developed by Thambidurai and Park~\cite{hybrid}.
-Using the PVS interactive theorem-prover~\cite{pvs}, Lincoln and Rushby demonstrate a flaw in Thambidurai and Park's algorithm and present a fix to it~\cite{csl-93-2}.
+Using the PVS interactive theorem-prover~\cite{pvs}, Lincoln and Rushby demonstrate a flaw in Thambidurai and Park's algorithm and present a fix to it~\cite{Lincoln-Rushby}.
 
@@
 We hypothesize that an ADSL will make refactoring even easier, and we can generate invariants or invariant templates useful for verification.
-Indeed, we have developed a preliminary ADSL that generates C code as well as formal models SRI's Sally~\cite{}, a model-checking descendent to SAL that will be described in future work.\footnote{\url{https://github.com/GaloisInc/atom-sally}}
+Indeed, we have developed a preliminary ADSL that generates C code as well as formal models SRI's Sally~\cite{pdr-kind}, a model-checking descendent to SAL that will be described in future work.\footnote{\url{https://github.com/GaloisInc/atom-sally}}
 
@@
 While our model of node behavior is low-level, there are gaps.
-For example, our work is in SAL's language of guarded commands~\cite{} and needs to be either refined or verified to be equivalent to a software implementation's semantics.
+For example, our work is in SAL's language of guarded commands~\cite{SAL} and needs to be either refined or verified to be equivalent to a software implementation's semantics.
 Another aspect is that behavior related to networking and benign fault diagnosis (e.g., computing a CRC) is are left abstract.
@@
 We are indebted to our collaborators Brendan Hall and Srivatsan Varadarajan at Honeywell Labs, and to Wilfredo Torres-Pomales at NASA Langley for their discussions and insights.
+Additionally, we acknowledge that this work is heavily inspired by a series of papers authored by John Rushby.
 

### 2016-12-07 doc/om-paper/paper.tex
wip on ASM section

--- 2016-12-07 doc/om-paper/paper.tex
@@
 
-% ------------------------------------------------------------
-\subsection{Abstract State Machines Observers}\label{sec:abstract}\label{sec:asms}
+\subsection{Proof by Abstraction}\label{sec:abstract}\label{sec:asms}
 \lee{These are really observers. Let's call them abstract state machine observers and add a bit of the synchronous observers text here. Also, cite disjunctive invariants paper?}
@@
 In model-checking, counterexamples to properties are traces of states.
-A single state includes assignments to each of the variables, and the counterexample traces we examine (see Section~\ref{sec:experimental}) can contain hundreds of state variables.
-In practice, the essence of a counterexample depends on a small subset of these.
+A state is an assignment to each of the system variables and the counterexample traces we examine (see Section~\ref{sec:experimental}) can contain hundreds of system  variables.
+In practice however, the essence of a counterexample depends on a small subset of these.
 The user has to manually separate the state variables that matter from those that do not for a specific property.
@@
 
-But the relationship between many of the state variables often has an intuitive structure, one familiar to the designer, that can be characterized by a state machine imposed over it.
-The state machine describes discrete modes or configurations.
+But the relationship between many of the state variables often has an intuitive structure, one familiar to the designer, that can be characterized by an auxiliary state machine imposed over it.
+The state space of this auxiliary machine comprises discrete modes or configurations of the real system.
 Each mode captures a relation on a subset of the state variables, as well as the set of possible modes that follow.
 
-This idea is an implementation of abstract state machines (ASMs).
+\cite{timed-systems}
+\cite{Rushby00:CAV}
+
+% This idea is an implementation of abstract state machines (ASMs).
 ASMs are a generalization of finite state machines originally developed by Gurevich to generalize state machines to operate over arbitrary data structures~\cite{asm}.
-In our context, we require only a restricted form of ASMs known as a \emph{sequential small-step ASM}:
+In our modeling framework we take the ASM idea and turn it into both and auxiliary state machine and an observer of the real system as follows.
 
-\begin{definition}[Abstract state machine (ASM)]
+% In our context, we require only a restricted form of ASMs known as a \emph{sequential small-step ASM}:
 
-\lee{Ben will redo. I'd just remove the Gurevich reference}
-\ben{I still need to do this!}
+\begin{definition}[Abstract state machine observer (ASMO)]
 
@@
 \begin{itemize}
-\item {\bf $\OMH(0)$}: $g$ broadcasts a value to each lieutenant, and the lieutenants return the value received (or $\ERR$).
-\item {\bf $\OMH(m)$, $m > 0$}:
+\item {\bf $\OM(0)$}: $g$ broadcasts a value to each lieutenant, and the lieutenants return the value received (or $\ERR$).
+\item {\bf $\OM(m)$, $m > 0$}:
   \begin{enumerate}
@@
 $l$ sends $l_v$ or $R$, if $l_v = \ERR$.
-  \item For each lieutenant $l \in L$, remove all $\ERR$ values received in Step~\ref{om:two} from executing $\OMH(m-1)$.
+  \item For each lieutenant $l \in L$, remove all $\ERR$ values received in Step~\ref{om:two} from executing $OHM(m-1)$.
 Compute the majority value over the remaining values, or $V$ if there is no majority.
@@
 
-We have implemented $\OMH(1)$ (as well as the variants, described in Section~\ref{sec:modular}) in Symbolic Analysis Laboratory (SAL)~\cite{SAL}.
-SAL contains a suite of model-checkers.
+We have implemented $\OMH(1)$ (as well as many variants, described in Section~\ref{sec:modular}) in the Symbolic Analysis Laboratory~\cite{SAL} (SAL) language.
+The verification is also carried out in SAL, which contains a suite of model-checkers.
 In our work, we use infinite-state (SMT-based) $k$-induction~\cite{Moura_boundedmodel,Armando-et-al}.
 
-We follow Rushby~\cite{Rushby:OM1} in ``unrolling'' the communication among lieutenants into two sets of logical nodes: \emph{relays} and \emph{receivers}. Relays \lee{finish}
-
-We refer to the general as the \emph{source}.
-Thus, for example, in the case where four nodes execute $\OMH(1)$, our model consists of one source, three relays, and three receivers.
+We have chosen to follow Rushby~\cite{Rushby:OM1} in ``unrolling'' the communication among lieutenants into two sets of logical nodes: relays and receivers.
+A special node called the source takes the place of the general above.
+Thus, for example, in the case where 4 nodes execute $\OMH(1)$, our model consists of one source, three relays, and three receivers.
 We model communication through one-way, typed channels.

### 2016-12-07 doc/om-paper/paper.tex

--- 2016-12-07 doc/om-paper/paper.tex
@@
 
-But the relationship between many of the state variables often has an intuitive structure, one familiar to the designer, that can be characterized by a state machine imposed over it.
-The state machine describes discrete modes or configurations.
+But the relationship between many of the state variables often has an intuitive structure, one familiar to the designer, that can be characterized by an auxiliary state machine imposed over it.
+The state space of this auxiliary machine comprises discrete modes or configurations of the real system.
 Each mode captures a relation on a subset of the state variables, as well as the set of possible modes that follow.
 
-This idea is an implementation of abstract state machines (ASMs).
-ASMs are a generalization of finite state machines originally developed by Gurevich to generalize state machines to operate over arbitrary data structures~\cite{}.
-In our context, we require only a restricted form of ASMs known as a \emph{sequential small-step ASM}:
+\cite{timed-systems}
+\cite{Rushby00:CAV}
 
-\begin{definition}[Abstract state machine (ASM)]
+% This idea is an implementation of abstract state machines (ASMs).
+ASMs are a generalization of finite state machines originally developed by Gurevich to generalize state machines to operate over arbitrary data structures~\cite{asm}.
+In our modeling framework we take the ASM idea and turn it into both and auxiliary state machine and an observer of the real system as follows.
 
-\lee{Ben will redo. I'd just remove the Gurevich reference}
-\ben{I still need to do this!}
+% In our context, we require only a restricted form of ASMs known as a \emph{sequential small-step ASM}:
+
+\begin{definition}[Abstract state machine observer (ASMO)]
 

### 2016-12-07 doc/om-paper/paper.tex
intro changes.

@@
 abstractions come at the price of divorcing the model from its implementation
-and making refactoring difficult. In this work, we present a systematic model for
-fault-tolerant distributed system verification that combines calendar automata,
-symbolic fault injection, and abstract state machines observers specialized for
-infinite-state model-checking. We demonstrate the efficacy of the framework by presenting the
-first model-checked \emph{implementation} of the Hybrid Oral Messages algorithm,
-using a fast majority vote, then demonstrate how modifications along axes such
-as the fault model, timing constraints, and local node behavior require only local modifications
-to the model and proofs. This work is carried out in SAL model-checker.
+and making refactoring difficult. In this work, we present a framework for
+fault-tolerant distributed system verification that combining ideas from the literature including calendar automata,
+symbolic fault injection, and abstract state machines observers, and then we apply the framework to model-checking various implementations of the Hybrid Oral Messages algorithm that differ in the fault model, timing model, and local node behavior.
+We show that despite being low-level models, the verifications are scalable and modular, insofar as isolated changes to an implementation require modular changes to the model and proofs.
+This work is carried out in SAL model-checker.
 \end{abstract}
@@
 
-In this paper, we present a framework for modeling fault-tolerant distributed systems, and use that framework to verify Oral Messages Hybrid algorithm (\OMH)~\cite{Lincoln-Rushby}
+In this paper, we present a framework for modeling fault-tolerant distributed systems, and use that framework to verify several implementations of the Byzantine fault-tolerant Hybrid Oral Messages algorithm (\OMH)~\cite{Lincoln-Rushby}.
+The framework combines ideas from the literature and folklore to build scalable and modular formal models suitable for infinite-state model-checking, and it reduces the need for ad-hoc abstractions and optimizations while maintaining scalability.
+In Section~\ref{sec:model}, we present the important aspects of the framework, including \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal}, symbolic fault-injection, and abstract state machines for verification~\cite{}.
 
-.
-We combine ideas from the literature and folklore to build scalable and modular formal models of fault-tolerant distributed systems.
-We describe a generic formal framework for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}.
-There are three aspects of the framework: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) symbolic fault injection for decomposing the effects of faults from the system specification, and (3) \emph{abstract state machines observers}, useful for abstracting counterexamples.
+ %% There are three aspects of the framework: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) symbolic fault injection for decomposing the effects of faults from the system specification, and (3) \emph{abstract state machines observers}, useful for abstracting counterexamples.
 The framework is suitable for infinite-state model-checking supported by satisfiability modulo theories (SMT).
 
-To demonstrate the framework, in Section~\ref{sec:byz} focus on a particular representative algorithm as a case-study.
-We present a formally-verified implementation of the Oral Messages Hybrid algorithm (\OMH)~\cite{Lincoln-Rushby}, implementing the Boyer-Moore Fast Majority Vote algorithm for voting~\cite{mjrty}.
+We use the framework to verify low-level models of \OMH in which message passing is explicit, nodes are not enforced to execute strictly synchronously, and voting is explicit.
+In short, the models corresponds closely with an implementation of the algorithm.
+In Section~\ref{sec:byz}, we describe \OMH, an initial implementation model of it using the Boyer-Moore Fast Majority Vote algorithm~\cite{mjrty}, and a class of modular invariants; that is, invariants are grouped in a small set of classes with well-defined relationships.
 The verification is interesting in its own right, as it is the first fully parametric (on the number of nodes) model-checked \emph{implementation} of the algorithm.
 
-In Section~\ref{sec:experimental}, we give describe two sets of experiments on the model.
-First, we benchmark the model, showing that despite the additional complexity, it scales farther than other model-checking verifications at the algorithmic level.
-The scalability is due principally to specifying invariants and using $k$-induction in SRI's SAL model-checker~\cite{SAL}.
-Developing invariants requires some user guidance.
-A user should not have to discover all new invariants for new protocols or implementations.
-We show that most of the lemmas required are not protocol specific and are modular, concerning only specific aspects of the system, such as the timing, faults, or local node behavior.
-To demonstrate this, we modify the OHM implementation and show that in each case, only a small number of localized modifications are required to repair the model and proofs.
-The modifications we study are to add omissive-asymetric faults, make the implementation time-triggered, and replace the majority vote with a mid-value selection.
+In Section~\ref{sec:experimental}, we first show that despite being low-level, the model is scalable.
+Developing invariants requires some user guidance, and isolated changes to an implementation should require isolated modifications to the model and proof.
+To demonstrate this, we modify the \OMH implementation along the dimensions of faults (by adding an omissive-asymetric fault type~\cite{omissive}), time (by making a time-triggered model), and local behavior (by changing the majority vote to a mid-value selection) and show that in each case, the modifications are small and modular.
 
@@
 
-\section{Putting it Together: Compositional Modeling and Verification for Oral Messages}\label{sec:byz}
+\section{Modeling and Verification for Oral Messages}\label{sec:byz}
 
-To demonstrate the model and techniques, we model an implementation of the Oral Messages Hybrid ($\OMH$) algorithm~\cite{Lincoln-Rushby}, a more sophisticated version of the classic Oral Messages ($\OM$) algorithm.
-$\OM$ was first formulated by Lamport, Shostak, and Pease~\cite{om} to achieve agreement in a distributed system while tolerating Byzantine faults.
-$\OMH$ was originally developed by Thambidurai and Park~\cite{hybrid} to achieve the same result in the presence of a more sophisticated fault model that includes symmetric and benign faults.
+The Hybrid Oral Messages ($\OMH$) algorithm~\cite{Lincoln-Rushby} is a variant of the classic Oral Messages ($\OM$) algorithm~\cite{om}, originally developed by Thambidurai and Park~\cite{hybrid} to achieve distributed consensus in the presence of a hybrid fault model.
 However, $\OMH$ had a bug, as originally formulated, which was corrected and the mended algorithm was formally verified by Lincoln and Rushby using interactive theorem-proving~\cite{Lincoln-Rushby}.
@@
 \subsection{Model Sketch}\label{sec:sketch}
+
+\lee{we need to mention MJRTY here or in the modular verifiation section}
 

### 2016-12-07 doc/om-paper/paper.tex
more edits!

@@
 \usepackage{xcolor}
-
+\usepackage{xspace}
 \usepackage{graphicx}
@@
 
-\newcommand{\OM}[1]{\ensuremath{\mathrm{OM}(#1)}}
-\newcommand{\OMH}{\ensuremath{\mathrm{OMH}}}
-\newcommand{\Id}{\ensuremath{\mathrm{Id}}}
-\newcommand{\Msg}{\ensuremath{\mathrm{Msg}}}
-\newcommand{\ERR}{\ensuremath{\mathrm{ERR}}}
-\newcommand{\TO}{\ensuremath{\bm{\mathrm{T}}}}
+\newcommand{\OM}[1]{\ensuremath{\mathrm{OM}(#1)}\xspace}
+\newcommand{\OMH}{\ensuremath{\mathrm{OMH}}\xspace}
+\newcommand{\Id}{\ensuremath{\mathrm{Id}}\xspace}
+\newcommand{\Msg}{\ensuremath{\mathrm{Msg}}\xspace}
+\newcommand{\ERR}{\ensuremath{\mathrm{ERR}}\xspace}
+\newcommand{\TO}{\ensuremath{\bm{\mathrm{T}}}\xspace}
 
@@
 \noindent
-\begin{multicols}{2}
-\begin{align*}
+
+\begin{center}
+\begin{tabular}{l|l}
+\begin{minipage}[t]{0.5\linewidth}
+\[\begin{aligned}
 &nonfaulty\_constraint =\\
   &\quad \forall i, j \in \Id, t \in \mathbb{R}.\\
-  &\quad\quad faults(i) = none\\
-  &\quad\quad\quad \textnormal{ implies } recv(i, j, t) = send(i, j, t)
-\end{align*}\break
-\noindent
-\begin{align*}
+  &\quad\quad\quad faults(i) = none\\
+  &\quad\quad \textnormal{ implies } recv(i, j, t) = send(i, j, t)
+\end{aligned}\]
+\end{minipage}&
+\begin{minipage}[t]{0.5\linewidth}
+\[\begin{aligned}
 &sym\_constraint =\\
@@
   &\quad\quad \textnormal{ implies } recv(i, j, t) = recv(i, k, t)
-\end{align*}
-\end{multicols}
+\end{aligned}\]
+\end{minipage}
+\end{tabular}
+\end{center}
+
 \noindent
@@
 
-As a point of comparison, we also benchmarked Rushby's model~\cite{Rushby:OM1} on the same machine, using the same environment.
-When the number of receivers was fixed at 2, the verification time for increasing numbers of relays was: $0.12 s$, $0.20 s$, $2.56 s$, $8.24 s$, $28.0 s$, $549 s$, \TO, \TO, \TO (compare to the numbers in the second column of Figure~\ref{fig:benchmark}).
-Empirical evidence suggests that our verification, while slower for small numbers of nodes, scales much better for larger numbers of nodes.
+As a point of comparison, Rushby presents an elegant model of the simpler \OM algorithm, also in SAL, but using BDD-based model-checking~\cite{Rushby:OM1}.
+The model is at the algorithmic level.
+For small values, the verification is much faster, probably due to making only one call to SAL.
+However, for six relays and two receivers, it takes 449~seconds and timeouts for seven relays and two receivers.
 
@@
 We report on four aspects of the system: which transition systems making up the full model are modified (as described in Section~\ref{sec:model}), how many definitions have to be added or modified, the number of invariants that have to be added or modified, and which invariant classes (as defined in Section~\ref{sec:invariants}) those lemmas belong to.
+We modify the implementation along the axes of faults, time, and local node behavior.
+
+\lee{no paragraph on no faults in table?}
 
@@
 Two new invariants calendar invariants are introduced, stating that the calendar messages are either empty, or their time-stamps fall within the respective message windows.
-Then, three invariants classifying faults are relaxed to allow for the possibility of faulty nodes sending manifestly-faulty messages.
+Then, three invariants classifying faults are relaxed to allow for the possibility of faulty nodes sending benign messages.
 
@@
 
-The Oral Messages algorithm and its variants have a long history of formal verification, making it a good test-case for comparing and contrasting verification tools~\cite{pvs}.
-The Oral Messages algorithm was presented in the landmark paper by Lamport, Shostak, and Pease~\cite{om}.
-OM(1) was subsequently verified in both the PVS and ACL2 interactive theorem-provers~\cite{Young97:IC}.
-Also in ACL2, an implementation of a circuit design to implement OM(1) is given~\cite{om-acl2-impl}; the low-level model most closely relates to our level of detail.
-A refinement-based verification approach is used, and OM(1) is specialized to a four-node system.
+The Oral Messages algorithm and its variants and its variants have a long history of formal verification.
+\OM{1} was verified in both the PVS and ACL2 interactive theorem-provers~\cite{Young97:IC}.
+Also in ACL2, an implementation of a circuit design to implement \OM{1} is given~\cite{om-acl2-impl}; the low-level model most closely relates to our level of detail.
+A refinement-based verification approach is used, and \OM{1} is specialized to a four-node system.
 It appears that a general purpose voting algorithm (like Fast MJRTY) is not verified as part of the work.
@@
 The model is efficient for up to five nodes, but results are not presented beyond that.
-Very recently, Jovanovi{\'{c}} and Dutertre use a ``flattened'' high-level model of OM(1) as a benchmark for IC3 augmented with $k$-induction~\cite{pdr-kind}.
+Very recently, Jovanovi{\'{c}} and Dutertre use a ``flattened'' high-level model of \OM{1} as a benchmark for IC3 augmented with $k$-induction~\cite{pdr-kind}.
 
@@
 We hypothesize that an ADSL will make refactoring even easier, and we can generate invariants or invariant templates useful for verification.
-Indeed, we have developed a preliminary ADSL that generates C code as well as formal models SRI's Sally~\cite{pdr-kind}, a model-checking descendent to SAL that will be described in future work.\footnote{\url{https://github.com/GaloisInc/atom-sally}}
+Indeed, we have developed a preliminary ADSL that generates C code as well as formal models SRI's Sally~\cite{pdr-kind}, to be described in future work.\footnote{\url{https://github.com/GaloisInc/atom-sally}}
 

### 2016-12-08 doc/om-paper/paper.tex
Rewrote section 2.3 on abstract transition systems

--- 2016-12-07 doc/om-paper/paper.tex
@@
 \newcommand{\TO}{\ensuremath{\bm{\mathrm{T}}}}
+%\newcommand{\tld}[1]{\overset{\sim}{#1}}
+\newcommand{\tld}[1]{#1_{*}}
 
@@
 
-\subsection{Proof by Abstraction}\label{sec:abstract}\label{sec:asms}
-\lee{These are really observers. Let's call them abstract state machine observers and add a bit of the synchronous observers text here. Also, cite disjunctive invariants paper?}
+\subsection{Abstract Transition Systems}\label{sec:abstract}
 
-In model-checking, counterexamples to properties are traces of states.
+In order to scale verification of models up to more than a few nodes, and to avoid having to increase the depth of induction (which tends to be very expensive) during verification, we have followed Rushby's  application of abstract transition systems (and disjunctive invariants) \cite{Rushby00:CAV,timed-systems}.
+In this context, an abstract transition system, relative to a given transition system $\mathcal{M} = (S, I, \rightarrow)$, is a set of state predicates $A_1, \ldots, A_n$ over $S$ and a transition system $\tld{\mathcal{M}} = (\tld{S}, \tld{I}, \leadsto)$ such that:
+
+\begin{enumerate}
+    \item $\tld{S} = \{a_1, \ldots, a_n\}$ is a set of ``abstract states'' which correspond one-to-one with the state predicates $A_i$.
+    \item $\forall s \in I, \, \exists i : \, a_i \in \tld{I} \wedge A_i(s)$
+    \item $\forall a_i \in \tld{S} \, \forall \sigma, \sigma' : \, A_i(\sigma) \wedge \sigma \rightarrow \sigma' \, \implies \, A_{k_1}(\sigma') \vee \ldots \vee A_{k_m}(\sigma')$ where $\{a_{k_1}, \ldots, a_{k_m}\}$ are the abstract states to which $\tld{\mathcal{M}}$ may transition from $a_i$.
+\end{enumerate}
+
+For verification purposes, it is important to note that if $\mathcal{M}$ and $\tld{\mathcal{M}}$ satisfy the requirements above, then $A_1 \vee \ldots \vee A_n$ is an inductive invariant of $\mathcal{M}$.
+We may use such an invariant freely as a powerful assumption in the proof of other invariants (see Section~\ref{sec:invariants}).
+
+The use of abstract transition systems not only allows us to scale proofs farther, but also to improve traceability and debugging while developing a model.
+In models like the ones described in Section~\ref{sec:modular} where there are on the order of 100 state variables and counter-example traces could be 30 steps long, the designer can be easily lost trying to identify the essence.
+In such cases, the values of the abstract predicates can serve to focus the designer's attention on one particular mode of the system where the counter example is taking place.
+
+% In model-checking, counterexamples to properties are traces of states.
 A state is an assignment to each of the system variables and the counterexample traces we examine (see Section~\ref{sec:experimental}) can contain hundreds of system  variables.
@@
 The user has to manually separate the state variables that matter from those that do not for a specific property.
-Moreover, It is not enough to prevent variables that stutter from displaying since some state variables might be updated without affecting the property.
 The problem is particularly pertinent when formal models have sufficient detail for implementation synthesis, as is the goal in our work.
@@
 
-But the relationship between many of the state variables often has an intuitive structure, one familiar to the designer, that can be characterized by an auxiliary state machine imposed over it.
+% But the relationship between many of the state variables often has an intuitive structure, one familiar to the designer, that can be characterized by an auxiliary state machine imposed over it.
 The state space of this auxiliary machine comprises discrete modes or configurations of the real system.
 Each mode captures a relation on a subset of the state variables, as well as the set of possible modes that follow.
-
-\cite{timed-systems}
-\cite{Rushby00:CAV}
 
@@
 
-\begin{definition}[Abstract state machine observer (ASMO)]
-
-Let $\mathcal{M} = (S, I, \rightarrow)$ be a transition system and let $B \subseteq 2^S$ be a set of predicates over $S$.
+% \begin{definition}[Abstract state machine observer (ASMO)]
+% 
+% Let $\mathcal{M} = (S, I, \rightarrow)$ be a transition system and let $B \subseteq 2^S$ be a set of predicates over $S$.
 Then $\mathcal{M}$ is an ASM iff the transition relation $\rightarrow$ is of the form
-$$\textnormal{if } b_0(s_0) \textnormal{ then } s'_0 \textnormal{ else if } b_1(s_1) \textnormal{ then } s'_1 \ldots \textnormal{ else } b_n(s_n) \textnormal{ then } s'_n$$
-\noindent
-where $b_0, b_1, \ldots, b_n$ are arbitrary Boolean combinations of the predicates in $B$.
-\end{definition}
-
-\ben{I'm confused by this definition, but I haven't looked at the reference, so maybe it would be clear if I did.
-My intuitive notion of ASM involves a set of predicates B which *cover* the reachable states and a transition relation over B that abstracts the one over S.}
-
-Our use of ASMs to improve traceability is even more restrictive.
-We construct an ASM $\mathcal{M}$ that is synchronously composed with the system specification $\mathcal{S}$.
-Each state of the ASM represents a ``configurations'' of $S$ and is associated with some predicate over the states of $\mathcal{S}$.
+% $$\textnormal{if } b_0(s_0) \textnormal{ then } s'_0 \textnormal{ else if } b_1(s_1) \textnormal{ then } s'_1 \ldots \textnormal{ else } b_n(s_n) \textnormal{ then } s'_n$$
+% \noindent
+% where $b_0, b_1, \ldots, b_n$ are arbitrary Boolean combinations of the predicates in $B$.
+% \end{definition}
 
 % \begin{example}
+
 % Consider a simple distributed system involving 3 processors: $S$, $L_1$, and $L_2$.
@@
 
-Intuitively, there are subsets of the overall state space which are important to understanding the evolution of such a system.
+% Intuitively, there are subsets of the overall state space which are important to understanding the evolution of such a system.
 For example, let $A_1$ denote the subset of states in which the processor $S$ has not sent a message, let $A_2$ be the subset in which $S$ has broadcast a message, and let $A_3$ be the subset in which $L_1$ and $L_2$ have each received and processed the message.

### 2016-12-08 doc/om-paper/paper.tex

--- 2016-12-07 doc/om-paper/paper.tex
@@
 \newcommand{\TO}{\ensuremath{\bm{\mathrm{T}}}\xspace}
+\newcommand{\tld}[1]{#1_{*}\xspace}
 

### 2016-12-08 doc/om-paper/paper.tex
more wibbles.

@@
   \hline \hline
-  No Fault Model                  & receivers & 8 modified & 2 modified & ASM, faults \\
-  \hline
-  Omissive Asymmetric Faults       & none      & 4 modified & 2 modified & faults             \\
+  %% No Fault Model                  & receivers & 8 modified & 2 modified & ASM, faults \\
+  %% \hline
+  Omissive Asymmetric Faults       & none      & 1 new, 2 modified & 2 modified & faults             \\
   \hline
@@
 
-\lee{no paragraph on no faults in table?}
-
 \paragraph{Omissive Asymmetric Faults.}
@@
 Doing so requires modifying none of transition systems.
-However, we modify four definitions (\texttt{msg()}, characterizing the result on the calendar due to a faulty message; \texttt{omissive\_asymmetric\_fault\_constraint()}, defining the fault, and two functions that characterize the overall maximum fault assumption).
+We add a new uninterpreted function definition for omissive asymmetric faults, then modify the type of faults, and their effect on the calendar.
 Two invariants, both in the class of invariants cover faults, are extended to cover the cases where a sender is omissive asymmetric.
@@
 It appears that a general purpose voting algorithm (like Fast MJRTY) is not verified as part of the work.
-A hybrid fault-model and modified Oral Messages algorithm was developed by Thambidurai and Park~\cite{hybrid}.
-Using the PVS interactive theorem-prover~\cite{pvs}, Lincoln and Rushby demonstrate a flaw in Thambidurai and Park's algorithm and present a fix to it~\cite{Lincoln-Rushby}.
+It is not clear how difficult it would be to model and verify changes to the implementation in that framework.
 
@@
 Another aspect is that behavior related to networking and benign fault diagnosis (e.g., computing a CRC) is are left abstract.
-In a software implementation, the \texttt{send()} and \texttt{recv()} functions would have to implement these behaviors.
-Finally, we have modeled messages as integers; but real data may be more complex.
-A synthesis approach may be suitable and has been used in the past for fault-tolerant clock-synchronization~\cite{ddd}.
+In a software implementation, the $send$ and $recv$ functions would have to implement these behaviors.
+
+ %% A synthesis approach may be suitable and has been used in the past for fault-tolerant clock-synchronization~\cite{ddd}.
 

### 2016-12-08 doc/om-paper/paper.tex
edits throughout.

@@
 
-\title{Modular Model-Checking Implementations of a Byzantine Fault-Tolerant Protocol}
+\title{Modular Model-Checking of a Byzantine Fault-Tolerant Protocol}
 
@@
 than ever before.  Still, fault-tolerant distributed systems are particularly
-challenging to model-check given their large state space and
+challenging to model-check given their large state spaces and
 non-determinism. The typical approach to controlling complexity is to construct
@@
 symbolic fault injection, and abstract state machines observers, and then we apply the framework to model-checking various implementations of the Hybrid Oral Messages algorithm that differ in the fault model, timing model, and local node behavior.
-We show that despite being low-level models, the verifications are scalable and modular, insofar as isolated changes to an implementation require modular changes to the model and proofs.
+We show that despite being implementation-level models, the verifications are scalable and modular, insofar as isolated changes to an implementation require modular changes to the model and proofs.
 This work is carried out in SAL model-checker.
@@
 \lee{grep for synchronous observers section references}
+\lee{grep for ASM vs abstract transition system}
 
@@
 
-In this paper, we present a framework for modeling fault-tolerant distributed systems, and use that framework to verify several implementations of the Byzantine fault-tolerant Hybrid Oral Messages algorithm (\OMH)~\cite{Lincoln-Rushby}.
+In this paper, we present a framework for modeling fault-tolerant distributed systems, and use that framework to verify several variant implementations of the Byzantine fault-tolerant Hybrid Oral Messages algorithm (\OMH)~\cite{Lincoln-Rushby}.
 The framework combines ideas from the literature and folklore to build scalable and modular formal models suitable for infinite-state model-checking, and it reduces the need for ad-hoc abstractions and optimizations while maintaining scalability.
-In Section~\ref{sec:model}, we present the important aspects of the framework, including \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal}, symbolic fault-injection, and abstract state machines for verification~\cite{}.
+In Section~\ref{sec:model}, we present the important aspects of the framework, including calendar automata, originally developed by Dutertre and Sorea~\cite{cal}, symbolic fault-injection, and abstract state machines for verification.
 
@@
 
-We use the framework to verify low-level models of \OMH in which message passing is explicit, nodes are not enforced to execute strictly synchronously, and voting is explicit.
+We use the framework to verify implementation-level models of \OMH in which message passing is explicit, nodes are not enforced to execute strictly synchronously, and voting is explicit.
 In short, the models corresponds closely with an implementation of the algorithm.
-In Section~\ref{sec:byz}, we describe \OMH, an initial implementation model of it using the Boyer-Moore Fast Majority Vote algorithm~\cite{mjrty}, and a class of modular invariants; that is, invariants are grouped in a small set of classes with well-defined relationships.
+In Section~\ref{sec:byz}, we first describe \OMH, then an implementation of it that uses the Boyer-Moore Fast Majority Vote algorithm~\cite{mjrty}.
+We then describe a set of modular invariants, such that the invariants only concern specific aspects of the model (e.g., faults, local node behavior, or the passage of time).
 The verification is interesting in its own right, as it is the first fully parametric (on the number of nodes) model-checked \emph{implementation} of the algorithm.
 
-In Section~\ref{sec:experimental}, we first show that despite being low-level, the model is scalable.
+In Section~\ref{sec:experimental}, we first show that despite being implementation-level, the model is scalable.
 Developing invariants requires some user guidance, and isolated changes to an implementation should require isolated modifications to the model and proof.
@@
 
-Proofs of these two lemmas are straightforward so we omit them here.
+Proofs of these two lemmas are straightforward and omitted.
 
@@
 The typical approach to modeling faults is to add new state variables to each process representing its fault state.
-Then additional state variables are added to represent the non-determinism introduced by faults into the state or messages by a process.
+Then a node chooses actions based on its fault state.
 As a simple example, we might define a node that sends a good message if it is non-faulty and a bad message otherwise.
@@
 \noindent
-But this typical approach mixes the specification of a node's behavior with the fault model, an aspect of the environment.
-Generally, nodes do not contain state variables assigned to their faults, or use their fault-status to determine their behavior!\footnote{There are exceptions.
-Benign faults may be detected by a node itself (e.g., in a built-in-test).} The upshot is that combining faults and node state divorces the specification from its implementation.
+But this approach mixes the specification of a node's behavior with the fault model, an aspect of the environment.
+Generally, nodes do not contain state variables assigned to their faults, or use their fault-status to determine their behavior!\footnote{There are exceptions; for example, benign faults may be detected by a node itself (e.g., in a built-in-test).} The upshot is that combining faults and node state divorces the specification from its implementation.
 
@@
 
-In order to scale verification of models up to more than a few nodes, and to avoid having to increase the depth of induction (which tends to be very expensive) during verification, we have followed Rushby's  application of abstract transition systems (and disjunctive invariants) \cite{Rushby00:CAV,timed-systems}.
+Due to the sheer size of implementation-level models, manually examining counterexamples is tedious.
+To scale up verification, we use abstract transition systems (also known as disjunctive invariants)~\cite{Rushby00:CAV,timed-systems}.
 In this context, an abstract transition system, relative to a given transition system $\mathcal{M} = (S, I, \rightarrow)$, is a set of state predicates $A_1, \ldots, A_n$ over $S$ and a transition system $\tld{\mathcal{M}} = (\tld{S}, \tld{I}, \leadsto)$ such that:
@@
 
-The Hybrid Oral Messages ($\OMH$) algorithm~\cite{Lincoln-Rushby} is a variant of the classic Oral Messages ($\OM$) algorithm~\cite{om}, originally developed by Thambidurai and Park~\cite{hybrid} to achieve distributed consensus in the presence of a hybrid fault model.
+The Hybrid Oral Messages ($\OMH$) algorithm~\cite{Lincoln-Rushby} is a variant of the classic Oral Messages ($\OM{m}$) algorithm~\cite{om}, originally developed by Thambidurai and Park~\cite{hybrid} to achieve distributed consensus in the presence of a hybrid fault model.
 However, $\OMH$ had a bug, as originally formulated, which was corrected and the mended algorithm was formally verified by Lincoln and Rushby using interactive theorem-proving~\cite{Lincoln-Rushby}.
@@
 Under that fault model, agreement and validity hold if $2a+2s+b+1 \geq n$, where $n$ is the total number of nodes, $a$ is the number of Byzantine (or asymmetric) faults, $s$ is the number of symmetric faults, and $b$ is the number of benign faults.
-Additionally, it is required that the number of rounds $m$ be greater or equal to the number of Byzantine faults, $a$~\cite{Lincoln-Rushby,hybrid}.
+Additionally, the number of rounds $m$ must be greater or equal to the number of Byzantine faults, $a$~\cite{Lincoln-Rushby,hybrid}.
 
@@
 
-We have implemented $\OMH(1)$ (as well as the variants described in Section~\ref{sec:modular}) in Symbolic Analysis Laboratory (SAL)~\cite{SAL}.
+We have implemented $\OMH(1)$ (as well as the variants described in Section~\ref{sec:modular}) in the Symbolic Analysis Laboratory (SAL)~\cite{SAL}.
 SAL contains a suite of model-checkers.
@@
   \item \emph{Abstract state machine (ASM)}: Lemmas relating the ASM synchronous observer states to the implementation states.
-  \item \emph{Lieutenant local state-machine (SM) behavior}: Lemmas describing the modes of behavior of the lieutenants.
+  \item \emph{Receiver local behavior}: Lemmas describing the modes of behavior of the receivers.
 The major modes of the state machine are receiving messages, then once it has filled its buffer, it votes, and after voting returns the result.
@@
 Arrows denote dependencies.
-For example, the ASM lemmas depend on both the calendar automata and lieutenant state-machine lemmas.
+For example, the ASM lemmas depend on both the calendar automata and receiver state-machine lemmas.
 As can be seen, the proof structure is modular.
 The calendar lemmas are general and independent of any particular protocol or fault model.
-Similarly, lemmas about the internal behavior of a lieutenant is independent of the global protocol behavior.
-It is also independent of the effect of faults on the system---the only ``knowledge'' of faults that lieutenant has is whether a fault is benign or not.
+Similarly, lemmas about the internal behavior of a receiver is independent of the global protocol behavior.
+It is also independent of the effect of faults on the system---the only ``knowledge'' of faults that receiver has is whether a fault is benign or not.
 Lemmas about the behavior of faults in the system are also independent of the particular protocol being modeled.
-Likewise, lemmas about the particular voting algorithm used depend only on the lieutenant's internal behavior.
+Likewise, lemmas about the particular voting algorithm used depend only on the receiver's internal behavior.
 Only the ASM depends on both calendar-specific and local state-machine results, since it is an abstraction of the entire system implementation.
@@
 
-As a point of comparison, Rushby presents an elegant model of the simpler \OM algorithm, also in SAL, but using BDD-based model-checking~\cite{Rushby:OM1}.
+As a point of comparison, Rushby presents an elegant model of the simpler \OM{m} algorithm, also in SAL, but using BDD-based model-checking~\cite{Rushby:OM1}.
 The model is at the algorithmic level.
 For small values, the verification is much faster, probably due to making only one call to SAL.
-However, for six relays and two receivers, it takes 449~seconds and timeouts for seven relays and two receivers.
+However, for six relays and two receivers, it takes 449~seconds and timeouts (at one hour) for seven relays and two receivers.
 
@@
 In the table, for each modification, we report how much of the model must be modified.
-We report on four aspects of the system: which transition systems making up the full model are modified (as described in Section~\ref{sec:model}), how many definitions have to be added or modified, the number of invariants that have to be added or modified, and which invariant classes (as defined in Section~\ref{sec:invariants}) those lemmas belong to.
+We report on four aspects of the system: which transition systems are modified (as described in Section~\ref{sec:sketch}), how many definitions have to be added or modified, the number of invariants that have to be added or modified, and which invariant classes (as defined in Section~\ref{sec:invariants}) those lemmas belong to.
 We modify the implementation along the axes of faults, time, and local node behavior.
@@
 Consider adding \emph{omissive asymmetric faults}, a restriction of Byzantine faults in which a broadcaster either sends the correct value or a benign fault~\cite{omissive}, to the fault model.
-Doing so requires modifying none of transition systems.
+Doing so requires modifying none of transition systems, because of the synchronous kibitzer.
 We add a new uninterpreted function definition for omissive asymmetric faults, then modify the type of faults, and their effect on the calendar.
@@
 \paragraph{Time-Triggered Messaging.}
-A \emph{time-triggered} distributed system is one in which nodes are independently clocked, but clock constraints (ensured by periodic clock synchronization) allow the model to appear as if it is executing synchronously~\cite{kopetz}.
+A \emph{time-triggered} distributed system is one in which nodes are independently clocked, but clock constraints allow the model to appear as if it is executing synchronously~\cite{kopetz}.
 
@@
 Messages received outside the window are marked as coming from manifest-faulty senders.
-To implement a time-triggered model, we create three definitions: encoding nondeterministic message delay, and two definitions that are helpers in invariants.
+To implement a time-triggered model, we create three definitions; one encodes nondeterministic message delay, and two are small helpers called by invariants.
 The guards in the relays and receivers are modified to latch messages received outside the specified windows as being manifest faults.
 Additionally, the ASM definition is modified to track the times in the calendar, not just the messages.
-Two new invariants calendar invariants are introduced, stating that the calendar messages are either empty, or their time-stamps fall within the respective message windows.
+Two new calendar invariants are introduced, stating that the calendar messages are either empty, or their time-stamps fall within the respective message windows.
 Then, three invariants classifying faults are relaxed to allow for the possibility of faulty nodes sending benign messages.
@@
 \paragraph{Mid-Value Selection.}
-Our $\OMH(1)$ model leverages majority vote in order to tolerate faults.
-Another choice for the fault masking algorithm used here is mid-value selection (or median).
-This choice is more common in applications involving hardware, signal selection (where signal values cannot be relied on to exactly match even in the absence of faults), or cases where information about congruence is useful.
+Our $\OMH(1)$ model leverages a majority vote in order to tolerate faults.
+Another choice for the fault masking algorithm used is mid-value selection.
+This choice is common in applications involving hardware, signal selection, or cases where information about congruence is useful.
 To implement mid-value selection in our model, we allow messages sent to take values in $\mathbb{R}$ and the receiver transition system is modified in two ways.
@@
 
-As shown in Section~\ref{sec:modular}, modifications to the implementation are modular.
-In particular, any protocol specified in the model can reuse the lemmas associated with the calendar automata and fault-model (insofar as they use the same fault model).
+The invariants and model are surprisingly modular.
+The modifications we explored took at most hours to develop.
+Moreover, most of the invariants do not concern the specific protocol modeled at all, and we hypothesize that for completely different protocols, only the modeling aspects related to the protocol behavior and local node behavior would change.
 

### 2016-12-08 doc/om-paper/paper.tex
pretty bar for agreement, validity.
more edits.

@@
 \newboolean{submission}  %set to true for the submission version
-%\setboolean{submission}{false}
-\setboolean{submission}{true}
+\setboolean{submission}{false}
+%\setboolean{submission}{true}
 \ifthenelse
@@
 fault-tolerant distributed system verification that combining ideas from the literature including calendar automata,
-symbolic fault injection, and abstract state machines observers, and then we apply the framework to model-checking various implementations of the Hybrid Oral Messages algorithm that differ in the fault model, timing model, and local node behavior.
+symbolic fault injection, and abstract transition systems, and then we apply the framework to model-checking various implementations of the Hybrid Oral Messages algorithm that differ in the fault model, timing model, and local node behavior.
 We show that despite being implementation-level models, the verifications are scalable and modular, insofar as isolated changes to an implementation require modular changes to the model and proofs.
@@
 \end{abstract}
-
-\lee{grep for synchronous observers section references}
-\lee{grep for ASM vs abstract transition system}
 
@@
 The framework combines ideas from the literature and folklore to build scalable and modular formal models suitable for infinite-state model-checking, and it reduces the need for ad-hoc abstractions and optimizations while maintaining scalability.
-In Section~\ref{sec:model}, we present the important aspects of the framework, including calendar automata, originally developed by Dutertre and Sorea~\cite{cal}, symbolic fault-injection, and abstract state machines for verification.
+In Section~\ref{sec:model}, we present the important aspects of the framework, including calendar automata, originally developed by Dutertre and Sorea~\cite{cal}, symbolic fault-injection, and abstract transition systems for verification.
 
@@
 In short, the models corresponds closely with an implementation of the algorithm.
-In Section~\ref{sec:byz}, we first describe \OMH, then an implementation of it that uses the Boyer-Moore Fast Majority Vote algorithm~\cite{mjrty}.
+In Section~\ref{sec:byz}, we first describe \OMH, then an implementation of it that uses the Boyer-Moore Fast Majority Vote algorithm (Fast~MJRT)~\cite{mjrty}.
 We then describe a set of modular invariants, such that the invariants only concern specific aspects of the model (e.g., faults, local node behavior, or the passage of time).
@@
 Finally, in Section~\ref{sec:related} we describe related work, and we make concluding remarks in Section~\ref{sec:conclusions}.
+
+The models and experiments reported herein can be found at \url{https://github.com/GaloisInc/mmc-paper}.
 
@@
 Here we describe our formal framework specialized for fault-tolerant distributed systems.
-The framework draws on three principal abstractions: calendar automata, symbolic fault injection, and abstract state machines observers; we describe each below.
+The framework draws on three principal abstractions: calendar automata, symbolic fault injection, and abstract transition sytems; we describe each below.
 
@@
 
-The event of receiving a message initiates a process to update its local state machine and generate additional messages to send.
-When the process is updating its local state machine, the event calendar is paused.
-That is, updating an event $recv(i, j, t, m)$ also includes updating $j$'s state machine.
+The event of receiving a message initiates a process to update its local transition system and generate additional messages to send.
+When the process is updating its local transition system, the event calendar is paused.
+That is, updating an event $recv(i, j, t, m)$ also includes updating $j$'s transition system.
 
@@
 %% \subsection{Synchronous Observers and }\label{sec:sync}
-
-\lee{Cutting synchronous observer section; need to talk about use for lemmas elsewhere.}
 
@@
 \noindent
-\begin{multicols}{2}
-\begin{equation}
-  \tag{Validity}
-    \forall \,i. \quad v \neq \ERR \Rightarrow l_i = v
-\end{equation}\break
-\begin{equation}
-  \tag{Agreement}
-    \forall \,i, j. \quad l_i = l_j
-\end{equation}
-\end{multicols}
+\begin{center}
+\begin{tabular}{l|l}
+\begin{minipage}[t]{0.5\linewidth}
+\begin{equation*}
+%  \tag{Validity}
+    \forall \,i. \quad l_i = v \textnormal{ (Validity)}
+\end{equation*}
+\end{minipage}&
+\begin{minipage}[t]{0.5\linewidth}
+\begin{equation*}
+%  \tag{Agreement}
+    \forall \,i, j. \quad l_i = l_j \textnormal{ (Agreement)}
+\end{equation*}
+\end{minipage}
+\end{tabular}
+\end{center}
 %
@@
 \subsection{Model Sketch}\label{sec:sketch}
-
-\lee{we need to mention MJRTY here or in the modular verifiation section}
 
@@
 
+The relays and receivers explicitly send and receive messages and store them in local buffers as needed. In addition, the receivers implement the Fast~MJRY algorithm~\cite{mjrty}.
+
 Our SAL model defines seven transition systems in total: \texttt{clock}, {\texttt{source}, \texttt{relay} (parametrized over an ID), \texttt{receiver} (parametrized over an ID), \texttt{observer}, \texttt{abstractor}, and \texttt{abstract\_monitor}.
@@
 
-The \texttt{observer} is a synchronous observer~\cite{Rushby:SAS14} that encodes the validity and agreement properties as state machines.
+The \texttt{observer} is a synchronous observer~\cite{Rushby:SAS14} that encodes the validity and agreement properties as synchronously-composed transition systems.
 State variables denoting validity and agreement are set to be false if the receivers have completed their vote but the respective properties do not hold.
 
-Finally, the \texttt{abstractor} and \texttt{abstract\_monitor} encode an abstract state machine observer for the system, as described in Section~\ref{sec:abstract}.
+Finally, the \texttt{abstractor} and \texttt{abstract\_monitor} encode an abstract transition system for the system, as described in Section~\ref{sec:abstract}.
 
@@
     \node[state] (calendar) {calendar (1)};
-    \node[state] (lieutenant) [right=1.75cm of calendar] {\begin{tabular}{c}lieutenant\\state-machine (2)\end{tabular}};
+    \node[state] (lieutenant) [right=1.75cm of calendar] {\begin{tabular}{c}receiver\\local behavior (2)\end{tabular}};
     \node[state] (faults) [below=0.5cm of calendar] {faults (5)};
-    \node[state] (asm) [right=0.5cm of faults] {ASM (1)};
+    \node[state] (asm) [right=0.5cm of faults] {ATS (1)};
     \node[state] (voting) [below=0.5cm of lieutenant] {voting (2)};
@@
   \item \emph{Calendar automata}: Lemmas relating to the calendar automata model.
-These include lemmas such as time being monotonic, channels missing messages if there is no calendar event, and only nodes associated with a calendar event may execute their local state machines.
-  \item \emph{Abstract state machine (ASM)}: Lemmas relating the ASM synchronous observer states to the implementation states.
+These include lemmas such as time being monotonic, channels missing messages if there is no calendar event, and only nodes associated with a calendar event may execute their local transition systems.
+  \item \emph{Abstract transition sytem (ATS)}: Lemmas relating the ATS states to the implementation states.
   \item \emph{Receiver local behavior}: Lemmas describing the modes of behavior of the receivers.
-The major modes of the state machine are receiving messages, then once it has filled its buffer, it votes, and after voting returns the result.
-One additional lemma notes that the messages currently received plus missing messages equals the total number of expected messages.
+The major modes of their behavior are receiving messages, then once it has filled its buffer, it votes, and after voting returns the result.
+An additional lemma notes that the messages currently received plus missing messages equals the total number of expected messages.
   \item \emph{Faults}: Lemmas characterizing the effect of a fault in a single broadcast.
@@
 Another example is that the faults of messages latched by a node in its buffer match the faults ascribed to the sender in the calendar event.
-  \item \emph{Voting}: Lemmas proving that the Fast Majority Vote algorithm implements a majority vote, if one exists.
-These lemmas are nearly verbatim transcriptions from the journal proofs~\cite{mjrty}.
+  \item \emph{Voting}: Lemmas proving that the Fast~MJRTY algorithm implements a majority vote, if one exists.
+These lemmas are nearly verbatim transcriptions from the journal proofs for the algorithm~\cite{mjrty}.
 \end{enumerate}
@@
 Arrows denote dependencies.
-For example, the ASM lemmas depend on both the calendar automata and receiver state-machine lemmas.
+For example, the ATS lemmas depend on both the calendar automata and receiver state-machine lemmas.
 As can be seen, the proof structure is modular.
@@
 Likewise, lemmas about the particular voting algorithm used depend only on the receiver's internal behavior.
-Only the ASM depends on both calendar-specific and local state-machine results, since it is an abstraction of the entire system implementation.
-Recall, however, that the ASM is a convenience for debugging and can be elided.
+Only the ATS depends on both calendar-specific and local state-machine results, since it is an abstraction of the entire system implementation.
+Recall, however, that the ATS is a convenience for debugging and can be elided.
 
@@
   \hline \hline
-  %% No Fault Model                  & receivers & 8 modified & 2 modified & ASM, faults \\
+  %% No Fault Model                  & receivers & 8 modified & 2 modified & ATS, faults \\
   %% \hline
@@
   \hline
-  Time-Triggered Messaging        & source, relays, receivers, ASM & 3 new & 2 modified, 3 modified & calendar, faults \\
+  Time-Triggered Messaging        & source, relays, receivers, ATS & 3 new & 2 modified, 3 modified & calendar, faults \\
   \hline
-  Mid-Value Selection & receivers & 4 new, 3 modified & 2 modified & ASM, voting \\
+  Mid-Value Selection & receivers & 4 new, 3 modified & 2 modified & ATS, voting \\
   \hline
@@
 The guards in the relays and receivers are modified to latch messages received outside the specified windows as being manifest faults.
-Additionally, the ASM definition is modified to track the times in the calendar, not just the messages.
+Additionally, the ATS definition is modified to track the times in the calendar, not just the messages.
 Two new calendar invariants are introduced, stating that the calendar messages are either empty, or their time-stamps fall within the respective message windows.
@@
 Second, a mid-value select function is called on the sorted buffer and the result is stored as the receiver's vote.
-The only invariants needing modification were the ASM definition (to account for the values stored by the new buffer and the relation between it and the main buffer) and the voting invariant.
+The only invariants needing modification were the ATS definition (to account for the values stored by the new buffer and the relation between it and the main buffer) and the voting invariant.
 
@@
 The counterexamples returned by SAL are very useful for strengthening invariants, but tedious to analyze---a model with five relays and two receivers contains 90 state variables, and there are known counterexamples to models that size~\cite{Lincoln-Rushby}.
-Once we developed the ASM synchronously-composed observer, the verification effort was sped up considerably.
+Once we developed the synchronously-composed ATS observer, the verification effort was sped up considerably.
 
@@
 Also in ACL2, an implementation of a circuit design to implement \OM{1} is given~\cite{om-acl2-impl}; the low-level model most closely relates to our level of detail.
-A refinement-based verification approach is used, and \OM{1} is specialized to a four-node system.
-It appears that a general purpose voting algorithm (like Fast MJRTY) is not verified as part of the work.
-It is not clear how difficult it would be to model and verify changes to the implementation in that framework.
+A refinement-based verification approach is used, and \OM{1} is specialized to fixed number of nodes.
 
@@
 Very recently, Jovanovi{\'{c}} and Dutertre use a ``flattened'' high-level model of \OM{1} as a benchmark for IC3 augmented with $k$-induction~\cite{pdr-kind}.
+
+Our work was heavily influenced by previous verifications of fault-tolerant systems in SAL~\cite{Rushby:OM1,cal,timed-systems}.
 
@@
 For example, our work is in SAL's language of guarded commands~\cite{SAL} and needs to be either refined or verified to be equivalent to a software implementation's semantics.
-Another aspect is that behavior related to networking and benign fault diagnosis (e.g., computing a CRC) is are left abstract.
+Another aspect is that behavior related to networking and benign fault diagnosis (e.g., computing a CRC) is left abstract.
 In a software implementation, the $send$ and $recv$ functions would have to implement these behaviors.
@@
  %% A synthesis approach may be suitable and has been used in the past for fault-tolerant clock-synchronization~\cite{ddd}.
-
-\lee{link to all results in github}
-\lee{talk about compositional verification of properties, even though SAL doesn't directly support it---maybe not so important}
-\lee{talk about k-induction vs. PDR---actually I don't think we have much to say}
-\lee{talk about lack of axiomatization in model-checking (multiple rushby bugs), but  tradeoff of deadlock. see proglema paper}
 

### 2016-12-08 doc/om-paper/paper.tex
sec 4 and below edits.

@@
 We present benchmarks in Figure~\ref{fig:benchmark}. The benchmarks were
-performed on a server with Intel Xeon E312xx (Sandy Bridge) CPUs (mid 2011).
+performed on a server with Intel Xeon E312xx (Sandy Bridge) CPUs.
 The table provides execution times in seconds, with a timeout limit of one hour, for verifying the model, given a selected number of relays and receivers.
@@
 
-As a point of comparison, Rushby presents an elegant model of the simpler \OM{m} algorithm, also in SAL, but using BDD-based model-checking~\cite{Rushby:OM1}.
-The model is at the algorithmic level.
-For small values, the verification is much faster, probably due to making only one call to SAL.
+As a point of comparison, Rushby presents an elegant high-level model of \OM{1}, also in SAL, but using BDD-based model-checking~\cite{Rushby:OM1}.
+For small values, the verification is much faster, likely due to making only one call to SAL.
 However, for six relays and two receivers, it takes 449~seconds and timeouts (at one hour) for seven relays and two receivers.
@@
 
-Changing the model to be time-triggered principally requires making the source, relays, and receivers driven explicitly by the passage of time.
+Changing the model to be time-triggered principally requires making the source, relays, and receivers driven explicitly by the passage of time (we do not model clock drift or skew).
 As well, a ``receive window'' is defined at which messages from non-faulty nodes should be received.
 Messages received outside the window are marked as coming from manifest-faulty senders.
-To implement a time-triggered model, we create three definitions; one encodes nondeterministic message delay, and two are small helpers called by invariants.
-The guards in the relays and receivers are modified to latch messages received outside the specified windows as being manifest faults.
-Additionally, the ATS definition is modified to track the times in the calendar, not just the messages.
+The model requires three new definitions to  encode nondeterministic message delay and two are small helper functions.
+The guards in the relays and receivers are modified to latch messages received outside the receive windows as being manifest faults.
+The ATS definition is modified to track the times in the calendar, not just the messages.
 Two new calendar invariants are introduced, stating that the calendar messages are either empty, or their time-stamps fall within the respective message windows.
@@
 
-The invariants and model are surprisingly modular.
-The modifications we explored took at most hours to develop.
-Moreover, most of the invariants do not concern the specific protocol modeled at all, and we hypothesize that for completely different protocols, only the modeling aspects related to the protocol behavior and local node behavior would change.
+The invariants are surprisingly modular.
+One benefit of a model-checking based approach is that it is automated to rerun a proof of a theorem omitting lemmas to see if the proof still holds.
+This allowed us to explore reducing dependencies between invariants related to different aspects of the system.
+
+The modifications to the implementation described in Section~\ref{sec:modular} took at most hours to develop.
+Moreover, most of the invariants do not concern the specific protocol modeled at all, and we hypothesize that for completely different fault-tolerance protocols, only the modeling aspects related to the protocol behavior and local node behavior would change, and the invariant structure would remain modular.
 
@@
 IC3 can even be strengthened using $k$-induciton~\cite{pdr-kind}.
-Tools like Ivy can present graphical counterexamples to help interactively develop invariants~\cite{ivy}.
+%% Tools like Ivy can present graphical counterexamples to help interactively develop invariants~\cite{ivy}.
 
@@
 A refinement-based verification approach is used, and \OM{1} is specialized to fixed number of nodes.
-
-Using automated techniques, Gasc{\'{o}}n and Tiwari use SAT-based bounded synthesis to partially synthesize a variant of Oral Messages that tolerates transient faults, specialized for three or four nodes~\cite{om1-synth}.
 Bokor~\emph{et~al.} describe a message-passing model for synchronous distributed algorithms that is particularly amenable to partial-order reduction for explicit-state model-checking~\cite{Bokor2010}.
@@
 
-Our work was heavily influenced by previous verifications of fault-tolerant systems in SAL~\cite{Rushby:OM1,cal,timed-systems}.
+Moreover, our work is heavily influenced by previous verifications of fault-tolerant and real-time systems in SAL~\cite{Rushby:OM1,cal,timed-systems}.
 
@@
 This work fits within a larger project, in collaboration with Honeywell Labs, to build an \emph{architectural domain-specific language} (ADSL) for specifying and verifying distributed fault-tolerant systems.
-The goal of the language is to synthesize both software and/or hardware implementations as well as formal models for verification.
+The ADSL should be able to synthesize both software and/or hardware implementations as well as formal models for verification.
 Before building such an ADSL, we needed a scalable general formal model to which to compile, leading to the work presented in this paper.
-We hypothesize that an ADSL will make refactoring even easier, and we can generate invariants or invariant templates useful for verification.
+We hypothesize that the ADSL will make refactoring even easier, and we can generate invariants or invariant templates useful for verification.
 Indeed, we have developed a preliminary ADSL that generates C code as well as formal models SRI's Sally~\cite{pdr-kind}, to be described in future work.\footnote{\url{https://github.com/GaloisInc/atom-sally}}
@@
 For example, our work is in SAL's language of guarded commands~\cite{SAL} and needs to be either refined or verified to be equivalent to a software implementation's semantics.
-Another aspect is that behavior related to networking and benign fault diagnosis (e.g., computing a CRC) is left abstract.
-In a software implementation, the $send$ and $recv$ functions would have to implement these behaviors.
+Another aspect is that behavior related to networking, serialization, etc.
+is left abstract, implicit in the $send$ and $recv$ functions.
+
+%% and benign fault diagnosis (e.g., computing a CRC) is left abstract. In a software implementation, the $send$ and $recv$ functions would have to implement these behaviors.
 

### 2016-12-08 doc/om-paper/paper.tex
final-ish proof read of sections 1-3

--- 2016-12-08 doc/om-paper/paper.tex
@@
 \newcommand{\OMH}{\ensuremath{\mathrm{OMH}}\xspace}
-\newcommand{\Id}{\ensuremath{\mathrm{Id}}\xspace}
-\newcommand{\Msg}{\ensuremath{\mathrm{Msg}}\xspace}
-\newcommand{\ERR}{\ensuremath{\mathrm{ERR}}\xspace}
+\newcommand{\Id}{\ensuremath{\mathit{Id}}\xspace}
+\newcommand{\Msg}{\ensuremath{\mathit{Msg}}\xspace}
+\newcommand{\ERR}{\ensuremath{\mathit{ERR}}\xspace}
 \newcommand{\TO}{\ensuremath{\bm{\mathrm{T}}}\xspace}
 \newcommand{\tld}[1]{#1_{*}\xspace}
+\newcommand{\mfa}{\ensuremath{\mathit{mfa}}\xspace}
 
@@
 symbolic fault injection, and abstract transition systems, and then we apply the framework to model-checking various implementations of the Hybrid Oral Messages algorithm that differ in the fault model, timing model, and local node behavior.
-We show that despite being implementation-level models, the verifications are scalable and modular, insofar as isolated changes to an implementation require modular changes to the model and proofs.
+We show that despite being implementation-level models, the verifications are scalable and modular, insofar as isolated changes to an implementation require isolated changes to the model and proofs.
 This work is carried out in SAL model-checker.
@@
 In this paper, we present a framework for modeling fault-tolerant distributed systems, and use that framework to verify several variant implementations of the Byzantine fault-tolerant Hybrid Oral Messages algorithm (\OMH)~\cite{Lincoln-Rushby}.
-The framework combines ideas from the literature and folklore to build scalable and modular formal models suitable for infinite-state model-checking, and it reduces the need for ad-hoc abstractions and optimizations while maintaining scalability.
+The framework combines ideas from the literature and folklore to build scalable and modular formal models suitable for infinite-state model-checking, and it reduces the need for ad-hoc abstractions and optimizations.
 In Section~\ref{sec:model}, we present the important aspects of the framework, including calendar automata, originally developed by Dutertre and Sorea~\cite{cal}, symbolic fault-injection, and abstract transition systems for verification.
@@
 
-We use the framework to verify implementation-level models of \OMH in which message passing is explicit, nodes are not enforced to execute strictly synchronously, and voting is explicit.
+We use the framework to verify implementation-level models of $\OMH$ in which message passing is explicit, nodes are not forced to execute strictly synchronously, and voting is explicit.
 In short, the models corresponds closely with an implementation of the algorithm.
@@
 Here we describe our formal framework specialized for fault-tolerant distributed systems.
-The framework draws on three principal abstractions: calendar automata, symbolic fault injection, and abstract transition sytems; we describe each below.
+The framework draws on three principal abstractions: calendar automata, symbolic fault injection, and abstract transition systems; we describe each below.
 
@@
 Real-time system verification in general-purpose model-checkers requires an explicit formalism of real-time progression.
-Trying to encode real-time clocks directly is difficult; in particular, one must avoid Zeno's paradox in which no progress is made because state transitions simply update real-valued variables by an infinitesimally small amount.
+Trying to encode real-time clocks directly is difficult; in particular, one must avoid Zeno's paradox in which no progress is made because state transitions simply update real-valued variables by an infinite sequence of decreasing amounts whose sum is finite.
 To avoid this problem, Dutetre and Sorea developed \emph{calendar automata}~\cite{cal}, which is itself inspired by event calendars used in discrete-event simulation.
@@
 For now, we do not define events; intuitively, an event is a set of state variables (shortly, we will associate events with messages sent in a distributed system).
-When an event is \emph{enabled}, the transitions over events are enabled; otherwise, the variables stutter and maintain the same value.
+When an event is \emph{enabled}, the transitions over events are enabled; otherwise, the variables stutter (maintain the same value).
 
@@
 
-Let $cal$ be an event calendar and $c_i, c_j \in c$ be calendar events.
-Define an ordering on event calendars such that $c_i \leq c_j$ iff $t_i \leq t_j$, and $min(c) = \{ c_i | \forall c_j \in cal, \, c_i \leq c_j  \}$ are the minimum elements of $cal$.
+Let $cal$ be an event calendar and $c_i, c_j \in cal$ be calendar events.
+Define an ordering on calendar events such that $c_i \leq c_j$ iff $t_i \leq t_j$, and $\min(cal) = \{ c_i | \forall c_j \in cal, \, c_i \leq c_j  \}$ are the minimum elements of $cal$.
 
 Let a transition system $\mathcal{M} = (S, I, \rightarrow)$, be a set of states $S$, a set of initial states $I \subseteq S$, and a transition relation $\rightarrow \subseteq S \times S$.
-We implicitly assume a set of state variables such that a state $\sigma \in S$ is a total function that maps state variables to values.
+We implicitly assume a set of state variables such that each $\sigma \in S$ is a total function that maps state variables to values.
 We distinguish two special state variables in a transition system: (1) $now \in \mathbb{R}$ denotes the current time in the state, and (2) $cal$ is an event calendar.
@@
 
-\item \label{cal:c} In all states, if the current time is strictly less than every calendar event, then the only enabled transition is a \emph{time progress} update: $\forall \sigma \in S$, $\forall (e_i, t_i) \in \sigma(cal)$, if $\sigma(now) < t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma' = \sigma[now := min(c)]$.
+\item \label{cal:c} In all states, if the current time is strictly less than every calendar event, then the only enabled transition is a \emph{time progress} update: $\forall \sigma \in S$, $\forall (e_i, t_i) \in \sigma(cal)$, if $\sigma(now) < t_i$, then $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma' = \sigma[now := \min(cal)]$.
 
@@
 
-Assume processes are indexed from a finite set $\Id \subset \mathbb{N}$.
+Assume processes are indexed from a finite set $\Id$.
 A \emph{channel} from process $i$ to $j$ is an ordered pair $(i,j)$.
@@
 That is, updating an event $recv(i, j, t, m)$ also includes updating $j$'s transition system.
-
-In specifying a system, a \emph{clock} module is synchronously composed (under SAL semantics for synchronous composition~\cite{SAL}) with the system.
-The clock module updates time to the next calendar event, and stutters otherwise.
 
@@
 
-A second difficulty with model-checking fault-tolerant systems in general is that modeling faults requires adding state and nondeterminism.
+A second difficulty with model-checking fault-tolerant systems in general is that modeling faults requires adding state and non-determinism.
 The minimum number of additional states that must be introduced may depend non-obviously on other aspects of the fault model, specific protocol, and system size.
@@
 It  applies to broadcast systems in which a process is expected to broadcast the same value to multiple receivers.
-A \emph{Byzantine} (or \emph{arbitrary} fault is one in which a process that is intended to broadcast the same value to other processes may instead broadcast arbitrary values to different receivers (including no value or the correct value).
+A \emph{Byzantine} (or \emph{arbitrary}) fault is one in which a process that is intended to broadcast the same value to other processes may instead broadcast arbitrary values to different receivers (including no value or the correct value).
 A \emph{symmetric} fault is one in which a process may broadcast the same, but incorrect, value to other processes.
-Finally, a \emph{manifest} (or \emph{benign} fault is one in which a process's broadcast fault is detectable by the receivers; e.g., by performing a cyclic redundancy check (CRC) or because the value arrives outside of a predetermined window.
+Finally, a \emph{manifest} (or \emph{benign}) fault is one in which a process's broadcast fault is detectable by the receivers; e.g., by performing a cyclic redundancy check (CRC) or because the value arrives outside of a predetermined window.
 
@@
 
-Define a set of fault types $$\textnormal{Faults} = \{none, byz, sym, man\}.$$ As in the previous section, let $\Id \subset \mathbb{N}$ be a finite set of process indices, and let the variable $$faults: \Id \rightarrow \textnormal{Faults}$$ range over possible mappings from processes to faults.
+Define a set of fault types $$\textnormal{Faults} = \{none, byz, sym, man\}.$$ As in the previous section, let $\Id$ be a finite set of process indices, and let the variable $$faults: \Id \rightarrow \textnormal{Faults}$$ range over possible mappings from processes to faults.
 
@@
 
-\paragraph{Transient faults.} If the $faults$ mapping is a constant, then faults are permanently but nondeterministically assigned to nodes.
-However, we can easily model \emph{transient faults} in which nodes are faulty temporarily by making $faults$ a state variable that is updated nondeterministically.
+If the $faults$ mapping is a constant, then faults are permanently but non-deterministically assigned to nodes.
+However, we can easily model \emph{transient faults} in which nodes are faulty temporarily by making $faults$ a state variable that is updated non-deterministically.
 Whether we model permanent or transient faults, a \emph{maximum fault assumption} (MFA) describes the maximum number of faults permitted in the system.
-The $faults$ mapping can be nondeterministically updated during execution while satisfying the MFA using a constraint such as $faults \in \{f \, | \, mfa(f) \}$, where the MFA is defined by $mfa()$.
+The $faults$ mapping can be non-deterministically updated during execution while satisfying the MFA using a constraint such as $faults \in \{f \, | \, \mfa(f) \}$, where the MFA is defined by the function $\mfa$.
 
@@
 
-The algorithm is recursively defined:
+The algorithm is recursively defined for $m \ge 0$:
 
 \begin{itemize}
-\item {\bf $\OMH(0)$}: $g$ broadcasts a value to each lieutenant, and the lieutenants return the value received (or $\ERR$).
+\item {\bf $\OMH(0)$}: $g$ broadcasts a value to each lieutenant and the lieutenants return the value received (or $\ERR$).
 \item {\bf $\OMH(m)$, $m > 0$}:
@@
 Then for each $l$, execute $\OMH(m-1)$, assigning $l$ to be the general and $L \setminus \{l\}$ to be the lieutenants.
-$l$ sends $l_v$ or $R$, if $l_v = \ERR$.
+$l$ sends $l_v$, or $R$ if $l_v = \ERR$.
   \item\label{om:three} For each lieutenant $l \in L$, remove all $\ERR$ values received in Step~\ref{om:two} from executing $\OMH(m-1)$.
@@
 \noindent
-In particular, $\OMH(1)$ includes two rounds of broadcast communication, one in which the general broadcasts, and one in which the lieutenants exchange their values.
+In particular, $\OMH(1)$ includes two rounds of broadcast communication: one in which the general broadcasts, and one in which the lieutenants exchange their values.
 
@@
 More formally, Let $l_i, l_j$ denote the outputs of lieutenants $i, j \in L$, respectively, and let $v$ be the value the general broadcasts:
+\vspace{-0.4cm}
 \noindent
@@
 We described a hybrid fault model in Section~\ref{sec:kibitzer}.
-Under that fault model, agreement and validity hold if $2a+2s+b+1 \geq n$, where $n$ is the total number of nodes, $a$ is the number of Byzantine (or asymmetric) faults, $s$ is the number of symmetric faults, and $b$ is the number of benign faults.
+Under that fault model, validity and agreement hold if $2a+2s+b+1 \geq n$, where $n$ is the total number of nodes, $a$ is the number of Byzantine (or asymmetric) faults, $s$ is the number of symmetric faults, and $b$ is the number of benign faults.
 Additionally, the number of rounds $m$ must be greater or equal to the number of Byzantine faults, $a$~\cite{Lincoln-Rushby,hybrid}.
@@
 The first four of these are composed asynchronously, in an intermediate system we label \texttt{system}, and share access to a global calendar consisting of event slots $(\text{message}, \text{time})$, one for each channel in the system.
-The \texttt{clock} transition system is responsible for updating a global variable \texttt{t} representing time according to the rules in Section~\ref{sec:calendar}.
+The \texttt{clock} transition system is responsible for updating a global variable \texttt{t} (called $now$ in Section~\ref{sec:calendar}) representing time according to the rules for calendar automata.
 

### 2016-12-08 doc/om-paper/paper.tex
final-ish proof read for sections 4+

@@
 The use of abstract transition systems not only allows us to scale proofs farther, but also to improve traceability and debugging while developing a model.
-In models like the ones described in Section~\ref{sec:modular} where there are on the order of 100 state variables and counter-example traces could be 30 steps long, the designer can be easily lost trying to identify the essence.
+In models like the ones described in Section~\ref{sec:modular} where there are on the order of 100 state variables and counterexample traces could be 30 steps long, the designer can be easily lost trying to identify the essence.
 In such cases, the values of the abstract predicates can serve to focus the designer's attention on one particular mode of the system where the counter example is taking place.
@@
 For example, let $A_1$ denote the subset of states in which the processor $S$ has not sent a message, let $A_2$ be the subset in which $S$ has broadcast a message, and let $A_3$ be the subset in which $L_1$ and $L_2$ have each received and processed the message.
-The union $A_1 \cup A_2 \cup A_3$ covers the set of reachable states of the system and defines an ASM that is useful for reasoning about its dynamics and understanding counter-example traces.
+The union $A_1 \cup A_2 \cup A_3$ covers the set of reachable states of the system and defines an ASM that is useful for reasoning about its dynamics and understanding counterexample traces.
 
@@
 
-To demonstrate the modularity of the model and verification approach, in this section, we explore variants to the model and report the effort required to implement the modifications and repair the proofs.
+To demonstrate the modularity of the modelling and verification approach, in this section, we explore variants to the model and report the effort required to implement the modifications and repair the proofs.
 The results are summarized in Figure~\ref{fig:effort} and sketched below.
@@
 We hypothesize that the ADSL will make refactoring even easier, and we can generate invariants or invariant templates useful for verification.
-Indeed, we have developed a preliminary ADSL that generates C code as well as formal models SRI's Sally~\cite{pdr-kind}, to be described in future work.\footnote{\url{https://github.com/GaloisInc/atom-sally}}
+Indeed, we have developed a preliminary ADSL that generates C code as well as formal models in SRI's Sally~\cite{pdr-kind}, to be described in future work.\footnote{\url{https://github.com/GaloisInc/atom-sally}}
 

### 2017-05-02 doc/slides/slides-2016-04/state-machine.tex
Rename directories under doc/slides

\documentclass{article}

\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}
\usepackage[latin1]{inputenc}
\usepackage{verbatim}

\begin{document}

\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2.8cm,
                    semithick]

  %\tikzstyle{every state}=[fill=red,draw=none,text=white]
  \tikzstyle{every state}=[fill=red,text=white]

  \node[initial,state]
    (A1) {$a_1$};
  \node[state]
    (A2) [right of=A1, label=above:source done] {$a_2$};
  \node[state]
    (A3) [right of=A2, label=above:1 relay done] {$a_3$};
  \node[state]
    (A3b) [below of=A1, label=below:all relays done] {$a_{3b}$};
  \node[state]
    (A4) [below of=A2, label=below:1 receiver done] {$a_4$};
  \node[state, accepting]
    (A5) [below of=A3, label=below:all done] {$a_4$};

  \path (A1)  edge (A2)
        (A2)  edge (A3)
        (A3)  edge (A3b)
        (A3b) edge (A4)
        (A4)  edge (A5);

\end{tikzpicture}

\end{document}
